### 2022-11-08 main.tex
initial commit

\documentclass[12pt]{article}
\usepackage{graphicx}
\usepackage[margin=2cm]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{tikz}
\usepackage[export]{adjustbox}
\usepackage{indentfirst}
\usepackage{wrapfig}
\usepackage{listings}
\usepackage{color}
\usepackage{enumerate}
\usepackage{amssymb, bm}
\usepackage{csvsimple}
\usepackage{tikz}
\usetikzlibrary{positioning}

\newcommand\tab[1][1cm]{\hspace*{#1}}
\definecolor{mygreen}{rgb}{0,0.6,0}
\definecolor{mygray}{rgb}{0.5,0.5,0.5}
\definecolor{mymauve}{rgb}{0.58,0,0.82}
\lstset{ %
    backgroundcolor=\color{gray!10!white},
  basicstyle=\tiny, %footnotesize,        % the size of the fonts that are used for the code
  breakatwhitespace=false,         % sets if automatic breaks should only happen at whitespace
  breaklines=true,                 % sets automatic line breaking
  captionpos=b,                    % sets the caption-position to bottom
  commentstyle=\color{mygreen},    % comment style
  deletekeywords={...},            % if you want to delete keywords from the given language
  escapeinside={\%*}{*)},          % if you want to add LaTeX within your code
  extendedchars=true,              % lets you use non-ASCII characters; for 8-bits encodings only, does not work with UTF-8
  frame=single,	                   % adds a frame around the code
  keepspaces=true,                 % keeps spaces in text, useful for keeping indentation of code (possibly needs columns=flexible)
  keywordstyle=\color{blue},       % keyword style
  language=Python,                 % the language of the code
  morekeywords={*,...},           % if you want to add more keywords to the set
  numbers=left,                    % where to put the line-numbers; possible values are (none, left, right)
  numbersep=10pt,                   % how far the line-numbers are from the code
  numberstyle=\tiny\color{mygray}, % the style that is used for the line-numbers
  rulecolor=\color{black},         % if not set, the frame-color may be changed on line-breaks within not-black text (e.g. comments (green here))
  showspaces=false,                % show spaces everywhere adding particular underscores; it overrides 'showstringspaces'
  showstringspaces=false,          % underline spaces within strings only
  showtabs=false,                  % show tabs within strings adding particular underscores
  stepnumber=1,                    % the step between two line-numbers. If it's 1, each line will be numbered
  stringstyle=\color{mymauve},     % string literal style
  tabsize=1,	                   % sets default tabsize to 2 spaces
  title=\lstname                   % show the filename of files included with \lstinputlisting; also try caption instead of title
}
\graphicspath{ }
\usetikzlibrary{arrows}

\title{\textbf{COMP0086 Summative Assignment}}
%\author{Jian Shu (James) Wu \\ }
\date{Nov 14, 2022}

\begin{document}
\maketitle
\section*{Question 1}

\begin{enumerate}

\item[(a)] Our sample space for images is $\{0, 1\}^D$, a binary space with $D$ dimensions, the number of pixels in the image.
Thus, picking the exponential family best suited on this sample space is the D-dimensional multivariate Bernoulli distribution which shares the same sample space.
On the other hand, a D-dimensional multivariate Gaussian has the sample space $\mathbb{R}^D$, which does not match the sample space of our data.
To match our data sample space, we would have to define additional mapping between our data and model spaces which adds unnecessary complexity.
Thus it would be inappropriate to model this dataset of images with a multivariate Gaussian.

\item[(b)] For $\mathcal{D} := \{x^{(n)}\}_{n=1}^N$ a data set of N images, the joint likelihood (assuming images are independently and identically distributed) is the product of N D-dimensional multivariate Bernoulli distributions, one for each image:

$$P(\mathcal{D}|\textbf{p}) = \prod_{n=1}^N P(x^{(n)} | \textbf{p})$$

Substituting the D-dimensional multivariate Bernoulli:

$$P(\mathcal{D}|\textbf{p}) = \prod_{n=1}^{N}\prod_{d=1}^{D} p_d^{x_d^{(n)}} (1-p_d)^{1-x_d^{(n)}}$$

Taking the logarithm of this, we get the log likelihood:

$$\mathcal{L}(\mathcal{D}|\textbf{p}) = \sum_{n=1}^{N}\sum_{d=1}^{D} [x_d^{(n)}\log(p_d) +  (1-x_d^{(n)})\log(1-p_d)]$$

Note that since the logarithm of the mean is a monotonic increasing function on $\mathbb{R}_+$, the maximisers and minimisers of the likelihood do not change.

To solve for the maximum likelihood estimate, $\hat p_d$, we can take the derivative of $\mathcal{L}(\mathcal{D}|\textbf{p})$ with respect to $p_d$, the $d^{th}$ element of $\textbf{p}$:

$$\frac{\partial\mathcal{L}(\mathcal{D}|\textbf{p})}{\partial p_d} = \sum_{n=1}^{N} (\frac{x_d^{(n)}}{p_d} -  \frac{1-x_d^{(n)}}{1-p_d})$$
$$\frac{\partial\mathcal{L}(\mathcal{D}|\textbf{p})}{\partial p_d} = \frac{\sum_{n=1}^{N} x_d^{(n)}}{p_d} -  \frac{\sum_{n=1}^{N} (1-x_d^{(n)})}{1-p_d}$$

and set the derivative to zero and solve for $\hat p_d$:

$$\frac{\sum_{n=1}^{N} x_d^{(n)}}{\hat p_d} -  \frac{\sum_{n=1}^{N} (1-x_d^{(n)})}{1-\hat p_d} = 0$$
$$ \sum_{n=1}^{N} x_d^{(n)} - \hat p_d\sum_{n=1}^{N} x_d^{(n)} - \hat p_d  \cdot N + \hat p_d \sum_{n=1}^{N}x_d^{(n)} = 0$$
$$  \hat p_d = \frac{1}{N}\sum_{n=1}^{N} x_d^{(n)}$$

Moreover, the second derivative with respect to $p_d$:

$$\frac{\partial\mathcal{L}(\mathcal{D}|\textbf{p})}{\partial p_d^2} = \frac{-\sum_{n=1}^{N} x_d^{(n)}}{p_d^2} +  \frac{-\sum_{n=1}^{N} (1-x_d^{(n)})}{(1-p_d)^2}$$

For a maximum, we need to show that the second derivative is negative.
Since $x_d^{(n)} \in \{0, 1\}$, in the worst case, of $N=1$, the single pixel must either be white ($\sum_{n=1}^{N} > 0$) or black ($\sum_{n=1}^{N} 1-x_d^{(n)} > 0$) so $\frac{\partial\mathcal{L}(\mathcal{D}|\textbf{p})}{\partial p_d^2} < 0$ will be guaranteed and $\hat p_d$ is a maximum as required for the maximum likelihood estimate.

Because we assume that each pixel is independent (we are taking the product of D one dimensional Bernoulli distributions), we can express the maximum likelihood for $\textbf{p}$ in matrix form as $\hat \textbf{p}$:

$$\hat \textbf{p} = \frac{1}{N}\sum_{n=1}^{N} \textbf{x}^{(n)}$$

%\item[(c)] Assuming independent Beta priors on the parameters $p_d$:
%
%$$P(p_d) = \frac{1}{B(\alpha, \beta)} p^{\alpha-1}_d (1-p_d)^{\beta-1}$$
%
%and $P(\textbf{p}) = \prod_d P(p_d)$ Find the maximum a posteriori (MAP) estimator for \textbf{p}. [5 marks]

\item[(c)] From Bayes' Theorem:

$$P(\textbf{p}|\mathcal{D}) = \frac{ P(\mathcal{D}|\textbf{p})P(\textbf{p})}{P(\mathcal{D})}$$

Taking the logarithm:

$$\mathcal{L}(\textbf{p}|\mathcal{D}) = \mathcal{L}(\mathcal{D}|\textbf{p})+ \mathcal{L}(\textbf{p}) - \mathcal{L}(\mathcal{D})$$

Taking the derivative with respect to $p_d$:

$$\frac{\partial\mathcal{L}(\textbf{p}|\mathcal{D})}{\partial p_d} = \frac{\partial\mathcal{L}(\mathcal{D}|\textbf{p})}{\partial p_d} + \frac{\partial\mathcal{L}(\textbf{p})}{\partial p_d}$$

where $\frac{\partial\mathcal{L}(\mathcal{D})}{\partial p_d}$=0 because it doesn't depend on $p_d$.

We know (b):

$$\frac{\partial\mathcal{L}(\mathcal{D}|\textbf{p})}{\partial p_d} = \frac{\sum_{n=1}^{N} x_d^{(n)}}{p_d} -  \frac{\sum_{n=1}^{N} (1-x_d^{(n)})}{1-p_d}$$

For the second term $\frac{\partial\mathcal{L}(\textbf{p})}{\partial p_d}$, we start with $P(\textbf{p}))$:

$$P(\textbf{p})) = \prod_{d=1}^D P(p_d)$$

Assuming independent Beta priors on the parameters $p_d$:

$$P(\textbf{p}) = \prod_{d=1}^D \frac{1}{B(\alpha, \beta)} p^{\alpha-1}_d (1-p_d)^{\beta-1}$$

Taking the logarithm:

$$\mathcal{L}(\textbf{p}) = \sum_{d=1}^{D} -\log (B(\alpha, \beta)) + (\alpha-1)\log p_d + (\beta-1)\log(1-p_d)$$

Taking the derivative with respect to $p_d$:

$$\frac{\partial\mathcal{L}(\textbf{p})}{\partial p_d} = -\log (B(\alpha, \beta)) + (\alpha-1)\log p_d + (\beta-1)\log(1-p_d))$$

Since we are only concerned with $p_d$, we are only left with a single element of the summation pertaining to $p_d$.

Simplifying:

$$\frac{\partial\mathcal{L}(\textbf{p})}{\partial p_d} = \frac{(\alpha-1)}{p_d} -  \frac{(\beta-1)}{1-p_d}$$

Combining to have an expression for the log posterior derivative $\frac{\partial\mathcal{L}(\textbf{p}|\mathcal{D})}{\partial p_d}$:

$$\frac{\partial\mathcal{L}(\textbf{p}|\mathcal{D})}{\partial p_d} = \frac{\sum_{n=1}^{N} x_d^{(n)}}{p_d} -  \frac{\sum_{n=1}^{N} (1-x_d^{(n)})}{1-p_d} + \frac{(\alpha-1)}{p_d} -  \frac{(\beta-1)}{1-p_d}$$
$$\frac{\partial\mathcal{L}(\textbf{p}|\mathcal{D})}{\partial p_d} = \frac{(\alpha-1) + \sum_{n=1}^{N} x_d^{(n)}}{p_d} -  \frac{(\beta-1) + \sum_{n=1}^{N} (1-x_d^{(n)})}{1-p_d}$$

To find the maximum a posteriori (MAP) estimate $\hat{p_d}$ with $\frac{\partial\mathcal{L}(\textbf{p}|\mathcal{D})}{\partial p_d} = 0$:

$$0 = \frac{(\alpha-1) + \sum_{n=1}^{N} x_d^{(n)}}{\hat{p_d}} -  \frac{(\beta-1) + \sum_{n=1}^{N} (1-x_d^{(n)})}{1-\hat{p_d}}$$
$$0 = (1-\hat{p_d})(\alpha-1) + (1-\hat{p_d})\bigg(\sum_{n=1}^{N} x_d^{(n)}\bigg) -  \hat{p_d}(\beta-1) - \hat{p_d}\bigg(\sum_{n=1}^{N} (1-x_d^{(n)})\bigg)$$
$$0 = (\alpha-\alpha \hat{p_d} + \hat{p_d} - 1) +\bigg(\sum_{n=1}^{N} x_d^{(n)} - \hat{p_d} \sum_{n=1}^{N} x_d^{(n)}\bigg) -  (\hat{p_d}\beta-\hat {p_d}) - \bigg(\hat{p_d}\cdot N - \hat{p_d}\sum_{n=1}^{N}x_d^{(n)}\bigg)$$

Cancelling the $\hat p_d\sum_{n=1}^{N} x_d^{(n)}$ terms:

$$0 = \alpha-\alpha \hat{p_d} + \hat{p_d} - 1 +\sum_{n=1}^{N} x_d^{(n)} -  \hat{p_d}\beta+\hat{p_d} - \hat{p_d}\cdot N$$
$$0 = \hat{p_d}(2-\alpha-\beta-N) + \alpha - 1 +\sum_{n=1}^{N} x_d^{(n)}$$
$$\hat{p_d} =  \frac{\alpha - 1 +\sum_{n=1}^{N} x_d^{(n)}}{(N+\alpha+\beta-2)}$$

To show that this is a maximum, the second derivative is:

$$\frac{\partial^2\mathcal{L}(\textbf{p}|\mathcal{D})}{(\partial p_d)^2} = \frac{(1-\alpha) - \sum_{n=1}^{N} x_d^{(n)}}{(p_d)^2} +  \frac{(1-\beta) - \sum_{n=1}^{N} (1-x_d^{(n)})}{(1-p_d)^2}$$.

For a maximum, we need $\frac{\partial^2\mathcal{L}(\textbf{p}|\mathcal{D})}{(\partial p_d)^2} < 0$ meaning that we need at least one of the strict inequalities $\alpha < 1- \sum_{n=1}^{N} x_d^{(n)}$ or $\beta < 1-\sum_{n=1}^{N} (1-x_d^{(n)})$ to be satisified, where the other can be $\leq$. The Beta distribution requires $\alpha > 0$ and $\beta > 0$ so this requirement will always be satisfied (in the worst case of a single image, either $x_d^{(1)}=1$ or $1-x_d^{(1)}=1$).

Due to independence of our likelihood and priors for each dimension, we can express the maximum a priori for \textbf{p} in matrix form as \hat \textbf{p}:

$$\hat \textbf{p} =   \frac{\alpha - 1 +\sum_{n=1}^{N} \textbf{x}^n}{(N+\alpha+\beta-2)}$$

\newpage
\item[(d\&e)] The Python code for MLE and MAP:
\lstinputlisting[language=Python, caption: MLE and MAP Implementation,]{src/solutions/q1.py}
\newpage
%  \begin{figure}[h]
%  \centering
%  \includegraphics[scale=0.5]{outputs/python/q1/q1d}
%  \caption{ML parameters}
%  \label{fig:1d}
%  \end{figure}
%  \begin{figure}[h]
%  \centering
%  \includegraphics[scale=0.5]{outputs/python/q1/q1e}
%  \caption{MAP parameters}
%  \label{fig:1e}
%  \end{figure}

Displaying the learned parameters:
\begin{figure}[h]
\centering
\begin{minipage}{0.5\textwidth}
  \centering
  \includegraphics[scale=0.5]{outputs/python/q1/q1d}
  \caption{ML parameters}
  \label{fig:1d}
\end{minipage}%
\begin{minipage}{0.5\textwidth}
  \centering
  \includegraphics[scale=0.5]{outputs/python/q1/q1e}
  \caption{MAP parameters}
  \label{fig:1e}
\end{minipage}
\end{figure}

Comparing the equations:

    $$\hat{\textbf{p}}^{MLE} = \frac{1}{N}\sum_{n=1}^{N} \textbf{x}^{(n)}$$

    and

    $$\hat{\textbf{p}}^{MAP}  =  \frac{\alpha - 1 +\sum_{n=1}^{N} \textbf{x}^n}{(N+\alpha+\beta-2)}$$

We see that when the number of data points increases, the MAP estimate approaches $\frac{1}{N}\sum_{n=1}^{N} \textbf{x}^{(n)}$, the MLE. This result makes sense because as our data set gets bigger, are less reliant on our prior.  Moreover, if in our data set of images, a specific pixel in all of these images are white or all black, the MLE for that pixel will be binary. This may not be be representative of our intuitions about images, where there should be some non-zero probability of a pixel being black or white. Thus, introducing an appropriate prior will ensure that the probability of that pixel will never be zero or one. Moreover, a prior can also help ensure numerical stability during calculations. The logarithm of zero is negative infinity, so having probability zero can be problematic for log-likelihoods calculations. A prior can ensure non-zero probabilities. In our case, by using a small Beta distribution prior on each pixel, our parameter values are biased be close to 0.5 and to never be at the extremities 0 and 1. We can see this in our MAP parameters figure where the range of our parameters is smaller than the range for the ML parameters. Also, the MLE has values of zero whereas the MAP does not. Interestingly, when $\alpha=1$ and $\beta=1$, $\hat{\textbf{p}}^{MLE} = \hat{\textbf{p}}^{MAP}$, intuitively this is when the prior is a uniform distribution and so there aren't any biases on the location of $\textbf{p}$ and we recover the MLE. On the other hand, a misspecified prior can be problematic, as the estimated parameters might be skewed by the prior and not properly represent the underlying data generating process and this can be worse than using the MLE.
%In this example, we're using images of different digits as part of our data set. If a particular digit occurs more frequently in the data set, we will see that digit when visualising the MLE or MAP (the digit zero in this case), they simply provides the "average" image pixels. With a set of images with more equal representation of each digit, the parameter visualisation could be an image overlaying the different digits, which is not representative of the "most likely" image in the data. In this case where our data is likely clustered, using a multimodal likelihood is probably a better idea than the unimodal likelihood we've used.

    We can also visualise $\hat{\textbf{p}}^{MAP}-\hat{\textbf{p}}^{MLE}$ to see that for likelihoods greater than 0.5 in the MLE, the MAP has a lower value and for likelihoods less than 0.5, the MAP has a higher value, confirming our intutions.

  \end{figure}
  \begin{figure}[h]
  \centering
  \includegraphics[scale=0.5]{outputs/python/q1/q1e-mle-vs-map}
  \caption{$\hat{\textbf{p}}^{MAP}-\hat{\textbf{p}}^{MLE}$}
  \label{fig:1e}
  \end{figure}

%With MAP, if we have priors skewed towards the parameters for under represented digits in our data, it is more likely to "even out" the estimated parameters. In our case, we used the same Beta prior on all pixels, essentially placing a "uniform shading" on the entire image, so we could not

\end{enumerate}

\newpage

\section*{Question 2}

\begin{enumerate}

\item[(a)] When all D components are generated from a Bernoulli distribution with $p_d = 0.5$, we have the likelihood function for model $M_1$:

$$P(\textbf{x}^{(n)|\textbf{p}^{(1)}}=[0.5, 0.5,...,0.5]^T) = \prod_{d=1}^{D} (0.5)^{x_d^{(n)}} (0.5)^{1-x_d^{(n)}}$$

\item[(b)] When all D components are generated from Bernoulli distributions with unknown, but identical, $p_d$, we have the likelihood function for model $M_2$:

$$P(\textbf{x}^{(n)}|\textbf{p}^{(2)}}=[p_d, p_d, ..., p_d]^T) = \prod_{d'=1}^{D} p_d^{x_{d'}^{(n)}} (1-p_d)^{1-x_{d'}^{(n)}}$$

\item[(c)] When each component is Bernoulli distributed with separate, unknown $p_d$, we have the likelihood function for model $M_3$:

$$P(\textbf{x}^{(n)}|\textbf{p}^{(3)}}=[p_1, p_2, ..., p_D]^T) = \prod_{d=1}^{D} p_d^{x_d^{(n)}} (1-p_d)^{1-x_d^{(n)}}$$

For each model $M_i$, we can marginalise out $P$ to get $P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_i)$:

$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_i) = \int_0^1...\int_0^1 P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|p_d, M_i)P(p_d|M_i) dp_1...dp_D$$

where $d=1,...,D$ and $ \{\textbf{x}^{(n)}\}_{n=1}^{N}$ is our data set.

Given that the prior of any unknown probabilities is uniform, i.e. $P(p_d | M_i) = 1$. We can simplify:

$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_i) = \int_0^1...\int_0^1 P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|p_d, M_i) dp_1...dp_D$$

For $M_1$, we have that all pixels have probability 0.5:

$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_1) = \int_0^1...\int_0^1 \prod_{j=n}^{N} \prod_{d=1}^{D} (0.5)^{x_d^{(n)}} (1-0.5)^{1-{x_d}^{(n)}} d\theta_1...d\theta_D$$

We can remove the integrals and knowing that either $x_d^{(n)}$ or $1-x_d^{(n)}$ will be 1 and the other zero, we can simplify $(0.5)^{x_d^{(n)}} (1-0.5)^{1-{x_d}^{(n)}}$ to $0.5$:

$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_1) = \prod_{j=1}^{N} \prod_{d=1}^D (0.5)$$
$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_1) = (0.5)^{N\cdot D}$$

For $M_2$, we have that all pixels share some probability $p_d$ so we only need to integrate over a single variable $p_d$:

$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_2) = \int_0^1 \prod_{n=1}^{N} \prod_{d'=1}^D p^{x_{d'}^{(n)}} (1-p_d)^{1-{x_{d'}}^{(n)}} dp_d$$

Changing the products to sums:

$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_2) = \int_0^1  p_d^{\sum_{n=1}^{N} \sum_{d'=1}^D x_{d'}^{(n)}} (1-p)^{\sum_{j=1}^{N} \sum_{d'=1}^D 1-{x_{d'}^{(n)}}} d p_d$$

Rewriting:

$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_2) = \int_0^1  p_d^{k} (1- p_{d'=1})^{N\cdot D-k} d p_d$$

where $k =\sum_{n=1}^{N} \sum_{d'=1}^D x_{d'}^{(n)}$.

This integral is the beta function:

$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_2) = \frac{k! (N\cdot D-k)!}{(N\cdot D+1)!}$$

For $M_3$, we need an integral for each $p_d$:

$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_3) = \int_0^1 ... \int_0^1 \prod_{n=1}^{N} \prod_{d=1}^D  p_d^{x_d^{(n)}} (1- p_d)^{1-{x_d}^{(n)}} d p_1 ... d p_D$$

We can separate the integrals to only contain the relevant $p_d$:

$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_3) = \prod_{d=1}^D \Bigg( \int_0^1 \prod_{n=1}^{N}   p_d^{x_d^{(n)}} (1- p_d)^{1-{x_d}^{(n)}} d p_d \Bigg)$$

Changing the products to sums:

$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_3) = \prod_{d=1}^D \Bigg( \int_0^1   p_d^{\sum_{n=1}^{N} x_d^{(n)}} (1- p_d)^{\sum_{n=1}^{N} 1-{x_d}^{(j)}} d p_d \Bigg)$$

In this case, we have the product of integrals where each evaluates to a beta function:

$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_3) = \prod_{d=1}^D \frac{k_d! (N-k_d)!}{(N+1)!}$$

where $k_d = \sum_{n=1}^{N} x_d^{(in}$.

The posterior probability of a model $M_i$ can be expressed:

$$P(M_i |  \{\textbf{x}^{(n)}\}_{n=1}^{N}) = \frac{P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_i)P(M_i)}{P( \{\textbf{x}^{(n)}\}_{n=1}^{N})}$$

We only ave three models, so in this case the normalisation $P(D)$ can be expressed as a sum:

$$P(M_i| \{\textbf{x}^{(n)}\}_{n=1}^{N}) = \frac{P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_i)P(M_i)}{\sum_{i\in \{1,2,3\}} P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_i)P(M_i)}$$

Given that $P(M_i) = \frac{1}{3}$ for all $i \in \{1,2,3\}$:

$$P(M_i| \{\textbf{x}^{(n)}\}_{n=1}^{N}) = \frac{P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_i)}{\sum_{i\in \{1,2,3\}} P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_i)}$$

Calculating the posterior probabilities of each of the three models having generated the data in binarydigits.txt using python we can show the values in the table below:

\begin{center}
\begin{tabular}{l|c}%
 \bfseries $i$ & \bfseries$ P(M_i|\mathcal{D})$% specify table head
\csvreader[head to column names]{outputs/python/q2/q2c.csv}{}% use head of csv as column names
{\\\hline\csvcoli&\csvcolii}% specify your coloumns here
\end{tabular}
\end{center}

% TODO

\newpage
The Python code for calculating the posterior probabilities of the three models:
\lstinputlisting[language=Python]{src/solutions/q2.py}
\end{enumerate}
\newpage
\section*{Question 3}

\begin{enumerate}

%\item[(a)] Write down the likelihood for a model consisting of a mixture of K multivariate Bernoulli
%distributions. Use the parameters $\pi_1, . . . , \pi_K$ to denote the mixing proportions $(0 \leq \pi_k \leq 1; \sum_k \pi_k = 1)$ and arrange the K Bernoulli parameter vectors into a matrix P with elements
%$p_{kd}$ denoting the probability that pixel d takes value 1 under mixture component k. Assume the images are iid under the model, and that the pixels are independent of each other within each component distribution. [5 marks]
%
%
%Just as we can for a mixture of Gaussians, we can formulate this mixture as a latent variable
%model, introducing a discrete hidden variable $s(n) \in \{1, . . . , K \}$ where $P (s(n) = k|\pi) = \pi_k$.

\item[(a)] The likelihood for a model consisting of a mixture of K multivariate Bernoulli distributions can be expressed as the product across $N$ data points:

$$P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\theta) = \prod_{i=1}^{N}P(x_i|\theta)$$

where $\{\textbf{x}^{(n)}\}_{n=1}^{N}$ is our data set with $\textbf{x}^{(n)} \in \mathbb{R}^{D \times 1}$ and $\theta = \{\pi, \mathbf{P}\}$, $\pi = [\pi_1, . . . , \pi_K] \in \mathbb{R}^{K \times 1}$$ our mixing proportions $(0 \leq \pi_k \leq 1; \sum_k \pi_k = 1)$ and $\mathbf{P} \in \mathbb{R}^{D \times K}$ the K Bernoulli parameter vectors with elements $p_{kd}$ denoting the probability that pixel d takes value 1 under mixture component k. We are also the images are iid under the model, and that the pixels are independent of each other within each component distribution.

For each $P(\textbf{x}^{(n)}|\theta)$:

$$P(\textbf{x}^{(n)}|\theta) = \sum_{k=1}^{K} \pi_k \prod_{d=1}^D (p_{kd})^{\textbf{x}^{(n)}_{d}} (1-p_{kd})^{1-\textbf{x}^{(n)}_{d}}$$

The log-likelihood $\mathcal{L}(\textbf{x}^{(n)}|\theta)$ can be expressed in matrix form:

$$\mathcal{L}(\textbf{x}^{(n)}|\theta) = \log \sum_{k=1}^{K}  \pi_k \exp\bigg(\textbf{x}^{(n)}\log(\mathbf{P}_{k}) + (1-\textbf{x}^{(n)})\log(1-\mathbf{P}_{k})\bigg) $$

which can be further vectorised using Python scipy's $logsumexp$ operation.

Moreover, the log-likelihood $\mathcal{L}(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\theta)$ can be expressed:

$$\mathcal{L}(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\theta) = \sum_{i=1}^{N} \Bigg( \log \sum_{k=1}^{K}  \pi_k \exp\bigg(\textbf{x}^{(n)}\log(\mathbf{P}_{k}) + (1-\textbf{x}^{(n)})\log(1-\mathbf{P}_{k})\bigg) \Bigg)$$

\item[(b)] The expression for the responsibility of mixture component k for data vector
$x^{(n)}$, i.e. $r_{nk} \equiv P (s^{(n)} = k|x^{(n)}, \pi, \mathbf{P})$. This computation provides the E-step for an EM algorithm.

We know that:

$$P(A|B) \propto P(B|A)P(A)$$

Thus,

$$P(s^{(n)} = k|x^{(n)}, \pi, \mathbf{P}) \propto  P(x^{(n)} | s^{(n)} = k, \pi, \mathbf{P}) P(s^{(n)} = k|\pi, \mathbf{P})$$

where $s^{(n)} \in \{1,...,K\}$ a discrete latent variable where $P(s^{(n)} = k|x^{(n)} | \pi) = \pi_k$. Note that $P(s^{(n)} = k|x^{(n)} | \pi) = P(s^{(n)} = k|x^{(n)} | \pi, \mathbf{P})$ as $s^{(n)$ isn't dependent on $\mathbf{P}$.

Let $P(s^{(n)} = k|x^{(n)}, \pi, \mathbf{P}) \propto P(s^{(n)}$ be the unnormalised responsibility $\tilde{r}_{nk}$. Using the mixture for component $k$ and the likelihood function of component $k$:

$$\tilde{r}_{nk} = \pi_k \prod_{d=1}^D (p_{kd})^{x^{(n)}_{d}} (1-p_{kd})^{1-x^{(n)}_{d}}$$

Normalising across the components:

$$r_{nk} = \frac{\tilde{r}_{nk}}{\sum_{j=1}^K\tilde{r}_{nj}}$$

and $r_{nk}$, we have calculated $P(s^{(n)} = k|x^{(n)}, \pi, \mathbf{P})$ for the E step.

Moreover,

$$ \log \tilde{r}_{nk} = \log \pi_k + \sum_{d=1}^D \bigg( x^{(n)}_{d}\log(p_{kd}) + (1-x^{(n)}_{d}) \log(1-\exp(\log(p_{kd}))) \bigg)$$

and

$$\log r_{nk} = \log\tilde{r}_{nk} - \log\sum_{j=1}^K\exp(\log\tilde{r}_{nj})$$

which can be vectorised using Python scipy's $logsumexp$ operation.

\item[(c)] We know that the expectation log joint can be expressed:

$$\Bigg\langle \sum_n \log P(x^{(n)}, s^{(n)}| \pi, \mathbf{P})\Bigg\rangle_{q(\{s^{(n)}\})} = \sum_{n=1}^N q(s^{(n)}) \log P(x^{(n)}, s^{(n)}| \pi, \mathbf{P})$$

Let this quantity be $E$.

Filling in the appropriate quantities:

$$ E = \sum_n r_n^T[\log(\pi)+\log(\mathbf{P})^T x^{(n)} + \log(1-\mathbf{P})^T(1-x^{(n)})]$$

where $q(s^{(n)}) = r_n^T$ and $\log P(x^{(n)}, s^{(n)}) = \log(\pi)+\log(\mathbf{P})^T x^{(n)} + \log(1-\mathbf{P})^T(1-x^{(n)})$. Noting that $P(x^{(n)}, s^{(n)}| \pi, \mathbf{P}) = P(x^{(n)}|s^{(n)}) \pi, \mathbf{P}) P(s^{(n)})| \pi, \mathbf{P}) = \tilde{r}_{n}$

To maximise with respect to $\pi$ and $P$ for the M step, we want to take the derivative, set to zero, and solve for $\hat\pi$ and $\hat P$.

For the $k^{th}$ element of $\pi$:

$$\frac{\partial E}{\partial \pi_k} = \sum_n r_{nk} \frac{1}{\pi_m}$$

and

$$\frac{\partial E}{\partial \pi_k} + \lambda = 0$$

where $\lambda$ is a Lagrange multiplier ensuring that the mixing proportions sum to unity.

Thus,

$$\hat \pi_m = \frac{1}{N}\sum_{n}r_{nk}$$

For the $dk^{th}$ element of P:

$$\frac{\partial E}{\partial P_{dk}} = \sum_n r_{nk}\frac{\partial}{\partial P_{dk}}[x^{(n)}_{d}\log P_{dk}+(1-x^{(n)}_{d}) \log(1-P_{dk})]$$

Simplying:

$$\frac{\partial E}{\partial P_{dk}} = \sum_n r_{nk}(\frac{x^{(n)}_{d}}{ P_{dk}}-\frac{1-x^{(n)}_{d}}{1-P_{dk}})$$

Setting the derivative to zero:

$$\frac{\sum_n  x^{(n)}_{d}r_{nk}}{ P_{dk}}-\frac{\sum_n r_{nk}-\sum_n  x^{(n)}_{d}r_{nk}}{1-P_{dk}} = 0$$

Solving for $P_{dk}$:

$$P_{dk}\sum_n r_{nk}-P_{dk}\sum_n x^{(n)}_{d}r_{nk} = \sum_n  x^{(n)}_{d}r_{nk}-P_{dk}\sum_n  x^{(n)}_{d}r_{nk}$$

Thus,

$$\hat{P}_{dk} = \frac{\sum_n x^{(n)}_{d}r_{nk}}{\sum_n r_{nk}}$$

We have the maximizing parameters for the expected log-joint

$$\arg \max_{\pi, P} \Bigg\langle \sum_n \log P(x^{(n)}, s^{(n)}| \pi, P)\Bigg\rangle_{q(\{s^{(n)}\})}$$

thus obtaining an iterative update for the parameters $\pi$ and $P$ in the M-step of EM.

\newpage
\item[(d)] The Python code for the EM algorithm:
\lstinputlisting[language=Python]{src/solutions/q3.py}
\newpage
Run your algorithm on the data set for values of K in {2, 3, 4, 7, 10}. Plot the log likelihood as a function of the iteration number, and display the parameters found. [30 marks]

\item[(e)] Run the algorithm a few times starting from randomly chosen initial conditions. Do you obtain the same solutions (up to permutation)? Does this depend on K? Show the learned probability vectors as images.

Comment on how well the algorithm works, whether it finds good clusters (look at the cluster means and responsibilities and try to interpret them), and how you might improve the model. [10 marks]

\end{enumerate}

\newpage
\section*{Question 5}

\begin{enumerate}

%\item[] \textbf{[70 points] Decrypting Messages with MCMC.} You are given a passage of English text that has been encrypted by remapping each symbol to a (usually) different one. For example,
%
%$$a \rightarrow s$$
%$$b \rightarrow !$$
%$$\langle space \rangle \rightarrow v$$
%$$\cdots$$
%
%Thus a text like ‘a boy. . . ’ might be encrypted by ‘sv!op. . . ’. Assume that the mapping between symbols is one-to-one. The file symbols.txt gives the list of symbols, one per line (the second line is 〈space〉). The file message.txt gives the encrypted message.
%
%Decoding the message by brute force is impossible, since there are 53 symbols and thus 53! possible permutations to try. Instead we will set up a Markov chain Monte Carlo sampler to find modes in the space of permutations.
%
%We model English text, say $s_1s_2\cdots s_n$ where $s_i$ are symbols, as a Markov chain, so that each symbol is independent of the preceding text given only the symbol before:

%$$p(s_1s_2\cdots s_n) = p(s_1)\prod_{i=2}^{n}p(s_i|s_{i-1})$$

\item[(a)] Let $p(\textbf{s}_i,\textbf{s}_{i-1})$ be the probability of the pair of symbols $\textbf{s}_i$ and $\textbf{s}_{i-1}$ occurring together in the text ($\textbf{s}_{i-1}$ followed by $\textbf{s}_i$ where order matters).
We can model $p(\textbf{s}_i,\textbf{s}_{i-1})$ as a multinomial distribution with $N=1$ and $D=53^2$:

$$p(\textbf{s}_i,\textbf{s}_{i-1}) = \frac{1}{s^1! s^2! \cdots s^{53^2}} \prod_{j \in \{1,...53\}, k \in \{1,...53\}} p_{s_i, s_{i-1}}^{t^{s_i, s_{i-1}}}$$

where $t^{s_i, s_{i-1}}$ is an indicator of transition $s_{i-1}$ to $s_i$.
For convenience we will denote $t^{(\alpha, \beta)})$ as the transition from
A multinomial distribution is appropriate because there can only be only one of 53 symbols chosen as $s$ (i.e. a 53 dimensional dice).
Thus, $p(s=s_i)$ represents the probability of the symbol $s_i$ in the text.

We can convert p(s) into exponential family form:

$$p(\textbf{s}) = \frac{1}{s^1! s^2! \cdots s^D} \exp{(\textbf{s}^T \log(\textbf{p}))}$$

where $\textbf{p}$ is the vector of $p_i$'s.
Thus the sufficient statistic is $T(\textbf{s}) = \textbf{s}^T$.

The ML estimate

%Learn the transition statistics of letters and punctuation in English: Download a large text [say the English translation of War and Peace] from the web and estimate the transition probabilities $p(s_i = \alpha|s_{i-1} = \beta) := \Psi(\alpha, \beta)$, as well as the stationary distribution $\lim_{i\rightarrow \infty} p(s_i = \gamma) := \phi(\gamma)$. Assume that the first letter of your text (and also that of the encrypted text provided) is itself sampled from the stationary distribution. Give formulae for the ML estimates of these probabilities as functions of the counts of numbers of occurrences of symbols and pairs of symbols.
%Compute the estimated probabilities. Report the values as a table. [6 marks]

\item[(b)] The latent variables $\sigma(s)$ for different symbols $s$ are not independent.
This is because by choosing an encoding for one symbol $e = \sigma(s)$, the encoding for a second symbol $\sigma(s')$ cannot be $e$.
We have 53 symbols but only 52 degrees of freedom, because once we have defined the encoding for 52 symbols, the encoding for the $53^{rd}$ symbol cannot be chosen.
Thus, there exists a dependence between the symbols for a given $\sigma$.

The joint probability of the encrypted text $e_1 e_2 \cdots e_n$ given $\sigma$:

$$P(e_1, e_2,...,e_n|\sigma) = \psi(\gamma=\sigma^{-1}(e_1))\prod_{i=2}^n \psi(\alpha=\sigma^{-1}(e_i), \beta=\sigma^{-1}(e_{i-1}))$$

because $\sigma$ is the encoding function, mapping to a symbol $s$ into the encoded text as $e$, we require $\sigma^{-1}$ the decoding function mapping the encoded symbol $e$ back to $s$.

\item[(c)] The proposal probability $S(\sigma \rightarrow \sigma')$ depends on the permutations of $\sigma$ and $\sigma'$ because we only choose a proposal $\sigma'$ that differs at \textit{two} spots:

$$\sigma'(s^i) = \sigma(s^j)$$
$$\sigma'(s^j) = \sigma(s^i)$$

for any two symbols $s^i$ and $s^j$ of the 53 possible symbols ($s^i \neq s^j$).

Therefore, if the above doesn't hold for $\sigma'$, $S(\sigma \rightarrow \sigma')=0$, because with our method of choosing proposals, it is not possible to choose $\sigma'$.
At $\sigma$ there are ${53 \choose 2}$ possible proposal $\sigma'$'s with the above property.
Because we are assuming a uniform prior distribution over $\sigma$'s, the transition probability of a $\sigma'$ that satisfies the above property is $S(\sigma \rightarrow \sigma')=\frac{1}{{53 \choose 2}}$.

The MH acceptance probability is given as:

$$A(\sigma \rightarrow \sigma'| \mathcal{D}) = \min \{1, \frac{S(\sigma' \rightarrow \sigma)P(\sigma'|\mathcal{D})}{S(\sigma \rightarrow \sigma')P(\sigma|\mathcal{D})})\}$$

where $S(\sigma \rightarrow \sigma')$ is the conditional transition probability of $\sigma'$ given $\sigma$ and $\mathcal{D}$ is our encrypted text $e_1, e_2,...,e_n$.

$S(\sigma \rightarrow \sigma') = S(\sigma' \rightarrow \sigma)$ for all $\sigma$ and $\sigma'$ that differ only at two spots because the probability in this case will always be $\frac{1}{{53 \choose 2}}$, we can simplify:

$$A(\sigma \rightarrow \sigma'| \mathcal{D}) = \min \{1, \frac{P(\sigma'|\mathcal{D})}{P(\sigma|\mathcal{D})})\}$$

From Bayes' Theorem:

$$P(\sigma|\mathcal{D}) = \frac{P(\mathcal{D}|\sigma)P(\sigma)}{\sum_{\sigma'}P(\mathcal{D}|\sigma')P(\sigma')}$$

We are assuming a uniform prior for $\sigma$, so $P(\sigma)$ is a constant and we can simplify further:

$$A(\sigma \rightarrow \sigma'| \mathcal{D}) = \min \{1, \frac{P(\mathcal{D}|\sigma')}{P(\mathcal{D}|\sigma)})\}$$

This is the acceptance probability for a given proposal $\sigma'$.
The expression for $P(\mathcal{D}|\sigma)$ is $P(e_1, e_2,...,e_n|\sigma)$ described in the previous part.

\newpage
\item[(d)] The Python code for the MH sampler:
\lstinputlisting[language=Python]{src/solutions/q5.py}
\newpage
Implement the MH sampler, and run it on the provided encrypted text. Report the current decryption of the first 60 symbols after every 100 iterations. Your Markov chain should converge to give you a fairly sensible message. (Hint: it may help to initialize your chain intelligently and to try multiple times; in any case, please describe what you did). [30 marks]

TODO

\item[(e)] Note that some $\Psi(\alpha, \beta)$ values may be zero. Does this affect the ergodicity of the chain? If the chain remains ergodic, give a proof; if not, explain and describe how you can restore ergodicity. [5 marks]

TODO

\item[(f)] Analyse this approach to decoding. For instance, would symbol probabilities alone (rather than transitions) be sufficient? If we used a second order Markov chain for English text, what problems might we encounter? Will it work if the encryption scheme allows two symbols to be mapped to the same encrypted value? Would it work for Chinese with $>$ 10000 symbols? [13 marks]

TODO

\end{enumerate}
\newpage
\section*{Question 7}
\begin{enumerate}
\item[(a)] To find the local extrema of the function $f(x,y) = x + 2y$ subject to the constraint $y^2 + xy = 1$, first we define $g(x, y)$:

$$g(x, y) = y^2 + xy - 1$$

where $g(x, y)=0$ is an equivalent representation of the given constraint.

We can therefore construct the optimisation problem:

$$\min_{\textbf{x}} f(\textbf{x})$$

such that $g(\textbf{x})=\textbf{0}$ and $\textbf{x} := [x, y]^T$.

We can calculate $\nabla f(\textbf{x})$:

$$\nabla f(\textbf{x}) = [\frac{\partial}{\partial x} (x+2y), \frac{\partial}{\partial y} (x+2y)]^T$$
$$\nabla f(\textbf{x}) = [1, 2]^T$$

and calculating $\nabla g(\textbf{x})$:

$$\nabla g(\textbf{x}) = [\frac{\partial}{\partial x} (y^2 + xy - 1), \frac{\partial}{\partial y} (y^2 + xy - 1)]^T$$
$$\nabla g(\textbf{x}) = [y, 2y+x]^T$$

Solving the constraint optimisation problem with Lagrange multipliers, we set up the equations:

$$\nabla f(\textbf{x})+\lambda \nabla g(\textbf{x}) = \textbf{0}$$

and

$$g(\textbf{x}) = 0$$

Giving us the three equations:

$$1+\lambda y = 0$$
$$2+\lambda (2y+x)= 0$$
$$y^2 + xy - 1 = 0$$

Substituting $\lambda y = -1$ from the first equation into the second equation:

$$2+2(-1) + x = 0$$

We see that $x=0$.
Solving for $y$ in our third equation with $x=0$:

$$y^2-1=0$$

We see that $y=\pm 1$ and from the first equation $\lambda\mp 1$.

The local extrema are $(x=0, y=1)$ when our $\lambda=-1$ and $(x=0, y=-1)$ when our $\lambda=1$.

\item[(b)] _

\begin{enumerate}
             \item[(i)] Given that $g(a) = \ln(a)$, we want to transform this to the form $f(x,a)=0$:
             $$x = \ln(a)$$
             $$\exp(x) - a = 0$$
             Thus,
             $$f(x, a) = \exp(x) - a$$
  \item[(ii)] We know that for Newton's method's
             $$x_{n+1} = x_{n} - \frac{f(x_n)}{f'(x_n)}$$
             where $f(x_n) = f(x_n, a) = \exp(x_n) - a$

             We can calculate:
             $$f'(x) = \frac{\partial f(x, a)}{\partial x}= \exp(x)$$
             Assuming we can evaluate $\exp(x)$, our update equation:
             $$x_{n+1} = x_{n} - \frac{\exp(x_n) - a}{\exp(x_n)}$$
             Simplifying:
             $$x_{n+1} = x_{n} + \frac{a}{\exp(x_n)} - 1$$
\end{enumerate}

\end{enumerate}
\newpage
\section*{Appendix: main.py}
\lstinputlisting[language=Python]{main.py}
\end{document}

### 2022-11-08 main.py

@@ -1,0 +1,50 @@
+import os
+
+import numpy as np
+
+from src.constants import BINARY_DIGITS_FILE_PATH, OUTPUTS_FOLDER
+from src.solutions import q1, q2, q3
+
+if __name__ == "__main__":
+
+    if not os.path.exists(OUTPUTS_FOLDER):
+        os.makedirs(OUTPUTS_FOLDER)
+
+    x = np.loadtxt(BINARY_DIGITS_FILE_PATH)
+    # Question 1
+    Q1_OUTPUT_FOLDER = os.path.join(OUTPUTS_FOLDER, "q1")
+    if not os.path.exists(Q1_OUTPUT_FOLDER):
+        os.makedirs(Q1_OUTPUT_FOLDER)
+
+    q1.d(
+        x,
+        figure_path=os.path.join(Q1_OUTPUT_FOLDER, "q1d.png"),
+        figure_title="Q1d: Maximum Likelihood Estimate",
+    )
+    q1.e(
+        x,
+        alpha=3,
+        beta=3,
+        figure_path=os.path.join(Q1_OUTPUT_FOLDER, "q1e"),
+        figure_title="Q1e: Maximum A Prior",
+    )
+
+    # Question 2
+    Q2_OUTPUT_FOLDER = os.path.join(OUTPUTS_FOLDER, "q2")
+    if not os.path.exists(Q2_OUTPUT_FOLDER):
+        os.makedirs(Q2_OUTPUT_FOLDER)
+    q2.c(x, table_path=os.path.join(Q2_OUTPUT_FOLDER, "q2c.csv"))
+
+    # Question 3
+    Q3_OUTPUT_FOLDER = os.path.join(OUTPUTS_FOLDER, "q3")
+    if not os.path.exists(Q3_OUTPUT_FOLDER):
+        os.makedirs(Q3_OUTPUT_FOLDER)
+    q3.e(
+        x,
+        number_of_trials=3,
+        ks=[2, 3, 4, 7, 10],
+        epsilon=1e-1,
+        max_number_of_steps=int(1e2),
+        figure_path=os.path.join(Q3_OUTPUT_FOLDER, "q3e"),
+        figure_title="Q3e",
+    )

### 2022-11-08 src/constants.py

@@ -1,0 +1,12 @@
+import os
+
+DATA_FOLDER = "data"
+
+BINARY_DIGITS_FILE_PATH = os.path.join(DATA_FOLDER, "binarydigits.txt")
+MESSAGE_FILE_PATH = os.path.join(DATA_FOLDER, "message.txt")
+SYMBOLS_FILE_PATH = os.path.join(DATA_FOLDER, "symbols.txt")
+TRAINING_TEXT_FILE_PATH = os.path.join(DATA_FOLDER, "war_and_peace.txt")
+
+OUTPUTS_FOLDER = "outputs/python"
+
+DEFAULT_SEED = 0

### 2022-11-08 src/solutions/q1.py

@@ -1,0 +1,59 @@
+import matplotlib.pyplot as plt
+import numpy as np
+
+
+def _compute_maximum_likelihood_estimate(x: np.ndarray) -> np.ndarray:
+    """
+    X: numpy array of shape (N, D)
+    """
+    return np.mean(x, axis=0)
+
+
+def _compute_maximum_a_priori_estimate(
+    x: np.ndarray, alpha: float, beta: float
+) -> np.ndarray:
+    """
+    X: numpy array of shape (N, D)
+    alpha: param of prior distribution
+    beta: param of prior distribution
+    """
+
+    n, _ = x.shape
+    return (alpha - 1 + np.sum(x, axis=0)) / (n + alpha + beta - 2)
+
+
+def d(x, figure_path, figure_title):
+    maximum_likelihood = _compute_maximum_likelihood_estimate(x)
+    plt.figure()
+    plt.imshow(
+        np.reshape(maximum_likelihood, (8, 8)),
+        interpolation="None",
+    )
+    plt.colorbar()
+    plt.axis("off")
+    plt.title(figure_title)
+    plt.savefig(figure_path)
+
+
+def e(x, alpha, beta, figure_path, figure_title):
+    maximum_a_priori = _compute_maximum_a_priori_estimate(x, alpha, beta)
+    plt.figure()
+    plt.imshow(
+        np.reshape(maximum_a_priori, (8, 8)),
+        interpolation="None",
+    )
+    plt.colorbar()
+    plt.axis("off")
+    plt.title(figure_title)
+    plt.savefig(f"{figure_path}.png")
+
+    maximum_likelihood = _compute_maximum_likelihood_estimate(x)
+    plt.figure()
+    plt.imshow(
+        np.reshape(maximum_a_priori - maximum_likelihood, (8, 8)),
+        interpolation="None",
+    )
+    plt.colorbar()
+    plt.axis("off")
+    plt.title(f"MAP vs MLE")
+    plt.savefig(f"{figure_path}-mle-vs-map.png")

### 2022-11-08 src/solutions/q2.py

@@ -1,0 +1,45 @@
+import numpy as np
+import pandas as pd
+from scipy.special import betaln, logsumexp
+
+
+def _log_p_d_given_m1(x):
+    n, d = x.shape
+    return n * d * np.log(0.5)
+
+
+def _log_p_d_given_m2(x):
+    n, d = x.shape
+    k = np.sum(x, axis=0).astype(int)
+    return betaln(np.sum(k) + 1, n * d - np.sum(k) + 1)
+
+
+def _log_p_d_given_m3(x):
+    n, _ = x.shape
+    k = np.sum(x, axis=0).astype(int)
+    return logsumexp(betaln(k + 1, n - k + 1))
+
+
+def c(x, table_path):
+    log_p_d_given_m = np.array(
+        [
+            _log_p_d_given_m1(x),
+            _log_p_d_given_m2(x),
+            _log_p_d_given_m3(x),
+        ]
+    )
+    log_p_m_given_d = log_p_d_given_m - logsumexp(log_p_d_given_m)
+    df = pd.DataFrame(
+        data=np.array(
+            [
+                np.arange(len(log_p_m_given_d)).astype(int) + 1,
+                [f"1E{int(x/np.log(10))}" for x in log_p_m_given_d[:-1]]
+                + [
+                    f"1-{'-'.join([f'(1E{int(x/np.log(10))})' for x in log_p_m_given_d[:-1]])}"
+                ],
+            ]
+        ).T,
+        columns=["Model", "P(M_i|D)"],
+    )
+    df.set_index("Model", inplace=True)
+    df.to_csv(table_path)

### 2022-11-08 src/solutions/q3.py

@@ -1,0 +1,331 @@
+from dataclasses import dataclass
+from typing import List, Tuple
+
+import matplotlib.pyplot as plt
+import numpy as np
+from scipy.special import logsumexp
+
+from src.constants import DEFAULT_SEED
+
+
+@dataclass
+class Theta:
+    """
+    log_pi: the logarithm of the mixing proportions (1, k)
+    log_p_matrix: the logarithm of the probability where the (i,j)th element is the probability that
+                  pixel j takes value 1 under mixture component i (d, k)
+    """
+
+    log_pi: np.ndarray
+    log_p_matrix: np.ndarray
+
+    @property
+    def pi(self):
+        return np.exp(self.log_pi)
+
+    @property
+    def p_matrix(self):
+        d, k = self.log_p_matrix.shape
+        image_dimension = int(np.sqrt(d))
+        return np.exp(self.log_p_matrix).reshape(image_dimension, image_dimension, -1)
+
+    @property
+    def log_one_minus_p_matrix(self) -> np.ndarray:
+        """
+        Compute log(1-P) where P=exp(log_p_matrix)
+        :return: an array of the same shape as log_p_matrix (d, k)
+        """
+        log_of_one = np.zeros(self.log_p_matrix.shape)
+        stacked_sum = np.stack((log_of_one, self.log_p_matrix))
+        weights = np.ones(stacked_sum.shape)
+        weights[1] = -1  # scale p matrix by -1 for subtraction
+        return np.array(logsumexp(stacked_sum, b=weights, axis=0))
+
+    def log_pi_repeated(self, n: int):
+        """
+        Repeats the log_pi vector n times along axis 0
+        :param n: number of repetitions
+        :return: an array of shape (n, k)
+        """
+        return np.repeat(self.log_pi, n, axis=0)
+
+
+def _init_params(k: int, d: int, seed: int = DEFAULT_SEED) -> Theta:
+    """
+    Random initialisation of theta parameters (log_pi and log_p_matrix)
+    :param k: Number of components
+    :param d: Image dimension (number of pixels in a single image)
+    :param seed: seed initialisation for random methods
+    :return: theta: the parameters of the model
+    """
+    np.random.seed(seed)
+    return Theta(
+        log_pi=np.log(np.random.dirichlet(np.ones(k), size=1)),
+        log_p_matrix=np.log(np.random.uniform(low=0, high=1, size=(d, k))),
+    )
+
+
[…]

### 2022-11-08 submission.tex
reorganising file structure

--- 2022-11-08 main.tex
@@
 %  \centering
-%  \includegraphics[scale=0.5]{outputs/python/q1/q1d}
+%  \includegraphics[scale=0.5]{outputs/q1/q1d}
 %  \caption{ML parameters}
@@
 %  \centering
-%  \includegraphics[scale=0.5]{outputs/python/q1/q1e}
+%  \includegraphics[scale=0.5]{outputs/q1/q1e}
 %  \caption{MAP parameters}
@@
   \centering
-  \includegraphics[scale=0.5]{outputs/python/q1/q1d}
+  \includegraphics[scale=0.5]{outputs/q1/q1d}
   \caption{ML parameters}
@@
   \centering
-  \includegraphics[scale=0.5]{outputs/python/q1/q1e}
+  \includegraphics[scale=0.5]{outputs/q1/q1e}
   \caption{MAP parameters}
@@
   \centering
-  \includegraphics[scale=0.5]{outputs/python/q1/q1e-mle-vs-map}
+  \includegraphics[scale=0.5]{outputs/q1/q1e-mle-vs-map}
   \caption{$\hat{\textbf{p}}^{MAP}-\hat{\textbf{p}}^{MLE}$}
@@
  \bfseries $i$ & \bfseries$ P(M_i|\mathcal{D})$% specify table head
-\csvreader[head to column names]{outputs/python/q2/q2c.csv}{}% use head of csv as column names
+\csvreader[head to column names]{outputs/q2/q2c.csv}{}% use head of csv as column names
 {\\\hline\csvcoli&\csvcolii}% specify your coloumns here

### 2022-11-08 src/constants.py

@@ -7,6 +7,6 @@ MESSAGE_FILE_PATH = os.path.join(DATA_FOLDER, "message.txt")
 SYMBOLS_FILE_PATH = os.path.join(DATA_FOLDER, "symbols.txt")
 TRAINING_TEXT_FILE_PATH = os.path.join(DATA_FOLDER, "war_and_peace.txt")
 
-OUTPUTS_FOLDER = "outputs/python"
+OUTPUTS_FOLDER = "outputs"
 
 DEFAULT_SEED = 0

### 2022-11-09 submission.tex
update submission file and add additional plots for q3

@@
 
-$$P(\textbf{p})) = \prod_{d=1}^D P(p_d)$$
+$$P(\textbf{p}) = \prod_{d=1}^D P(p_d)$$
 
@@
 
-$$\frac{\partial\mathcal{L}(\textbf{p})}{\partial p_d} = -\log (B(\alpha, \beta)) + (\alpha-1)\log p_d + (\beta-1)\log(1-p_d))$$
+$$\frac{\partial\mathcal{L}(\textbf{p})}{\partial p_d} = \frac{(\alpha-1)}{p_d} -  \frac{(\beta-1)}{1-p_d}$$
 
 Since we are only concerned with $p_d$, we are only left with a single element of the summation pertaining to $p_d$.
-
-Simplifying:
-
-$$\frac{\partial\mathcal{L}(\textbf{p})}{\partial p_d} = \frac{(\alpha-1)}{p_d} -  \frac{(\beta-1)}{1-p_d}$$
 
@@
   \centering
-  \includegraphics[scale=0.5]{outputs/q1/q1d}
+  \includegraphics[scale=0.4]{outputs/q1/q1d}
   \caption{ML parameters}
@@
   \centering
-  \includegraphics[scale=0.5]{outputs/q1/q1e}
+  \includegraphics[scale=0.4]{outputs/q1/q1e}
   \caption{MAP parameters}
@@
 Comparing the equations:
-
     $$\hat{\textbf{p}}^{MLE} = \frac{1}{N}\sum_{n=1}^{N} \textbf{x}^{(n)}$$
-
     and
-
     $$\hat{\textbf{p}}^{MAP}  =  \frac{\alpha - 1 +\sum_{n=1}^{N} \textbf{x}^n}{(N+\alpha+\beta-2)}$$
 
-We see that when the number of data points increases, the MAP estimate approaches $\frac{1}{N}\sum_{n=1}^{N} \textbf{x}^{(n)}$, the MLE.
-This result makes sense because as our data set gets bigger, are less reliant on our prior.
-Moreover, if in our data set of images, a specific pixel in all of these images are white or all black, the MLE for that pixel will be binary.
-This may not be be representative of our intuitions about images, where there should be some non-zero probability of a pixel being black or white.
-Thus, introducing an appropriate prior will ensure that the probability of that pixel will never be zero or one.
-Moreover, a prior can also help ensure numerical stability during calculations.
-The logarithm of zero is negative infinity, so having probability zero can be problematic for log-likelihoods calculations.
-A prior can ensure non-zero probabilities.
-In our case, by using a small Beta distribution prior on each pixel, our parameter values are biased be close to 0.5 and to never be at the extremities 0 and 1.
-We can see this in our MAP parameters figure where the range of our parameters is smaller than the range for the ML parameters.
-Also, the MLE has values of zero whereas the MAP does not.
-Interestingly, when $\alpha=1$ and $\beta=1$, $\hat{\textbf{p}}^{MLE} = \hat{\textbf{p}}^{MAP}$, intuitively this is when the prior is a uniform distribution and so there aren't any biases on the location of $\textbf{p}$ and we recover the MLE.
-On the other hand, a misspecified prior can be problematic, as the estimated parameters might be skewed by the prior and not properly represent the underlying data generating process and this can be worse than using the MLE.
+As the number of data points increases, the MAP estimate approaches $\frac{1}{N}\sum_{n=1}^{N} \textbf{x}^{(n)}$, the MLE.
+This makes sense because as our data set gets bigger, we are less reliant on our prior.
+Moreover, if a specific pixel in all of the images of our data set are white or all black, the MLE for that pixel will be binary.
+This may not be representative of our intuitions about image pixels, as there should be some non-zero probability of a pixel being black or white.
+By introducing an appropriate prior we can ensure that the probability of that pixel will never be exactly zero or one.
+In our case, with a Beta(3,3) prior on each pixel, our parameter values are biased to be closer to 0.5 and to never be at the extremities 0 and 1.
+We can see this in Figure $\ref{fig:1e}$ where the range of our parameters is smaller than the range of Figure $\ref{fig:1d}$.
+Figure $\ref{fig:1e-comparison}$ visualises $\hat{\textbf{p}}^{MAP}-\hat{\textbf{p}}^{MLE}$ and we can see that for likelihoods greater than 0.5 in the MLE, the MAP has a lower value and for likelihoods less than 0.5, the MAP has a higher value, confirming our intuitions.
+
+  \begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.4]{outputs/q1/q1e-mle-vs-map}
+  \caption{$\hat{\textbf{p}}^{MAP}-\hat{\textbf{p}}^{MLE}$}
+  \label{fig:1e-comparison}
+  \end{figure}
+
+Priors can also help ensure numerical stability during calculations.
+The logarithm of zero is negative infinity, so having if the MLE is zero it can be problematic for log-likelihoods calculations whereas MAP can ensure non-zero probabilities.
+We can see in Figure $\ref{fig:1d}$ and Figure $\ref{fig:1e}$ that the MLE has parameter values of zero whereas the MAP does not.
+Interestingly, when $\alpha=1$ and $\beta=1$, $\hat{\textbf{p}}^{MLE} = \hat{\textbf{p}}^{MAP}$.
+Intuitively this is when the prior is a uniform distribution and so there aren't any biases on the location of $\textbf{p}$ and we recover the MLE.
+
+On the other hand, a mis-specified prior can be problematic, as the estimated parameters might be skewed by the prior and not properly represent the underlying data generating process, this can result in parameter estimates that are worse than using the MLE.
 %In this example, we're using images of different digits as part of our data set.
@@
 In this case where our data is likely clustered, using a multimodal likelihood is probably a better idea than the unimodal likelihood we've used.
-
-    We can also visualise $\hat{\textbf{p}}^{MAP}-\hat{\textbf{p}}^{MLE}$ to see that for likelihoods greater than 0.5 in the MLE, the MAP has a lower value and for likelihoods less than 0.5, the MAP has a higher value, confirming our intutions.
-
-  \end{figure}
-  \begin{figure}[h]
-  \centering
-  \includegraphics[scale=0.5]{outputs/q1/q1e-mle-vs-map}
-  \caption{$\hat{\textbf{p}}^{MAP}-\hat{\textbf{p}}^{MLE}$}
-  \label{fig:1e}
-  \end{figure}
 
@@
 \newpage
-
 \section*{Question 2}
@@
 
-$$P(\textbf{x}^{(n)|\textbf{p}^{(1)}}=[0.5, 0.5,...,0.5]^T) = \prod_{d=1}^{D} (0.5)^{x_d^{(n)}} (0.5)^{1-x_d^{(n)}}$$
+$$P(\textbf{x}^{(n)|\textbf{p}^{(1)}}=[0.5, 0.5,...,0.5]^T, M_1) = \prod_{d=1}^{D} (0.5)^{x_d^{(n)}} (0.5)^{1-x_d^{(n)}}$$
 
@@
 
-$$P(\textbf{x}^{(n)}|\textbf{p}^{(2)}}=[p_d, p_d, ..., p_d]^T) = \prod_{d'=1}^{D} p_d^{x_{d'}^{(n)}} (1-p_d)^{1-x_{d'}^{(n)}}$$
+$$P(\textbf{x}^{(n)}|\textbf{p}^{(2)}}=[p_d, p_d, ..., p_d]^T, M_2) = \prod_{d'=1}^{D} p_d^{x_{d'}^{(n)}} (1-p_d)^{1-x_{d'}^{(n)}}$$
 
@@
 
-$$P(\textbf{x}^{(n)}|\textbf{p}^{(3)}}=[p_1, p_2, ..., p_D]^T) = \prod_{d=1}^{D} p_d^{x_d^{(n)}} (1-p_d)^{1-x_d^{(n)}}$$
+$$P(\textbf{x}^{(n)}|\textbf{p}^{(3)}}=[p_1, p_2, ..., p_D]^T, M_3) = \prod_{d=1}^{D} p_d^{x_d^{(n)}} (1-p_d)^{1-x_d^{(n)}}$$
 
-For each model $M_i$, we can marginalise out $P$ to get $P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_i)$:
+For each model $M_i$, we can marginalise out $\textbf{p}^{(i)}}$ to get $P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_i)$:
 
@@
 
-We only ave three models, so in this case the normalisation $P(D)$ can be expressed as a sum:
+We only have three models, so in this case the normalisation $P(\{\textbf{x}^{(n)}\}_{n=1}^{N})$ can be expressed as a sum:
 
@@
 \begin{tabular}{l|c}%
- \bfseries $i$ & \bfseries$ P(M_i|\mathcal{D})$% specify table head
+ \bfseries $i$ & \bfseries$ P(M_i|\{\textbf{x}^{(n)}\}_{n=1}^{N})$% specify table head
 \csvreader[head to column names]{outputs/q2/q2c.csv}{}% use head of csv as column names
@@
 
-% TODO
+We can see that for models specified to have the same parameter value for all pixels is very unlikely under the given data set.
+This makes sense because it is specifying models where the image is essentially blank, which is not reflective of the data.
+Moreover, $M_1$ specifies a specific value of 0.5 for all the parameters whereas $M_2$ specifies any value for all the parameters as long as it's the same.
+So the model $M_1$ is a subset of the models specified in $M_2$ and we can see this reflected in our probabilities when $P(M_2|\{\textbf{x}^{(n)}\}_{n=1}^{N}) > P(M_1|\{\textbf{x}^{(n)}\}_{n=1}^{N})$.
 
@@
 \item[(a)] The likelihood for a model consisting of a mixture of K multivariate Bernoulli distributions can be expressed as the product across $N$ data points:
-
 $$P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\theta) = \prod_{i=1}^{N}P(x_i|\theta)$$
-
 where $\{\textbf{x}^{(n)}\}_{n=1}^{N}$ is our data set with $\textbf{x}^{(n)} \in \mathbb{R}^{D \times 1}$ and $\theta = \{\pi, \mathbf{P}\}$, $\pi = [\pi_1, .
@@
 .
-, \pi_K] \in \mathbb{R}^{K \times 1}$$ our mixing proportions $(0 \leq \pi_k \leq 1; \sum_k \pi_k = 1)$ and $\mathbf{P} \in \mathbb{R}^{D \times K}$ the K Bernoulli parameter vectors with elements $p_{kd}$ denoting the probability that pixel d takes value 1 under mixture component k.
+, \pi_K] \in \mathbb{R}^{K \times 1}$ our mixing proportions $(0 \leq \pi_k \leq 1; \sum_k \pi_k = 1)$ and $\mathbf{P} \in \mathbb{R}^{D \times K}$ the K Bernoulli parameter vectors with elements $p_{kd}$ denoting the probability that pixel d takes value 1 under mixture component k.
 We are also the images are iid under the model, and that the pixels are independent of each other within each component distribution.
@@
 For each $P(\textbf{x}^{(n)}|\theta)$:
-
 $$P(\textbf{x}^{(n)}|\theta) = \sum_{k=1}^{K} \pi_k \prod_{d=1}^D (p_{kd})^{\textbf{x}^{(n)}_{d}} (1-p_{kd})^{1-\textbf{x}^{(n)}_{d}}$$
-
 The log-likelihood $\mathcal{L}(\textbf{x}^{(n)}|\theta)$ can be expressed in matrix form:
-
 $$\mathcal{L}(\textbf{x}^{(n)}|\theta) = \log \sum_{k=1}^{K}  \pi_k \exp\bigg(\textbf{x}^{(n)}\log(\mathbf{P}_{k}) + (1-\textbf{x}^{(n)})\log(1-\mathbf{P}_{k})\bigg) $$
-
 which can be further vectorised using Python scipy's $logsumexp$ operation.
@@
 Moreover, the log-likelihood $\mathcal{L}(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\theta)$ can be expressed:
-
 $$\mathcal{L}(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\theta) = \sum_{i=1}^{N} \Bigg( \log \sum_{k=1}^{K}  \pi_k \exp\bigg(\textbf{x}^{(n)}\log(\mathbf{P}_{k}) + (1-\textbf{x}^{(n)})\log(1-\mathbf{P}_{k})\bigg) \Bigg)$$
@@
 We know that:
-
 $$P(A|B) \propto P(B|A)P(A)$$
-
 Thus,
-
 $$P(s^{(n)} = k|x^{(n)}, \pi, \mathbf{P}) \propto  P(x^{(n)} | s^{(n)} = k, \pi, \mathbf{P}) P(s^{(n)} = k|\pi, \mathbf{P})$$
-
 where $s^{(n)} \in \{1,...,K\}$ a discrete latent variable where $P(s^{(n)} = k|x^{(n)} | \pi) = \pi_k$.
@@
 Let $P(s^{(n)} = k|x^{(n)}, \pi, \mathbf{P}) \propto P(s^{(n)}$ be the unnormalised responsibility $\tilde{r}_{nk}$. Using the mixture for component $k$ and the likelihood function of component $k$:
-
 $$\tilde{r}_{nk} = \pi_k \prod_{d=1}^D (p_{kd})^{x^{(n)}_{d}} (1-p_{kd})^{1-x^{(n)}_{d}}$$
-
 Normalising across the components:
-
 $$r_{nk} = \frac{\tilde{r}_{nk}}{\sum_{j=1}^K\tilde{r}_{nj}}$$
-
 and $r_{nk}$, we have calculated $P(s^{(n)} = k|x^{(n)}, \pi, \mathbf{P})$ for the E step.
@@
 Moreover,
-
 $$ \log \tilde{r}_{nk} = \log \pi_k + \sum_{d=1}^D \bigg( x^{(n)}_{d}\log(p_{kd}) + (1-x^{(n)}_{d}) \log(1-\exp(\log(p_{kd}))) \bigg)$$
-
 and
-
 $$\log r_{nk} = \log\tilde{r}_{nk} - \log\sum_{j=1}^K\exp(\log\tilde{r}_{nj})$$
-
-which can be vectorised using Python scipy's $logsumexp$ operation.
+which can be vectorised as $\log \textbf{r}_{n}$ using Python scipy's $logsumexp$ operation.
 
 \item[(c)] We know that the expectation log joint can be expressed:
+$$\Bigg\langle \sum_n \log P(x^{(n)}, s^{(n)}| \pi, \mathbf{P})\Bigg\rangle_{q(\{s^{(n)}\})} = \sum_{n=1}^N q(s^{(n)}) \log P(x^{(n)}, s^{(n)}| \pi, \mathbf{P})$$
+Let this quantity be $E$. Each term of $E$ can be expressed:
+$$q(s^{(n)}) = \textbf{r}_{n}$$
+and
+$$\log P(x^{(n)}, s^{(n)}| \pi, \mathbf{P}) = \log[P(x^{(n)}| s^{(n)}, \pi, \mathbf{P})P(s^{(n)}| \pi, \mathbf{P})]$$
+which is the vectorised version of $\log\tilde{r}_{nk}$ from part (b) so:
+$$\log P(x^{(n)}, s^{(n)}| \pi, \mathbf{P}) =\log(\pi)+\log(\mathbf{P})^T x^{(n)} + \log(1-\mathbf{P})^T(1-x^{(n)})$$
+Combining:
+$$ E = \sum_n \textbf{r}_{n}^T[\log(\pi)+\log(\mathbf{P})^T x^{(n)} + \log(1-\mathbf{P})^T(1-x^{(n)})]$$
 
-$$\Bigg\langle \sum_n \log P(x^{(n)}, s^{(n)}| \pi, \mathbf{P})\Bigg\rangle_{q(\{s^{(n)}\})} = \sum_{n=1}^N q(s^{(n)}) \log P(x^{(n)}, s^{(n)}| \pi, \mathbf{P})$$
-
-Let this quantity be $E$.
-
-Filling in the appropriate quantities:
-
-$$ E = \sum_n r_n^T[\log(\pi)+\log(\mathbf{P})^T x^{(n)} + \log(1-\mathbf{P})^T(1-x^{(n)})]$$
-
-where $q(s^{(n)}) = r_n^T$ and $\log P(x^{(n)}, s^{(n)}) = \log(\pi)+\log(\mathbf{P})^T x^{(n)} + \log(1-\mathbf{P})^T(1-x^{(n)})$.
-Noting that $P(x^{(n)}, s^{(n)}| \pi, \mathbf{P}) = P(x^{(n)}|s^{(n)}) \pi, \mathbf{P}) P(s^{(n)})| \pi, \mathbf{P}) = \tilde{r}_{n}$
-
-To maximise with respect to $\pi$ and $P$ for the M step, we want to take the derivative, set to zero, and solve for $\hat\pi$ and $\hat P$.
+To maximise with respect to $\pi$ and $\mathbf{P}$ for the M step, we want to take the derivative, set to zero, and solve for $\hat\pi$ and $\hat P$.
 
 For the $k^{th}$ element of $\pi$:
+$$\frac{\partial E}{\partial \pi_k} = \sum_n r_{nk} \frac{1}{\pi_k}$$
 
-$$\frac{\partial E}{\partial \pi_k} = \sum_n r_{nk} \frac{1}{\pi_m}$$
+The second derivative:
 
-and
+$$\frac{\partial E}{(\partial \pi_k)^2} = \sum_n r_{nk} \frac{-1}{(\pi_k)^2}$$
 
+is always negative because $r_{nk}\geq 0$, $\sum_{n}r_{nk}=1$, \pi_{k}\geq 0$, and $\sum_{n}\pi_{k}=1$, ensuring a maximum in the next step.
+
+We can calculate the maximiser:
 $$\frac{\partial E}{\partial \pi_k} + \lambda = 0$$
-
 where $\lambda$ is a Lagrange multiplier ensuring that the mixing proportions sum to unity.
@@
 Thus,
+$$\hat \pi_k = \frac{\sum_{n}r_{nk}}{N}$$
 
-$$\hat \pi_m = \frac{1}{N}\sum_{n}r_{nk}$$
+For the $dk^{th}$ element of $\mathbf{P}$:
 
-For the $dk^{th}$ element of P:
+$$\frac{\partial E}{\partial \mathbf{P}_{dk}} = \sum_n r_{nk}\frac{\partial}{\partial \mathbf{P}_{dk}}[x^{(n)}_{d}\log \mathbf{P}_{dk}+(1-x^{(n)}_{d}) \log(1-\mathbf{P}_{dk})]$$
 
-$$\frac{\partial E}{\partial P_{dk}} = \sum_n r_{nk}\frac{\partial}{\partial P_{dk}}[x^{(n)}_{d}\log P_{dk}+(1-x^{(n)}_{d}) \log(1-P_{dk})]$$
+Simplifying:
 
-Simplying:
+$$\frac{\partial E}{\partial \mathbf{P}_{dk}} = \sum_n r_{nk}(\frac{x^{(n)}_{d}}{ \mathbf{P}_{dk}}-\frac{1-x^{(n)}_{d}}{1-\mathbf{P}_{dk}})$$
 
-$$\frac{\partial E}{\partial P_{dk}} = \sum_n r_{nk}(\frac{x^{(n)}_{d}}{ P_{dk}}-\frac{1-x^{(n)}_{d}}{1-P_{dk}})$$
+Similar to Question 1, we can see that taking the derivative, the term in the brackets will always be less than zero and with $r_{nk}\geq 0$ and $\sum_{n}r_{nk}=1$, the second derivative will always be negative.
+This ensures that we have a maximum in the next step.
 
@@
 
-$$\frac{\sum_n  x^{(n)}_{d}r_{nk}}{ P_{dk}}-\frac{\sum_n r_{nk}-\sum_n  x^{(n)}_{d}r_{nk}}{1-P_{dk}} = 0$$
+$$\frac{\sum_n  x^{(n)}_{d}r_{nk}}{ \mathbf{P}_{dk}}-\frac{\sum_n r_{nk}-\sum_n  x^{(n)}_{d}r_{nk}}{1-\mathbf{P}_{dk}} = 0$$
 
-Solving for $P_{dk}$:
+Solving for $\hat{\mathbf{P}}_{dk}$:
 
-$$P_{dk}\sum_n r_{nk}-P_{dk}\sum_n x^{(n)}_{d}r_{nk} = \sum_n  x^{(n)}_{d}r_{nk}-P_{dk}\sum_n  x^{(n)}_{d}r_{nk}$$
+$$\hat{\mathbf{P}}_{dk}\sum_n r_{nk}-\hat{\mathbf{P}}_{dk}\sum_n x^{(n)}_{d}r_{nk} = \sum_n  x^{(n)}_{d}r_{nk}-\hat{\mathbf{P}}_{dk}\sum_n  x^{(n)}_{d}r_{nk}$$
 
@@
 
-$$\hat{P}_{dk} = \frac{\sum_n x^{(n)}_{d}r_{nk}}{\sum_n r_{nk}}$$
+$$\hat{\mathbf{P}}_{dk} = \frac{\sum_n x^{(n)}_{d}r_{nk}}{\sum_n r_{nk}}$$
 
@@
 
-$$\arg \max_{\pi, P} \Bigg\langle \sum_n \log P(x^{(n)}, s^{(n)}| \pi, P)\Bigg\rangle_{q(\{s^{(n)}\})}$$
+$$\arg \max_{\pi, \mathbf{P}} \Bigg\langle \sum_n \log P(x^{(n)}, s^{(n)}| \pi, \mathbf{P})\Bigg\rangle_{q(\{s^{(n)}\})}$$
 
-thus obtaining an iterative update for the parameters $\pi$ and $P$ in the M-step of EM.
+thus obtaining an iterative update for the parameters $\pi$ and $\mathbf{P}$ in the M-step of EM.
+
+For numerical stability, we can compute the maximisation step for the MAP of $\mathbf{P}$,$\hat{\mathbf{P}}_{dk}^{MAP}$:
+$$\frac{\partial E'}{\partial \mathbf{P}_{dk}} = 0$$
+where
+$$E' = \sum_{n=1}^N q(s^{(n)}) \log P(\mathbf{P} | \pi, x^{(n)}, s^{(n)})$$
+and
+$$\log P(\mathbf{P} | \pi, x^{(n)}, s^{(n)}) = \log P( x^{(n)}, s^{(n)} | \pi, \mathbf{P} ) + \log P(\mathbf{P}) - \log P(x^{(n)}, s^{(n)} | \pi)$$
+Assuming the same independent Beta prior on each pixel of each component:
+$$\log P(\mathbf{P}) = \sum_{k=1}^K \sum_{d=1}^D -\log (B(\alpha, \beta)) + (\alpha-1)\log \mathbf{P}_{dk} + (\beta-1)\log(1-\mathbf{P}_{dk})$$
+and
+$$\frac{\partial \log P(\mathbf{P})}{\partial \mathbf{P}_{dk}} = \frac{(\alpha-1)}{\mathbf{P}_{dk}} -  \frac{(\beta-1)}{1-\mathbf{P}_{dk}}$$`
+Thus, the derivative can be expressed as:
+$$\frac{\partial E'}{\partial \mathbf{P}_{dk}} =  \sum_n \Bigg(r_{nk} \bigg(\frac{\partial \log P( x^{(n)}, s^{(n)} | \pi, \mathbf{P} )}{\partial \mathbf{P}_{dk}} + \frac{\partial \log P(\mathbf{P})}{\partial \mathbf{P}_{dk}} \bigg) \Bigg)$$
+Substituting the appropriate expressions:
+$$\frac{\partial E'}{\partial \mathbf{P}_{dk}} =  \sum_n \Bigg(r_{nk} \bigg(\frac{x^{(n)}_{d}}{ \mathbf{P}_{dk}}-\frac{1-x^{(n)}_{d}}{1-\mathbf{P}_{dk}} + \frac{(\alpha-1)}{\mathbf{P}_{dk}} -  \frac{(\beta-1)}{1-\mathbf{P}_{dk}} \bigg) \Bigg)$$
+Simplifying:
+$$\frac{\partial E'}{\partial \mathbf{P}_{dk}} =  \frac{(\alpha-1)+\sum_n r_{nk} x^{(n)}_{d}}{ \mathbf{P}_{dk}}-\frac{(\beta-1)+\sum_n r_{nk} (1-x^{(n)}_{d})}{1-\mathbf{P}_{dk}}$$
+
+This form is very similar to Question 2 (c).
+By a similar argument, we can see that the second derivative will always be negative (additionally $r_{nk}\geq 0$ and $\sum_n r_{nk}=1$ ), ensuring a maximum in the calculation of the next step.
+
+Setting $\frac{\partial E'}{\partial \mathbf{P}_{dk}}=0$ we can calculate $\hat{\mathbf{P}}_{dk}^{MAP}$:
+
+$$\hat{\mathbf{P}}_{dk}^{MAP} = \frac{\alpha - 1 +\sum_n r_{nk} x^{(n)}_{d}}{(N+\alpha+\beta-2)}$$
 
 \newpage
-\item[(d)] The Python code for the EM algorithm:
+\item[(d)] Plotting the posterior likelihood as a function of the iteration number:
+
+\begin{figure}[h]
+\centering
+\includegraphics[scale=0.35]{outputs/q3/q3e-0-log-pos}
+\caption{Log Likelihood vs Iteration Number}
+\label{fig:3d-log-like}
+\end{figure}
+
+where $epsilon$ is the stopping condition for the posterior posterior converges.
+
+\newpage
+
+Displaying the parameters found for $K$ in $\{2, 3, 4, 7, 10\}$:
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.3]{outputs/q3/q3e-0-initialised-p}
+  \caption{Randomly initialised parameters}
+  \label{fig:3d-initialised-p}
+\end{figure}
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.3]{outputs/q3/q3e-0-optimised-p}
+  \caption{EM optimised parameters}
+  \label{fig:3d-optimised-p}
+\end{figure}
+
+%\begin{minipage}{0.5\textwidth}
+
+%\end{minipage}%
+%\begin{minipage}{0.5\textwidth}
+%  \centering
+%  \includegraphics[scale=0.3]{outputs/q3/q3e-0-optimised-p}
+%  \caption{EM optimised parameters}
+%  \label{fig:3d-optimised-p}
+%\end{minipage}
+%\end{figure}
+
+\newpage
+The Python code for the EM algorithm:
 \lstinputlisting[language=Python]{src/solutions/q3.py}
 \newpage
-Run your algorithm on the data set for values of K in {2, 3, 4, 7, 10}. Plot the log likelihood as a function of the iteration number, and display the parameters found. [30 marks]
+\item[(e)] Running the algorithm a few times starting from randomly chosen initial conditions and visualising the parameters:
 
-\item[(e)] Run the algorithm a few times starting from randomly chosen initial conditions.
-Do you obtain the same solutions (up to permutation)?
-Does this depend on K?
-Show the learned probability vectors as images.
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.3]{outputs/q3/q3e-0-optimised-p}
+  \caption{EM optimised parameters: Trial 0}
+  \label{fig:3e-initialised-p-trial-0}
+\end{figure}
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.3]{outputs/q3/q3e-1-optimised-p}
+  \caption{EM optimised parameters: Trial 1}
+  \label{fig:3e-initialised-p-trial-1}
+\end{figure}
+\newpage
 
-Comment on how well the algorithm works, whether it finds good clusters (look at the cluster means and responsibilities and try to interpret them), and how you might improve the model. [10 marks]
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.3]{outputs/q3/q3e-2-optimised-p}
+  \caption{EM optimised parameters: Trial 2}
+  \label{fig:3e-initialised-p-trial-2}
+\end{figure}
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.3]{outputs/q3/q3e-3-optimised-p}
+  \caption{EM optimised parameters: Trial 3}
+  \label{fig:3e-initialised-p-trial-3}
+\end{figure}
+
+%\begin{figure}[h]
+%\centering
+%\begin{minipage}{0.5\textwidth}
+%  \centering
+%  \includegraphics[scale=0.5]{outputs/q3/q3e-0-optimised-p}
+%  \caption{EM optimised parameters: Trial 0}
+%  \label{fig:3e-initialised-p-trial-0}
+%\end{minipage}%
+%\begin{minipage}{0.5\textwidth}
+%  \centering
+%  \includegraphics[scale=0.5]{outputs/q3/q3e-1-optimised-p}
+%  \caption{EM optimised parameters: Trial 1}
+%  \label{fig:3e-initialised-p-trial-1}
+%\end{minipage}
+%\end{figure}
+%\begin{figure}[h]
+%\centering
+%\begin{minipage}{0.5\textwidth}
+%  \centering
+%  \includegraphics[scale=0.5]{outputs/q3/q3e-2-optimised-p}
+%  \caption{EM optimised parameters: Trial 2}
+%  \label{fig:3e-initialised-p-trial-2}
+%\end{minipage}%
+%\begin{minipage}{0.5\textwidth}
+%  \centering
+%  \includegraphics[scale=0.5]{outputs/q3/q3e-3-optimised-p}
+%  \caption{EM optimised parameters: Trial 3}
+%  \label{fig:3e-initialised-p-trial-3}
+%\end{minipage}
+%\end{figure}
+
+For smaller k, we can visually see that we obtain very similar solutions (a 7 and a 0 for $k=2$).
+However for higher K, we see that this may not always be the case.
+For Trial 0 of $k=10$, we have two 0's whereas in Trial 1 we have three 0's.
+Interestingly, the zeros are all different, representing different variants of the written digit (i.e.
+a slanted zero, a slightly slanted zero, and a symmetric zero).
+
+\newpage
+
+Moreover, looking at the responsibilities of each mixture component, we can see that when k is relatively small they are relatively evenly distributed.
+However for $k=7$ and especially $k=10$, we can see some components have very small or zero probability (i.e.
+$\pi_1$ and $\pi_2$ of trial 2).
+The corresponding parameter visualisations will essentially never be utilised because no probability is assigned to the component.
+This can be verified when we perform a TSNE visualisation of the responsibility vector for each of the images.
+We can see that for large k, qualitatively the number of clusters no longer matches the k value, indicating that some clusters are redundant.
+For example for $k=7$ and $k=10$ we can only qualitatively see four or five clusters with TSNE.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.35]{outputs/q3/q3e-0-tsne}
+  \caption{TSNE Visualisation of Image responsibilities: Trial 0}
+  \label{fig:3e-tsne-0}
+\end{figure}
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.35]{outputs/q3/q3e-1-tsne}
+  \caption{TSNE Visualisation of Image responsibilities: Trial 1}
+  \label{fig:3e-tsne-1}
+\end{figure}
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.35]{outputs/q3/q3e-2-tsne}
+  \caption{TSNE Visualisation of Image responsibilities: Trial 2}
+  \label{fig:3e-tsne-2}
+\end{figure}
+\newpage
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.35]{outputs/q3/q3e-3-tsne}
+  \caption{TSNE Visualisation of Image responsibilities: Trial 3}
+  \label{fig:3e-tsne-3}
+\end{figure}
+
+Improvements to the model could include searching for an optimal $k$ by maximising the log posterior with regularisation on the magnitude of $k$ to balancing maximising log posterior with minimising model complexity.
+Additionally, adding a prior on the responsibility components can be helpful to ensure non-zero mixing components unlike the components visualised here.
+This can help promote more meaningful clusters as $k$ increases.
 

### 2022-11-09 main.py

@@ -41,7 +41,7 @@ if __name__ == "__main__":
         os.makedirs(Q3_OUTPUT_FOLDER)
     q3.e(
         x,
-        number_of_trials=3,
+        number_of_trials=4,
         ks=[2, 3, 4, 7, 10],
         epsilon=1e-1,
         max_number_of_steps=int(1e2),

### 2022-11-09 src/solutions/q3.py

@@ -4,6 +4,7 @@ from typing import List, Tuple
 import matplotlib.pyplot as plt
 import numpy as np
 from scipy.special import logsumexp
+from sklearn.manifold import TSNE
 
 from src.constants import DEFAULT_SEED
 
@@ -125,21 +126,7 @@ def _compute_log_pi_hat(log_responsibility: np.ndarray) -> np.ndarray:
     :return: an array of the maximised log mixing proportions (1, k)
     """
     n, _ = log_responsibility.shape
-    summed_responsibilities = logsumexp(log_responsibility, axis=0)
-
-    alpha = 2
-    beta = 2
-    summed_responsibilities = logsumexp(
-        np.stack(
-            (
-                (alpha - 1) * np.ones(summed_responsibilities.shape),
-                summed_responsibilities,
-            ),
-            axis=0,
-        ),
-        axis=0,
-    )
-    return (summed_responsibilities - np.log(n + alpha + beta - 2)).reshape(1, -1)
+    return (logsumexp(log_responsibility, axis=0) - np.log(n)).reshape(1, -1)
 
 
 def _compute_log_p_matrix_hat(
@@ -159,37 +146,37 @@ def _compute_log_p_matrix_hat(
         log_responsibility[:, np.newaxis, :], d, axis=1
     )  # (n, d, k)
 
-    log_p_matrix_unnormalised = logsumexp(
+    log_p_matrix_unnormalised_likelihood = logsumexp(
         log_responsibility_repeated, b=x_repeated, axis=0
     )  # (d, k)
-    log_p_matrix_normaliser = np.array(
+    log_p_matrix_normaliser_likelihood = np.array(
         logsumexp(log_responsibility_repeated, axis=0)
     )  # (d, k)
 
     alpha = 2
     beta = 2
-    log_p_matrix_unnormalised = logsumexp(
+    log_p_matrix_unnormalised_posterior = logsumexp(
         np.stack(
             (
-                (alpha - 1) * np.ones(log_p_matrix_unnormalised.shape),
-                log_p_matrix_unnormalised,
+                (alpha - 1) * np.ones(log_p_matrix_unnormalised_likelihood.shape),
+                log_p_matrix_unnormalised_likelihood,
             ),
             axis=0,
         ),
         axis=0,
     )
-    log_p_matrix_normaliser = logsumexp(
+    log_p_matrix_normaliser_posterior = logsumexp(
         np.stack(
             (
[…]

### 2022-11-10 submission.tex
major edits

@@
 \usetikzlibrary{positioning}
-
+\usepackage{lscape}
 \newcommand\tab[1][1cm]{\hspace*{#1}}
@@
 
-\item[(a)] Our sample space for images is $\{0, 1\}^D$, a binary space with $D$ dimensions, the number of pixels in the image.
-Thus, picking the exponential family best suited on this sample space is the D-dimensional multivariate Bernoulli distribution which shares the same sample space.
+\item[(a)] Our sample space for images is $\{0, 1\}^D$, where each of our D dimensions can only take binary values, D being the number of pixels in the image.
+The exponential family best suited on this sample space is the D-dimensional multivariate Bernoulli distribution because it shares the same sample space.
 On the other hand, a D-dimensional multivariate Gaussian has the sample space $\mathbb{R}^D$, which does not match the sample space of our data.
-To match our data sample space, we would have to define additional mapping between our data and model spaces which adds unnecessary complexity.
+To match our data sample space, we might have to define an additional mapping between our data and model spaces, adding unnecessary complexity.
 Thus it would be inappropriate to model this dataset of images with a multivariate Gaussian.
 
-\item[(b)] For $\mathcal{D} := \{x^{(n)}\}_{n=1}^N$ a data set of N images, the joint likelihood (assuming images are independently and identically distributed) is the product of N D-dimensional multivariate Bernoulli distributions, one for each image:
-
-$$P(\mathcal{D}|\textbf{p}) = \prod_{n=1}^N P(x^{(n)} | \textbf{p})$$
-
+\item[(b)] For $\{\textbf{x}^{(n)}\}_{n=1}^N$, a data set of N images, the joint likelihood (assuming images are independently and identically distributed) is the product of N D-dimensional multivariate Bernoulli distributions, one for each image:
+$$P(\{\textbf{x}^{(n)}\}_{n=1}^N|\textbf{p}) = \prod_{n=1}^N P(\textbf{x}^{(n)} | \textbf{p})$$
 Substituting the D-dimensional multivariate Bernoulli:
-
-$$P(\mathcal{D}|\textbf{p}) = \prod_{n=1}^{N}\prod_{d=1}^{D} p_d^{x_d^{(n)}} (1-p_d)^{1-x_d^{(n)}}$$
-
+$$P(\{\textbf{x}^{(n)}\}_{n=1}^N|\textbf{p}) = \prod_{n=1}^{N}\prod_{d=1}^{D} p_d^{x_d^{(n)}} (1-p_d)^{1-x_d^{(n)}}$$
 Taking the logarithm of this, we get the log likelihood:
-
-$$\mathcal{L}(\mathcal{D}|\textbf{p}) = \sum_{n=1}^{N}\sum_{d=1}^{D} [x_d^{(n)}\log(p_d) +  (1-x_d^{(n)})\log(1-p_d)]$$
-
-Note that since the logarithm of the mean is a monotonic increasing function on $\mathbb{R}_+$, the maximisers and minimisers of the likelihood do not change.
-
-To solve for the maximum likelihood estimate, $\hat p_d$, we can take the derivative of $\mathcal{L}(\mathcal{D}|\textbf{p})$ with respect to $p_d$, the $d^{th}$ element of $\textbf{p}$:
-
-$$\frac{\partial\mathcal{L}(\mathcal{D}|\textbf{p})}{\partial p_d} = \sum_{n=1}^{N} (\frac{x_d^{(n)}}{p_d} -  \frac{1-x_d^{(n)}}{1-p_d})$$
-$$\frac{\partial\mathcal{L}(\mathcal{D}|\textbf{p})}{\partial p_d} = \frac{\sum_{n=1}^{N} x_d^{(n)}}{p_d} -  \frac{\sum_{n=1}^{N} (1-x_d^{(n)})}{1-p_d}$$
-
+$$\mathcal{L}(\{\textbf{x}^{(n)}\}_{n=1}^N|\textbf{p}) = \sum_{n=1}^{N}\sum_{d=1}^{D} [x_d^{(n)}\log(p_d) +  (1-x_d^{(n)})\log(1-p_d)]$$
+Note that since the logarithm is a monotonically increasing function on $\mathbb{R}_+$, the maximisers and minimisers of the likelihood do not change.
+Thus, to solve for the maximum likelihood estimate, $\hat p_d$, we can take the derivative of $\mathcal{L}(\{\textbf{x}^{(n)}\}_{n=1}^N|\textbf{p})$ with respect to $p_d$, the $d^{th}$ element of $\textbf{p}$:
+$$\frac{\partial\mathcal{L}(\{\textbf{x}^{(n)}\}_{n=1}^N|\textbf{p})}{\partial p_d} = \sum_{n=1}^{N} (\frac{x_d^{(n)}}{p_d} -  \frac{1-x_d^{(n)}}{1-p_d})$$
+$$\frac{\partial\mathcal{L}(\{\textbf{x}^{(n)}\}_{n=1}^N|\textbf{p})}{\partial p_d} = \frac{\sum_{n=1}^{N} x_d^{(n)}}{p_d} -  \frac{\sum_{n=1}^{N} (1-x_d^{(n)})}{1-p_d}$$
 and set the derivative to zero and solve for $\hat p_d$:
-
 $$\frac{\sum_{n=1}^{N} x_d^{(n)}}{\hat p_d} -  \frac{\sum_{n=1}^{N} (1-x_d^{(n)})}{1-\hat p_d} = 0$$
@@
 $$  \hat p_d = \frac{1}{N}\sum_{n=1}^{N} x_d^{(n)}$$
+Moreover, the second derivative with respect to $p_d$:
+$$\frac{\partial\mathcal{L}(\{\textbf{x}^{(n)}\}_{n=1}^N|\textbf{p})}{\partial p_d^2} = \frac{-\sum_{n=1}^{N} x_d^{(n)}}{p_d^2} +  \frac{-\sum_{n=1}^{N} (1-x_d^{(n)})}{(1-p_d)^2}$$
+For a maximum, we need to show that the second derivative is negative.
+Since $x_d^{(n)} \in \{0, 1\}$, in the worst case, of $N=1$, the single pixel $x_d^{(1)}$ must either be white ($\sum_{n=1}^{N} x_d^{(n)}> 0$) or black ($\sum_{n=1}^{N} 1-x_d^{(n)} > 0$) with the other being zero,  $\frac{\partial\mathcal{L}(\{\textbf{x}^{(n)}\}_{n=1}^N|\textbf{p})}{\partial p_d^2} < 0$ will be guaranteed and $\hat p_d$ is a maximum as required for the maximum likelihood estimate.
 
-Moreover, the second derivative with respect to $p_d$:
-
-$$\frac{\partial\mathcal{L}(\mathcal{D}|\textbf{p})}{\partial p_d^2} = \frac{-\sum_{n=1}^{N} x_d^{(n)}}{p_d^2} +  \frac{-\sum_{n=1}^{N} (1-x_d^{(n)})}{(1-p_d)^2}$$
-
-For a maximum, we need to show that the second derivative is negative.
-Since $x_d^{(n)} \in \{0, 1\}$, in the worst case, of $N=1$, the single pixel must either be white ($\sum_{n=1}^{N} > 0$) or black ($\sum_{n=1}^{N} 1-x_d^{(n)} > 0$) so $\frac{\partial\mathcal{L}(\mathcal{D}|\textbf{p})}{\partial p_d^2} < 0$ will be guaranteed and $\hat p_d$ is a maximum as required for the maximum likelihood estimate.
-
-Because we assume that each pixel is independent (we are taking the product of D one dimensional Bernoulli distributions), we can express the maximum likelihood for $\textbf{p}$ in matrix form as $\hat \textbf{p}$:
-
-$$\hat \textbf{p} = \frac{1}{N}\sum_{n=1}^{N} \textbf{x}^{(n)}$$
+Because we assume that each pixel is independent (we are taking the product of D one dimensional Bernoulli distributions), we can express the maximum likelihood for $\textbf{p}$ in vectorised form as $\hat{\textbf{p}}^{MLE}$:
+$$\hat{\textbf{p}}^{MLE} = \frac{1}{N}\sum_{n=1}^{N} \textbf{x}^{(n)}$$
 
@@
 \item[(c)] From Bayes' Theorem:
+$$P(\textbf{p}|\{\textbf{x}^{(n)}\}_{n=1}^N) = \frac{ P(\{\textbf{x}^{(n)}\}_{n=1}^N|\textbf{p})P(\textbf{p})}{P(\{\textbf{x}^{(n)}\}_{n=1}^N)}$$
+Taking the logarithm:
+$$\mathcal{L}(\textbf{p}|\{\textbf{x}^{(n)}\}_{n=1}^N) = \mathcal{L}(\{\textbf{x}^{(n)}\}_{n=1}^N|\textbf{p})+ \mathcal{L}(\textbf{p}) - \mathcal{L}(\{\textbf{x}^{(n)}\}_{n=1}^N)$$
+Taking the derivative with respect to $p_d$:
+$$\frac{\partial\mathcal{L}(\textbf{p}|\{\textbf{x}^{(n)}\}_{n=1}^N)}{\partial p_d} = \frac{\partial\mathcal{L}(\{\textbf{x}^{(n)}\}_{n=1}^N|\textbf{p})}{\partial p_d} + \frac{\partial\mathcal{L}(\textbf{p})}{\partial p_d}$$
+where $\frac{\partial\mathcal{L}(\{\textbf{x}^{(n)}\}_{n=1}^N)}{\partial p_d}$=0 because it doesn't depend on $p_d$.
 
-$$P(\textbf{p}|\mathcal{D}) = \frac{ P(\mathcal{D}|\textbf{p})P(\textbf{p})}{P(\mathcal{D})}$$
-
+We know from (b):
+$$\frac{\partial\mathcal{L}(\{\textbf{x}^{(n)}\}_{n=1}^N|\textbf{p})}{\partial p_d} = \frac{\sum_{n=1}^{N} x_d^{(n)}}{p_d} -  \frac{\sum_{n=1}^{N} (1-x_d^{(n)})}{1-p_d}$$
+For the second term $\frac{\partial\mathcal{L}(\textbf{p})}{\partial p_d}$, we start with $P(\textbf{p}))$, assuming each pixel to have an independent prior:
+$$P(\textbf{p}) = \prod_{d=1}^D P(p_d)$$
+Assuming a Beta prior on each $p_d$:
+$$P(\textbf{p}) = \prod_{d=1}^D \frac{1}{B(\alpha, \beta)} p^{\alpha-1}_d (1-p_d)^{\beta-1}$$
 Taking the logarithm:
-
-$$\mathcal{L}(\textbf{p}|\mathcal{D}) = \mathcal{L}(\mathcal{D}|\textbf{p})+ \mathcal{L}(\textbf{p}) - \mathcal{L}(\mathcal{D})$$
-
+$$\mathcal{L}(\textbf{p}) = \sum_{d=1}^{D} -\log (B(\alpha, \beta)) + (\alpha-1)\log p_d + (\beta-1)\log(1-p_d)$$
 Taking the derivative with respect to $p_d$:
-
-$$\frac{\partial\mathcal{L}(\textbf{p}|\mathcal{D})}{\partial p_d} = \frac{\partial\mathcal{L}(\mathcal{D}|\textbf{p})}{\partial p_d} + \frac{\partial\mathcal{L}(\textbf{p})}{\partial p_d}$$
-
-where $\frac{\partial\mathcal{L}(\mathcal{D})}{\partial p_d}$=0 because it doesn't depend on $p_d$.
-
-We know (b):
-
-$$\frac{\partial\mathcal{L}(\mathcal{D}|\textbf{p})}{\partial p_d} = \frac{\sum_{n=1}^{N} x_d^{(n)}}{p_d} -  \frac{\sum_{n=1}^{N} (1-x_d^{(n)})}{1-p_d}$$
-
-For the second term $\frac{\partial\mathcal{L}(\textbf{p})}{\partial p_d}$, we start with $P(\textbf{p}))$:
-
-$$P(\textbf{p}) = \prod_{d=1}^D P(p_d)$$
-
-Assuming independent Beta priors on the parameters $p_d$:
-
-$$P(\textbf{p}) = \prod_{d=1}^D \frac{1}{B(\alpha, \beta)} p^{\alpha-1}_d (1-p_d)^{\beta-1}$$
-
-Taking the logarithm:
-
-$$\mathcal{L}(\textbf{p}) = \sum_{d=1}^{D} -\log (B(\alpha, \beta)) + (\alpha-1)\log p_d + (\beta-1)\log(1-p_d)$$
-
-Taking the derivative with respect to $p_d$:
-
 $$\frac{\partial\mathcal{L}(\textbf{p})}{\partial p_d} = \frac{(\alpha-1)}{p_d} -  \frac{(\beta-1)}{1-p_d}$$
-
 Since we are only concerned with $p_d$, we are only left with a single element of the summation pertaining to $p_d$.
 
-Combining to have an expression for the log posterior derivative $\frac{\partial\mathcal{L}(\textbf{p}|\mathcal{D})}{\partial p_d}$:
-
-$$\frac{\partial\mathcal{L}(\textbf{p}|\mathcal{D})}{\partial p_d} = \frac{\sum_{n=1}^{N} x_d^{(n)}}{p_d} -  \frac{\sum_{n=1}^{N} (1-x_d^{(n)})}{1-p_d} + \frac{(\alpha-1)}{p_d} -  \frac{(\beta-1)}{1-p_d}$$
-$$\frac{\partial\mathcal{L}(\textbf{p}|\mathcal{D})}{\partial p_d} = \frac{(\alpha-1) + \sum_{n=1}^{N} x_d^{(n)}}{p_d} -  \frac{(\beta-1) + \sum_{n=1}^{N} (1-x_d^{(n)})}{1-p_d}$$
-
-To find the maximum a posteriori (MAP) estimate $\hat{p_d}$ with $\frac{\partial\mathcal{L}(\textbf{p}|\mathcal{D})}{\partial p_d} = 0$:
-
+Combining, we have have an expression for $\frac{\partial\mathcal{L}(\textbf{p}|\{\textbf{x}^{(n)}\}_{n=1}^N)}{\partial p_d}$:
+$$\frac{\partial\mathcal{L}(\textbf{p}|\{\textbf{x}^{(n)}\}_{n=1}^N)}{\partial p_d} = \frac{\sum_{n=1}^{N} x_d^{(n)}}{p_d} -  \frac{\sum_{n=1}^{N} (1-x_d^{(n)})}{1-p_d} + \frac{(\alpha-1)}{p_d} -  \frac{(\beta-1)}{1-p_d}$$
+$$\frac{\partial\mathcal{L}(\textbf{p}|\{\textbf{x}^{(n)}\}_{n=1}^N)}{\partial p_d} = \frac{(\alpha-1) + \sum_{n=1}^{N} x_d^{(n)}}{p_d} -  \frac{(\beta-1) + \sum_{n=1}^{N} (1-x_d^{(n)})}{1-p_d}$$
+To find the maximum a posteriori (MAP) estimate $\hat{p_d}$ set $\frac{\partial\mathcal{L}(\textbf{p}|\{\textbf{x}^{(n)}\}_{n=1}^N)}{\partial p_d} = 0$ and solve:
 $$0 = \frac{(\alpha-1) + \sum_{n=1}^{N} x_d^{(n)}}{\hat{p_d}} -  \frac{(\beta-1) + \sum_{n=1}^{N} (1-x_d^{(n)})}{1-\hat{p_d}}$$
@@
 $$0 = (\alpha-\alpha \hat{p_d} + \hat{p_d} - 1) +\bigg(\sum_{n=1}^{N} x_d^{(n)} - \hat{p_d} \sum_{n=1}^{N} x_d^{(n)}\bigg) -  (\hat{p_d}\beta-\hat {p_d}) - \bigg(\hat{p_d}\cdot N - \hat{p_d}\sum_{n=1}^{N}x_d^{(n)}\bigg)$$
-
 Cancelling the $\hat p_d\sum_{n=1}^{N} x_d^{(n)}$ terms:
-
 $$0 = \alpha-\alpha \hat{p_d} + \hat{p_d} - 1 +\sum_{n=1}^{N} x_d^{(n)} -  \hat{p_d}\beta+\hat{p_d} - \hat{p_d}\cdot N$$
@@
 $$\hat{p_d} =  \frac{\alpha - 1 +\sum_{n=1}^{N} x_d^{(n)}}{(N+\alpha+\beta-2)}$$
-
 To show that this is a maximum, the second derivative is:
-
-$$\frac{\partial^2\mathcal{L}(\textbf{p}|\mathcal{D})}{(\partial p_d)^2} = \frac{(1-\alpha) - \sum_{n=1}^{N} x_d^{(n)}}{(p_d)^2} +  \frac{(1-\beta) - \sum_{n=1}^{N} (1-x_d^{(n)})}{(1-p_d)^2}$$.
-
-For a maximum, we need $\frac{\partial^2\mathcal{L}(\textbf{p}|\mathcal{D})}{(\partial p_d)^2} < 0$ meaning that we need at least one of the strict inequalities $\alpha < 1- \sum_{n=1}^{N} x_d^{(n)}$ or $\beta < 1-\sum_{n=1}^{N} (1-x_d^{(n)})$ to be satisified, where the other can be $\leq$.
+$$\frac{\partial^2\mathcal{L}(\textbf{p}|\{\textbf{x}^{(n)}\}_{n=1}^N)}{(\partial p_d)^2} = \frac{(1-\alpha) - \sum_{n=1}^{N} x_d^{(n)}}{(p_d)^2} +  \frac{(1-\beta) - \sum_{n=1}^{N} (1-x_d^{(n)})}{(1-p_d)^2}$$.
+For a maximum, we need $\frac{\partial^2\mathcal{L}(\textbf{p}|\{\textbf{x}^{(n)}\}_{n=1}^N)}{(\partial p_d)^2} < 0$ meaning that we need at least one of the strict inequalities $\alpha < 1- \sum_{n=1}^{N} x_d^{(n)}$ or $\beta < 1-\sum_{n=1}^{N} (1-x_d^{(n)})$ to be satisified, where the other can be $\leq$.
 The Beta distribution requires $\alpha > 0$ and $\beta > 0$ so this requirement will always be satisfied (in the worst case of a single image, either $x_d^{(1)}=1$ or $1-x_d^{(1)}=1$).
 
-Due to independence of our likelihood and priors for each dimension, we can express the maximum a priori for \textbf{p} in matrix form as \hat \textbf{p}:
-
-$$\hat \textbf{p} =   \frac{\alpha - 1 +\sum_{n=1}^{N} \textbf{x}^n}{(N+\alpha+\beta-2)}$$
+Due to independence of our likelihood and priors for each dimension, we can express the maximum a priori for \textbf{p} in vectorised form as $\hat{\textbf{p}}^{MAP}$:
+$$\hat{\textbf{p}}^{MAP} =   \frac{\alpha - 1 +\sum_{n=1}^{N} \textbf{x}^n}{(N+\alpha+\beta-2)}$$
 
@@
     $$\hat{\textbf{p}}^{MAP}  =  \frac{\alpha - 1 +\sum_{n=1}^{N} \textbf{x}^n}{(N+\alpha+\beta-2)}$$
-
-As the number of data points increases, the MAP estimate approaches $\frac{1}{N}\sum_{n=1}^{N} \textbf{x}^{(n)}$, the MLE.
+As the number of data points increases, $\hat{\textbf{p}}^{MAP}$ approaches $\frac{1}{N}\sum_{n=1}^{N} \textbf{x}^{(n)}$, $\hat{\textbf{p}}^{MLE}$.
 This makes sense because as our data set gets bigger, we are less reliant on our prior.
-Moreover, if a specific pixel in all of the images of our data set are white or all black, the MLE for that pixel will be binary.
-This may not be representative of our intuitions about image pixels, as there should be some non-zero probability of a pixel being black or white.
+However, if a specific pixel in all of the images of our data set are white or all black, the MLE for that pixel would either be 1 or 0.
+This may not be representative of our intuitions about images, as there should be some non-zero probability of a pixel being black or white.
 By introducing an appropriate prior we can ensure that the probability of that pixel will never be exactly zero or one.
 In our case, with a Beta(3,3) prior on each pixel, our parameter values are biased to be closer to 0.5 and to never be at the extremities 0 and 1.
-We can see this in Figure $\ref{fig:1e}$ where the range of our parameters is smaller than the range of Figure $\ref{fig:1d}$.
+We can see this in Figure $\ref{fig:1e}$ where the range of our parameters is smaller than the range of Figure $\ref{fig:1d}$ and doesn't include zero.
 Figure $\ref{fig:1e-comparison}$ visualises $\hat{\textbf{p}}^{MAP}-\hat{\textbf{p}}^{MLE}$ and we can see that for likelihoods greater than 0.5 in the MLE, the MAP has a lower value and for likelihoods less than 0.5, the MAP has a higher value, confirming our intuitions.
@@
 The logarithm of zero is negative infinity, so having if the MLE is zero it can be problematic for log-likelihoods calculations whereas MAP can ensure non-zero probabilities.
-We can see in Figure $\ref{fig:1d}$ and Figure $\ref{fig:1e}$ that the MLE has parameter values of zero whereas the MAP does not.
-Interestingly, when $\alpha=1$ and $\beta=1$, $\hat{\textbf{p}}^{MLE} = \hat{\textbf{p}}^{MAP}$.
-Intuitively this is when the prior is a uniform distribution and so there aren't any biases on the location of $\textbf{p}$ and we recover the MLE.
+Interestingly, when $\alpha=\beta=1$, $\hat{\textbf{p}}^{MLE} = \hat{\textbf{p}}^{MAP}$.
+This is when the prior is a uniform distribution and so there is uniform bias on the location of $\textbf{p}$ and we recover the MLE.
 
-On the other hand, a mis-specified prior can be problematic, as the estimated parameters might be skewed by the prior and not properly represent the underlying data generating process, this can result in parameter estimates that are worse than using the MLE.
+On the other hand, a mis-specified prior can be problematic, as the estimated parameters might be skewed by the prior and not properly represent the underlying data generating process, this can result in parameter estimates that are worse than using the MLE if our data set is limited.
 %In this example, we're using images of different digits as part of our data set.
@@
 
-\begin{enumerate}
-
-\item[(a)] When all D components are generated from a Bernoulli distribution with $p_d = 0.5$, we have the likelihood function for model $M_1$:
-
-$$P(\textbf{x}^{(n)|\textbf{p}^{(1)}}=[0.5, 0.5,...,0.5]^T, M_1) = \prod_{d=1}^{D} (0.5)^{x_d^{(n)}} (0.5)^{1-x_d^{(n)}}$$
-
-\item[(b)] When all D components are generated from Bernoulli distributions with unknown, but identical, $p_d$, we have the likelihood function for model $M_2$:
-
-$$P(\textbf{x}^{(n)}|\textbf{p}^{(2)}}=[p_d, p_d, ..., p_d]^T, M_2) = \prod_{d'=1}^{D} p_d^{x_{d'}^{(n)}} (1-p_d)^{1-x_{d'}^{(n)}}$$
-
-\item[(c)] When each component is Bernoulli distributed with separate, unknown $p_d$, we have the likelihood function for model $M_3$:
-
-$$P(\textbf{x}^{(n)}|\textbf{p}^{(3)}}=[p_1, p_2, ..., p_D]^T, M_3) = \prod_{d=1}^{D} p_d^{x_d^{(n)}} (1-p_d)^{1-x_d^{(n)}}$$
-
+When all D components are generated from a Bernoulli distribution with $p_d = 0.5$, we have the likelihood function for model $M_1$:
+$$P(\textbf{x}^{(n)|\textbf{p}^{(1)}}=[0.5, 0.5,...,0.5]^T, M_1) = \prod_{n=1}^{N}\prod_{d=1}^{D} (0.5)^{x_d^{(n)}} (0.5)^{1-x_d^{(n)}}$$
+When all D components are generated from Bernoulli distributions with unknown, but identical, $p_d$, we have the likelihood function for model $M_2$:
+$$P(\textbf{x}^{(n)}|\textbf{p}^{(2)}}=[p_d, p_d, ..., p_d]^T, M_2) = \prod_{n=1}^{N}\prod_{d'=1}^{D} p_d^{x_{d'}^{(n)}} (1-p_d)^{1-x_{d'}^{(n)}}$$
+When each component is Bernoulli distributed with separate, unknown $p_d$, we have the likelihood function for model $M_3$:
+$$P(\textbf{x}^{(n)}|\textbf{p}^{(3)}}=[p_1, p_2, ..., p_D]^T, M_3) = \prod_{n=1}^{N}\prod_{d=1}^{D} p_d^{x_d^{(n)}} (1-p_d)^{1-x_d^{(n)}}$$
 For each model $M_i$, we can marginalise out $\textbf{p}^{(i)}}$ to get $P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_i)$:
-
 $$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_i) = \int_0^1...\int_0^1 P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|p_d, M_i)P(p_d|M_i) dp_1...dp_D$$
-
 where $d=1,...,D$ and $ \{\textbf{x}^{(n)}\}_{n=1}^{N}$ is our data set.
@@
 Given that the prior of any unknown probabilities is uniform, i.e. $P(p_d | M_i) = 1$. We can simplify:
-
 $$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_i) = \int_0^1...\int_0^1 P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|p_d, M_i) dp_1...dp_D$$
-
 For $M_1$, we have that all pixels have probability 0.5:
-
-$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_1) = \int_0^1...\int_0^1 \prod_{j=n}^{N} \prod_{d=1}^{D} (0.5)^{x_d^{(n)}} (1-0.5)^{1-{x_d}^{(n)}} d\theta_1...d\theta_D$$
-
+$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_1) = \int_0^1...\int_0^1 \prod_{n=1}^{N} \prod_{d=1}^{D} (0.5)^{x_d^{(n)}} (1-0.5)^{1-{x_d}^{(n)}} d\theta_1...d\theta_D$$
 We can remove the integrals and knowing that either $x_d^{(n)}$ or $1-x_d^{(n)}$ will be 1 and the other zero, we can simplify $(0.5)^{x_d^{(n)}} (1-0.5)^{1-{x_d}^{(n)}}$ to $0.5$:
-
-$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_1) = \prod_{j=1}^{N} \prod_{d=1}^D (0.5)$$
+$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_1) = \prod_{n=1}^{N} \prod_{d=1}^D (0.5)$$
 $$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_1) = (0.5)^{N\cdot D}$$
-
 For $M_2$, we have that all pixels share some probability $p_d$ so we only need to integrate over a single variable $p_d$:
-
-$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_2) = \int_0^1 \prod_{n=1}^{N} \prod_{d'=1}^D p^{x_{d'}^{(n)}} (1-p_d)^{1-{x_{d'}}^{(n)}} dp_d$$
-
+$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_2) = \int_0^1 \prod_{n=1}^{N} \prod_{d'=1}^D p_d^{x_{d'}^{(n)}} (1-p_d)^{1-{x_{d'}}^{(n)}} dp_d$$
 Changing the products to sums:
-
-$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_2) = \int_0^1  p_d^{\sum_{n=1}^{N} \sum_{d'=1}^D x_{d'}^{(n)}} (1-p)^{\sum_{j=1}^{N} \sum_{d'=1}^D 1-{x_{d'}^{(n)}}} d p_d$$
-
+$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_2) = \int_0^1  p_d^{\sum_{n=1}^{N} \sum_{d'=1}^D x_{d'}^{(n)}} (1-p_d)^{\sum_{j=1}^{N} \sum_{d'=1}^D 1-{x_{d'}^{(n)}}} d p_d$$
 Rewriting:
-
-$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_2) = \int_0^1  p_d^{k} (1- p_{d'=1})^{N\cdot D-k} d p_d$$
-
-where $k =\sum_{n=1}^{N} \sum_{d'=1}^D x_{d'}^{(n)}$.
+$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_2) = \int_0^1  (p_d)^{K} (1- p_{d'=1})^{N\cdot D-K} d p_d$$
+where $K =\sum_{n=1}^{N} \sum_{d'=1}^D x_{d'}^{(n)}$.
 
 This integral is the beta function:
-
-$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_2) = \frac{k! (N\cdot D-k)!}{(N\cdot D+1)!}$$
-
+$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_2) = \frac{K! (N\cdot D-k)!}{(N\cdot D+1)!}$$
 For $M_3$, we need an integral for each $p_d$:
-
 $$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_3) = \int_0^1 ... \int_0^1 \prod_{n=1}^{N} \prod_{d=1}^D  p_d^{x_d^{(n)}} (1- p_d)^{1-{x_d}^{(n)}} d p_1 ... d p_D$$
-
 We can separate the integrals to only contain the relevant $p_d$:
-
 $$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_3) = \prod_{d=1}^D \Bigg( \int_0^1 \prod_{n=1}^{N}   p_d^{x_d^{(n)}} (1- p_d)^{1-{x_d}^{(n)}} d p_d \Bigg)$$
-
 Changing the products to sums:
-
 $$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_3) = \prod_{d=1}^D \Bigg( \int_0^1   p_d^{\sum_{n=1}^{N} x_d^{(n)}} (1- p_d)^{\sum_{n=1}^{N} 1-{x_d}^{(j)}} d p_d \Bigg)$$
-
 In this case, we have the product of integrals where each evaluates to a beta function:
-
-$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_3) = \prod_{d=1}^D \frac{k_d! (N-k_d)!}{(N+1)!}$$
-
-where $k_d = \sum_{n=1}^{N} x_d^{(in}$.
+$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_3) = \prod_{d=1}^D \frac{K_d! (N-K_d)!}{(N+1)!}$$
+where $K_d = \sum_{n=1}^{N} x_d^{(n}$.
 
 The posterior probability of a model $M_i$ can be expressed:
+$$P(M_i |  \{\textbf{x}^{(n)}\}_{n=1}^{N}) = \frac{P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_i)P(M_i)}{P( \{\textbf{x}^{(n)}\}_{n=1}^{N})}$$
+We only have three models, so in this case the normalisation $P(\{\textbf{x}^{(n)}\}_{n=1}^{N})$ can be expressed as a sum:
+$$P(M_i| \{\textbf{x}^{(n)}\}_{n=1}^{N}) = \frac{P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_i)P(M_i)}{\sum_{i\in \{1,2,3\}} P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_i)P(M_i)}$$
+Given that $P(M_i) = \frac{1}{3}$ for all $i \in \{1,2,3\}$:
+$$P(M_i| \{\textbf{x}^{(n)}\}_{n=1}^{N}) = \frac{P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_i)}{\sum_{i\in \{1,2,3\}} P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_i)}$$
+Calculating the posterior probabilities of each of the three models having generated the data in binarydigits.txt using python, we can show the values in the Table $\ref{table2}$:
 
-$$P(M_i |  \{\textbf{x}^{(n)}\}_{n=1}^{N}) = \frac{P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_i)P(M_i)}{P( \{\textbf{x}^{(n)}\}_{n=1}^{N})}$$
+\begin{table}[h]
+\begin{center}
 
-We only have three models, so in this case the normalisation $P(\{\textbf{x}^{(n)}\}_{n=1}^{N})$ can be expressed as a sum:
+\begin{tabular}{l|c}%
 
-$$P(M_i| \{\textbf{x}^{(n)}\}_{n=1}^{N}) = \frac{P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_i)P(M_i)}{\sum_{i\in \{1,2,3\}} P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_i)P(M_i)}$$
-
-Given that $P(M_i) = \frac{1}{3}$ for all $i \in \{1,2,3\}$:
-
-$$P(M_i| \{\textbf{x}^{(n)}\}_{n=1}^{N}) = \frac{P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_i)}{\sum_{i\in \{1,2,3\}} P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_i)}$$
-
-Calculating the posterior probabilities of each of the three models having generated the data in binarydigits.txt using python we can show the values in the table below:
-
-\begin{center}
-\begin{tabular}{l|c}%
  \bfseries $i$ & \bfseries$ P(M_i|\{\textbf{x}^{(n)}\}_{n=1}^{N})$% specify table head
@@
 \end{center}
+\caption{Posterior Probabilities}
+\label{table2}
+\end{table}
 
-We can see that for models specified to have the same parameter value for all pixels is very unlikely under the given data set.
-This makes sense because it is specifying models where the image is essentially blank, which is not reflective of the data.
+We can see that for models specified to have the same parameter value for all pixels like $M_1$ is very unlikely with the given data set.
+This makes sense because it is specifying models where the image is essentially blank (a uniform shade), which is not reflective of our digit images.
 Moreover, $M_1$ specifies a specific value of 0.5 for all the parameters whereas $M_2$ specifies any value for all the parameters as long as it's the same.
@@
 \lstinputlisting[language=Python]{src/solutions/q2.py}
-\end{enumerate}
 \newpage
@@
 , \pi_K] \in \mathbb{R}^{K \times 1}$ our mixing proportions $(0 \leq \pi_k \leq 1; \sum_k \pi_k = 1)$ and $\mathbf{P} \in \mathbb{R}^{D \times K}$ the K Bernoulli parameter vectors with elements $p_{kd}$ denoting the probability that pixel d takes value 1 under mixture component k.
-We are also the images are iid under the model, and that the pixels are independent of each other within each component distribution.
+We also assume the images are iid and that the pixels are independent of each other within each component distribution.
 
@@
 
-\item[(b)] The expression for the responsibility of mixture component k for data vector
-$x^{(n)}$, i.e. $r_{nk} \equiv P (s^{(n)} = k|x^{(n)}, \pi, \mathbf{P})$. This computation provides the E-step for an EM algorithm.
-
-We know that:
+\item[(b)] We know that:
 $$P(A|B) \propto P(B|A)P(A)$$
 Thus,
-$$P(s^{(n)} = k|x^{(n)}, \pi, \mathbf{P}) \propto  P(x^{(n)} | s^{(n)} = k, \pi, \mathbf{P}) P(s^{(n)} = k|\pi, \mathbf{P})$$
-where $s^{(n)} \in \{1,...,K\}$ a discrete latent variable where $P(s^{(n)} = k|x^{(n)} | \pi) = \pi_k$.
-Note that $P(s^{(n)} = k|x^{(n)} | \pi) = P(s^{(n)} = k|x^{(n)} | \pi, \mathbf{P})$ as $s^{(n)$ isn't dependent on $\mathbf{P}$.
+$$P(s^{(n)} = k|\textbf{x}^{(n)}, \pi, \mathbf{P}) \propto  P(\textbf{x}^{(n)} | s^{(n)} = k, \pi, \mathbf{P}) P(s^{(n)} = k|\pi, \mathbf{P})$$
+where $s^{(n)} \in \{1,...,K\}$ a discrete hidden variable with $P(s^{(n)} = k|\textbf{x}^{(n)}, \pi) = \pi_k$.
+Note that $P(s^{(n)} = k|\textbf{x}^{(n)} , \pi) = P(s^{(n)} = k|\textbf{x}^{(n)} , \pi, \mathbf{P})$ as $s^{(n)$ isn't dependent on $\mathbf{P}$.
 
-Let $P(s^{(n)} = k|x^{(n)}, \pi, \mathbf{P}) \propto P(s^{(n)}$ be the unnormalised responsibility $\tilde{r}_{nk}$. Using the mixture for component $k$ and the likelihood function of component $k$:
-$$\tilde{r}_{nk} = \pi_k \prod_{d=1}^D (p_{kd})^{x^{(n)}_{d}} (1-p_{kd})^{1-x^{(n)}_{d}}$$
+Let $P(s^{(n)} = k|\textbf{x}^{(n)}, \pi, \mathbf{P}) \propto P(s^{(n)}$ be the unnormalised responsibility $\tilde{r}_{nk}$.
+Using the mixture for component $k$, $\pi_k$ and the likelihood function of component $k$:
+$$\tilde{r}_{nk} = \pi_k \prod_{d=1}^D (p_{kd})^{\textbf{x}^{(n)}_{d}} (1-p_{kd})^{1-\textbf{x}^{(n)}_{d}}$$
 Normalising across the components:
 $$r_{nk} = \frac{\tilde{r}_{nk}}{\sum_{j=1}^K\tilde{r}_{nj}}$$
-and $r_{nk}$, we have calculated $P(s^{(n)} = k|x^{(n)}, \pi, \mathbf{P})$ for the E step.
+we have calculated $P(s^{(n)} = k|\textbf{x}^{(n)}, \pi, \mathbf{P})$ for the E step of an EM algorithm.
 
 Moreover,
-$$ \log \tilde{r}_{nk} = \log \pi_k + \sum_{d=1}^D \bigg( x^{(n)}_{d}\log(p_{kd}) + (1-x^{(n)}_{d}) \log(1-\exp(\log(p_{kd}))) \bigg)$$
+$$ \log \tilde{r}_{nk} = \log \pi_k + \sum_{d=1}^D \bigg( \textbf{x}^{(n)}_{d}\log(p_{kd}) + (1-\textbf{x}^{(n)}_{d}) \log(1-\exp(\log(p_{kd}))) \bigg)$$
 and
 $$\log r_{nk} = \log\tilde{r}_{nk} - \log\sum_{j=1}^K\exp(\log\tilde{r}_{nj})$$
-which can be vectorised as $\log \textbf{r}_{n}$ using Python scipy's $logsumexp$ operation.
+which can be vectorised as $\log \textbf{r}_{n}$ calculated with $\log\pi$ and $\log \textbf{P}$ using Python scipy's $logsumexp$ operation.
 
 \item[(c)] We know that the expectation log joint can be expressed:
-$$\Bigg\langle \sum_n \log P(x^{(n)}, s^{(n)}| \pi, \mathbf{P})\Bigg\rangle_{q(\{s^{(n)}\})} = \sum_{n=1}^N q(s^{(n)}) \log P(x^{(n)}, s^{(n)}| \pi, \mathbf{P})$$
+$$\Bigg\langle \sum_n \log P(\textbf{x}^{(n)}, s^{(n)}| \pi, \mathbf{P})\Bigg\rangle_{q(\{s^{(n)}\})} = \sum_{n=1}^N q(s^{(n)}) \log P(\textbf{x}^{(n)}, s^{(n)}| \pi, \mathbf{P})$$
 Let this quantity be $E$. Each term of $E$ can be expressed:
@@
 and
-$$\log P(x^{(n)}, s^{(n)}| \pi, \mathbf{P}) = \log[P(x^{(n)}| s^{(n)}, \pi, \mathbf{P})P(s^{(n)}| \pi, \mathbf{P})]$$
+$$\log P(\textbf{x}^{(n)}, s^{(n)}| \pi, \mathbf{P}) = \log[P(\textbf{x}^{(n)}| s^{(n)}, \pi, \mathbf{P})P(s^{(n)}| \pi, \mathbf{P})]$$
 which is the vectorised version of $\log\tilde{r}_{nk}$ from part (b) so:
-$$\log P(x^{(n)}, s^{(n)}| \pi, \mathbf{P}) =\log(\pi)+\log(\mathbf{P})^T x^{(n)} + \log(1-\mathbf{P})^T(1-x^{(n)})$$
+$$\log P(\textbf{x}^{(n)}, s^{(n)}| \pi, \mathbf{P}) =\log(\pi)+\log(\mathbf{P})^T \textbf{x}^{(n)} + \log(1-\mathbf{P})^T(1-\textbf{x}^{(n)})$$
 Combining:
-$$ E = \sum_n \textbf{r}_{n}^T[\log(\pi)+\log(\mathbf{P})^T x^{(n)} + \log(1-\mathbf{P})^T(1-x^{(n)})]$$
-
+$$ E = \sum_n \textbf{r}_{n}^T[\log(\pi)+\log(\mathbf{P})^T \textbf{x}^{(n)} + \log(1-\mathbf{P})^T(1-\textbf{x}^{(n)})]$$
 To maximise with respect to $\pi$ and $\mathbf{P}$ for the M step, we want to take the derivative, set to zero, and solve for $\hat\pi$ and $\hat P$.
@@
 $$\frac{\partial E}{\partial \pi_k} = \sum_n r_{nk} \frac{1}{\pi_k}$$
-
 The second derivative:
-
 $$\frac{\partial E}{(\partial \pi_k)^2} = \sum_n r_{nk} \frac{-1}{(\pi_k)^2}$$
-
 is always negative because $r_{nk}\geq 0$, $\sum_{n}r_{nk}=1$, \pi_{k}\geq 0$, and $\sum_{n}\pi_{k}=1$, ensuring a maximum in the next step.
 
-We can calculate the maximiser:
+We can calculate the maximiser with:
 $$\frac{\partial E}{\partial \pi_k} + \lambda = 0$$
@@
 $$\hat \pi_k = \frac{\sum_{n}r_{nk}}{N}$$
-
 For the $dk^{th}$ element of $\mathbf{P}$:
-
-$$\frac{\partial E}{\partial \mathbf{P}_{dk}} = \sum_n r_{nk}\frac{\partial}{\partial \mathbf{P}_{dk}}[x^{(n)}_{d}\log \mathbf{P}_{dk}+(1-x^{(n)}_{d}) \log(1-\mathbf{P}_{dk})]$$
-
+$$\frac{\partial E}{\partial \mathbf{P}_{dk}} = \sum_n r_{nk}\frac{\partial}{\partial \mathbf{P}_{dk}}[\textbf{x}^{(n)}_{d}\log \mathbf{P}_{dk}+(1-\textbf{x}^{(n)}_{d}) \log(1-\mathbf{P}_{dk})]$$
 Simplifying:
-
-$$\frac{\partial E}{\partial \mathbf{P}_{dk}} = \sum_n r_{nk}(\frac{x^{(n)}_{d}}{ \mathbf{P}_{dk}}-\frac{1-x^{(n)}_{d}}{1-\mathbf{P}_{dk}})$$
-
-Similar to Question 1, we can see that taking the derivative, the term in the brackets will always be less than zero and with $r_{nk}\geq 0$ and $\sum_{n}r_{nk}=1$, the second derivative will always be negative.
+$$\frac{\partial E}{\partial \mathbf{P}_{dk}} = \sum_n r_{nk}(\frac{\textbf{x}^{(n)}_{d}}{ \mathbf{P}_{dk}}-\frac{1-\textbf{x}^{(n)}_{d}}{1-\mathbf{P}_{dk}})$$
+Similar to Question 1, we can see that taking second derivative, the term in the brackets will always be less than zero and with $r_{nk}\geq 0$ and $\sum_{n}r_{nk}=1$, the second derivative will always be negative.
 This ensures that we have a maximum in the next step.
@@
 Setting the derivative to zero:
-
-$$\frac{\sum_n  x^{(n)}_{d}r_{nk}}{ \mathbf{P}_{dk}}-\frac{\sum_n r_{nk}-\sum_n  x^{(n)}_{d}r_{nk}}{1-\mathbf{P}_{dk}} = 0$$
-
+$$\frac{\sum_n  \textbf{x}^{(n)}_{d}r_{nk}}{ \mathbf{P}_{dk}}-\frac{\sum_n r_{nk}-\sum_n  \textbf{x}^{(n)}_{d}r_{nk}}{1-\mathbf{P}_{dk}} = 0$$
 Solving for $\hat{\mathbf{P}}_{dk}$:
-
-$$\hat{\mathbf{P}}_{dk}\sum_n r_{nk}-\hat{\mathbf{P}}_{dk}\sum_n x^{(n)}_{d}r_{nk} = \sum_n  x^{(n)}_{d}r_{nk}-\hat{\mathbf{P}}_{dk}\sum_n  x^{(n)}_{d}r_{nk}$$
-
+$$\hat{\mathbf{P}}_{dk}\sum_n r_{nk}-\hat{\mathbf{P}}_{dk}\sum_n \textbf{x}^{(n)}_{d}r_{nk} = \sum_n  \textbf{x}^{(n)}_{d}r_{nk}-\hat{\mathbf{P}}_{dk}\sum_n  \textbf{x}^{(n)}_{d}r_{nk}$$
 Thus,
-
-$$\hat{\mathbf{P}}_{dk} = \frac{\sum_n x^{(n)}_{d}r_{nk}}{\sum_n r_{nk}}$$
-
+$$\hat{\mathbf{P}}_{dk} = \frac{\sum_n \textbf{x}^{(n)}_{d}r_{nk}}{\sum_n r_{nk}}$$
 We have the maximizing parameters for the expected log-joint
-
-$$\arg \max_{\pi, \mathbf{P}} \Bigg\langle \sum_n \log P(x^{(n)}, s^{(n)}| \pi, \mathbf{P})\Bigg\rangle_{q(\{s^{(n)}\})}$$
-
+$$\arg \max_{\pi, \mathbf{P}} \Bigg\langle \sum_n \log P(\textbf{x}^{(n)}, s^{(n)}| \pi, \mathbf{P})\Bigg\rangle_{q(\{s^{(n)}\})}$$
 thus obtaining an iterative update for the parameters $\pi$ and $\mathbf{P}$ in the M-step of EM.
-
-For numerical stability, we can compute the maximisation step for the MAP of $\mathbf{P}$,$\hat{\mathbf{P}}_{dk}^{MAP}$:
+For numerical stability, we can compute the maximisation step for the MAP of $\mathbf{P}$,$\hat{\mathbf{P}}_{dk}^{MAP}$ by solving:
 $$\frac{\partial E'}{\partial \mathbf{P}_{dk}} = 0$$
 where
-$$E' = \sum_{n=1}^N q(s^{(n)}) \log P(\mathbf{P} | \pi, x^{(n)}, s^{(n)})$$
-and
-$$\log P(\mathbf{P} | \pi, x^{(n)}, s^{(n)}) = \log P( x^{(n)}, s^{(n)} | \pi, \mathbf{P} ) + \log P(\mathbf{P}) - \log P(x^{(n)}, s^{(n)} | \pi)$$
-Assuming the same independent Beta prior on each pixel of each component:
+$$E' = \sum_{n=1}^N q(s^{(n)}) \log P(\mathbf{P} | \pi, \textbf{x}^{(n)}, s^{(n)})$$
+and from Bayes':
+$$\log P(\mathbf{P} | \pi, \textbf{x}^{(n)}, s^{(n)}) = \log P( \textbf{x}^{(n)}, s^{(n)} | \pi, \mathbf{P} ) + \log P(\mathbf{P}) - \log P(\textbf{x}^{(n)}, s^{(n)} | \pi)$$
+Assuming an independent Beta prior on each pixel of each component:
 $$\log P(\mathbf{P}) = \sum_{k=1}^K \sum_{d=1}^D -\log (B(\alpha, \beta)) + (\alpha-1)\log \mathbf{P}_{dk} + (\beta-1)\log(1-\mathbf{P}_{dk})$$
 and
-$$\frac{\partial \log P(\mathbf{P})}{\partial \mathbf{P}_{dk}} = \frac{(\alpha-1)}{\mathbf{P}_{dk}} -  \frac{(\beta-1)}{1-\mathbf{P}_{dk}}$$`
+$$\frac{\partial \log P(\mathbf{P})}{\partial \mathbf{P}_{dk}} = \frac{(\alpha-1)}{\mathbf{P}_{dk}} -  \frac{(\beta-1)}{1-\mathbf{P}_{dk}}$$
 Thus, the derivative can be expressed as:
-$$\frac{\partial E'}{\partial \mathbf{P}_{dk}} =  \sum_n \Bigg(r_{nk} \bigg(\frac{\partial \log P( x^{(n)}, s^{(n)} | \pi, \mathbf{P} )}{\partial \mathbf{P}_{dk}} + \frac{\partial \log P(\mathbf{P})}{\partial \mathbf{P}_{dk}} \bigg) \Bigg)$$
+$$\frac{\partial E'}{\partial \mathbf{P}_{dk}} =  \sum_n \Bigg(r_{nk} \bigg(\frac{\partial \log P( \textbf{x}^{(n)}, s^{(n)} | \pi, \mathbf{P} )}{\partial \mathbf{P}_{dk}} + \frac{\partial \log P(\mathbf{P})}{\partial \mathbf{P}_{dk}} \bigg) \Bigg)$$
 Substituting the appropriate expressions:
-$$\frac{\partial E'}{\partial \mathbf{P}_{dk}} =  \sum_n \Bigg(r_{nk} \bigg(\frac{x^{(n)}_{d}}{ \mathbf{P}_{dk}}-\frac{1-x^{(n)}_{d}}{1-\mathbf{P}_{dk}} + \frac{(\alpha-1)}{\mathbf{P}_{dk}} -  \frac{(\beta-1)}{1-\mathbf{P}_{dk}} \bigg) \Bigg)$$
+$$\frac{\partial E'}{\partial \mathbf{P}_{dk}} =  \sum_n \Bigg(r_{nk} \bigg(\frac{\textbf{x}^{(n)}_{d}}{ \mathbf{P}_{dk}}-\frac{1-\textbf{x}^{(n)}_{d}}{1-\mathbf{P}_{dk}} + \frac{(\alpha-1)}{\mathbf{P}_{dk}} -  \frac{(\beta-1)}{1-\mathbf{P}_{dk}} \bigg) \Bigg)$$
 Simplifying:
-$$\frac{\partial E'}{\partial \mathbf{P}_{dk}} =  \frac{(\alpha-1)+\sum_n r_{nk} x^{(n)}_{d}}{ \mathbf{P}_{dk}}-\frac{(\beta-1)+\sum_n r_{nk} (1-x^{(n)}_{d})}{1-\mathbf{P}_{dk}}$$
-
-This form is very similar to Question 2 (c).
-By a similar argument, we can see that the second derivative will always be negative (additionally $r_{nk}\geq 0$ and $\sum_n r_{nk}=1$ ), ensuring a maximum in the calculation of the next step.
-
+$$\frac{\partial E'}{\partial \mathbf{P}_{dk}} =  \frac{\sum_n r_{nk}(\alpha-1 + \textbf{x}^{(n)}_{d})}{ \mathbf{P}_{dk}}-\frac{\sum_n r_{nk}(\beta-\textbf{x}^{(n)}_{d})}{1-\mathbf{P}_{dk}}$$
+For a maximum, we see that we need $\alpha > \textbf{x}^{(n)}_{d}-1$  or $\beta < \textbf{x}^{(n)}_{d}$, both of which are satisfied knowing that $\alpha > 0$ and $\beta > 0$.
 Setting $\frac{\partial E'}{\partial \mathbf{P}_{dk}}=0$ we can calculate $\hat{\mathbf{P}}_{dk}^{MAP}$:
-
-$$\hat{\mathbf{P}}_{dk}^{MAP} = \frac{\alpha - 1 +\sum_n r_{nk} x^{(n)}_{d}}{(N+\alpha+\beta-2)}$$
-
+$$\sum_n r_{nk}(\alpha-1 + \textbf{x}^{(n)}_{d})} - \hat{\mathbf{P}}_{dk}\sum_n r_{nk}(\alpha-1 + \textbf{x}^{(n)}_{d})} = \hat{\mathbf{P}}_{dk}\sum_n r_{nk}(\beta-\textbf{x}^{(n)}_{d})$$
+$$ \hat{\mathbf{P}}_{dk}^{MAP} = \frac{\sum_n r_{nk}(\textbf{x}^{(n)}_{d}+\alpha-1)}{(\alpha+\beta-1)( \sum_n r_{nk})}$$
+As a sense check, we can see when setting $\alpha=1$ and $\beta=1$ we recover $ \hat{\mathbf{P}}_{dk}^{MLE}$ as we would expect.
 \newpage
@@
 \end{figure}
-
-%\begin{minipage}{0.5\textwidth}
-
-%\end{minipage}%
-%\begin{minipage}{0.5\textwidth}
-%  \centering
-%  \includegraphics[scale=0.3]{outputs/q3/q3e-0-optimised-p}
-%  \caption{EM optimised parameters}
-%  \label{fig:3d-optimised-p}
-%\end{minipage}
-%\end{figure}
 
@@
 
-%\begin{figure}[h]
-%\centering
-%\begin{minipage}{0.5\textwidth}
-%  \centering
-%  \includegraphics[scale=0.5]{outputs/q3/q3e-0-optimised-p}
-%  \caption{EM optimised parameters: Trial 0}
-%  \label{fig:3e-initialised-p-trial-0}
-%\end{minipage}%
-%\begin{minipage}{0.5\textwidth}
-%  \centering
-%  \includegraphics[scale=0.5]{outputs/q3/q3e-1-optimised-p}
-%  \caption{EM optimised parameters: Trial 1}
-%  \label{fig:3e-initialised-p-trial-1}
-%\end{minipage}
-%\end{figure}
-%\begin{figure}[h]
-%\centering
-%\begin{minipage}{0.5\textwidth}
-%  \centering
-%  \includegraphics[scale=0.5]{outputs/q3/q3e-2-optimised-p}
-%  \caption{EM optimised parameters: Trial 2}
-%  \label{fig:3e-initialised-p-trial-2}
-%\end{minipage}%
-%\begin{minipage}{0.5\textwidth}
-%  \centering
-%  \includegraphics[scale=0.5]{outputs/q3/q3e-3-optimised-p}
-%  \caption{EM optimised parameters: Trial 3}
-%  \label{fig:3e-initialised-p-trial-3}
-%\end{minipage}
-%\end{figure}
-
 For smaller k, we can visually see that we obtain very similar solutions (a 7 and a 0 for $k=2$).
 However for higher K, we see that this may not always be the case.
-For Trial 0 of $k=10$, we have two 0's whereas in Trial 1 we have three 0's.
-Interestingly, the zeros are all different, representing different variants of the written digit (i.e.
+For Trial 2 of $k=10$, we have three 5's whereas in Trial 4 we have two 5's.
+Interestingly, different clusters of the same digits can be different, representing different variants of the written digit (i.e.
 a slanted zero, a slightly slanted zero, and a symmetric zero).
@@
 However for $k=7$ and especially $k=10$, we can see some components have very small or zero probability (i.e.
-$\pi_1$ and $\pi_2$ of trial 2).
-The corresponding parameter visualisations will essentially never be utilised because no probability is assigned to the component.
-This can be verified when we perform a TSNE visualisation of the responsibility vector for each of the images.
+$\pi_2$ of trial 2).
+It will be unlikely for those components to represent very distinct clusters (i.e.
+the parameters for $\pi_2$ and $\pi_9$ are very similar in trial 2) This can be verified when we perform a TSNE visualisation of the responsibility vector for each of the images (Note that for $k=2$, the responsibility vector is displayed).
 We can see that for large k, qualitatively the number of clusters no longer matches the k value, indicating that some clusters are redundant.
@@
 Additionally, adding a prior on the responsibility components can be helpful to ensure non-zero mixing components unlike the components visualised here.
-This can help promote more meaningful clusters as $k$ increases.
+This could help promote more meaningful clusters as $k$ increases.
 
 \end{enumerate}
+
+[BONUS] Express the log-likelihoods obtained in bits and relate these numbers to the length
+of the naive encoding of these binary data. How does your number compare to gzip (or
+another compression algorithm)? Why the difference? [5 marks]
+
+ [BONUS] Consider the total cost of encoding both the model parameters and the data given
+the model. How does this total cost compare to gzip (or similar)? How does it depend on
+K ? What might this tell you? [5 marks]
 
@@
 
-\item[(a)] Let $p(\textbf{s}_i,\textbf{s}_{i-1})$ be the probability of the pair of symbols $\textbf{s}_i$ and $\textbf{s}_{i-1}$ occurring together in the text ($\textbf{s}_{i-1}$ followed by $\textbf{s}_i$ where order matters).
-We can model $p(\textbf{s}_i,\textbf{s}_{i-1})$ as a multinomial distribution with $N=1$ and $D=53^2$:
+\item[(a)] The formulae for the ML estimates of $P(s_i = \alpha |s_{i-1} = \beta) = \Psi(\alpha, \beta)$:
+$$\Psi(\alpha, \beta) = \frac{N_{s_i, s_{i-1}}}{N_{s_{i-1}}}$$
+where $N_{s_i, s_{i-1}}$ is the count of the number of occurrences of the pair $(s_i, s_{i-1})$, where $s__{i-1}$ is followed by $s_{i}$ and $N_{s__{i-1}}$ is the number of occurrences of $s_{i-1}$.
 
-$$p(\textbf{s}_i,\textbf{s}_{i-1}) = \frac{1}{s^1! s^2! \cdots s^{53^2}} \prod_{j \in \{1,...53\}, k \in \{1,...53\}} p_{s_i, s_{i-1}}^{t^{s_i, s_{i-1}}}$$
+Moreover, the stationary distribution $\phi$ can be calculated using the power method:
+\begin{enumerate}
+  \item[(i)] Initialise any $\phi_0 \in \mathbb{R}^{53 \times 1}$
+  \item[(ii)] Repeat $\phi_{i+1} = \Psi \phi_{i}$
+  \item[(iii)] Terminate when $\phi_{i+1} - \phi_{i} < \epsilon$
+\end{enumerate}
 
-where $t^{s_i, s_{i-1}}$ is an indicator of transition $s_{i-1}$ to $s_i$.
-For convenience we will denote $t^{(\alpha, \beta)})$ as the transition from
-A multinomial distribution is appropriate because there can only be only one of 53 symbols chosen as $s$ (i.e. a 53 dimensional dice).
-Thus, $p(s=s_i)$ represents the probability of the symbol $s_i$ in the text.
+where $\Psi \in \mathbf{R}^{53\times53}$ containing the transition probabilities, $\Psi_{i, j} = P(\alpha_j | \alpha_i)$ where $\alpha_i$ is the $i^{th}$ symbol and $\alpha_j$ is the $j^{th}$ symbol, and $\epsilon$ is some small number indicating sufficient convergence of the distribution to be considered stationary.
+The function $\phi(\gamma)$ is simply the index of $\gamma$ in the vector $\phi$.
 
-We can convert p(s) into exponential family form:
-
-$$p(\textbf{s}) = \frac{1}{s^1! s^2! \cdots s^D} \exp{(\textbf{s}^T \log(\textbf{p}))}$$
-
-where $\textbf{p}$ is the vector of $p_i$'s.
-Thus the sufficient statistic is $T(\textbf{s}) = \textbf{s}^T$.
-
-The ML estimate
+%Let $p(\textbf{s}_i,\textbf{s}_{i-1})$ be the probability of the pair of symbols $\textbf{s}_i$ and $\textbf{s}_{i-1}$ occurring together in the text ($\textbf{s}_{i-1}$ followed by $\textbf{s}_i$ where order matters).
+%We can model $p(\textbf{s}_i,\textbf{s}_{i-1})$ as a multinomial distribution with $N=1$ and $D=53^2$:
+%
+%$$p(\textbf{s}_i,\textbf{s}_{i-1}) = \frac{1}{s^1! s^2! \cdots s^{53^2}} \prod_{j \in \{1,...53\}, k \in \{1,...53\}} p_{s_i, s_{i-1}}^{t^{s_i, s_{i-1}}}$$
+%
+%where $t^{s_i, s_{i-1}}$ is an indicator of transition $s_{i-1}$ to $s_i$.
+%For convenience we will denote $t^{(\alpha, \beta)})$ as the transition from
+%A multinomial distribution is appropriate because there can only be only one of 53 symbols chosen as $s$ (i.e. a 53 dimensional dice).
+%Thus, $p(s=s_i)$ represents the probability of the symbol $s_i$ in the text.
+%
+%We can convert p(s) into exponential family form:
+%
+%$$p(\textbf{s}) = \frac{1}{s^1! s^2! \cdots s^D} \exp{(\textbf{s}^T \log(\textbf{p}))}$$
+%
+%where $\textbf{p}$ is the vector of $p_i$'s.
+%Thus the sufficient statistic is $T(\textbf{s}) = \textbf{s}^T$.
+%
+%The ML estimate
 
@@
 
+The transition matrix $\Psi$:
+
+\scalebox{0.18}{
+    \centering
+    \begin{tabular}{|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|c|}
+    \hline
+        \textbf{} & \textbf{=} & \textbf{space} & \textbf{-} & \textbf{","} & \textbf{;} & \textbf{:} & \textbf{!} & \textbf{?} & \textbf{/} & \textbf{.} & \textbf{'} & \textbf{double quotes} & \textbf{(} & \textbf{)} & \textbf{[} & \textbf{]} & \textbf{*} & \textbf{0} & \textbf{1} & \textbf{2} & \textbf{3} & \textbf{4} & \textbf{5} & \textbf{6} & \textbf{7} & \textbf{8} & \textbf{9} & \textbf{a} & \textbf{b} & \textbf{c} & \textbf{d} & \textbf{e} & \textbf{f} & \textbf{g} & \textbf{h} & \textbf{i} & \textbf{j} & \textbf{k} & \textbf{l} & \textbf{m} & \textbf{n} & \textbf{o} & \textbf{p} & \textbf{q} & \textbf{r} & \textbf{s} & \textbf{t} & \textbf{u} & \textbf{v} & \textbf{w} & \textbf{x} & \textbf{y} & \textbf{z} \\ \hline
+        \textbf{=} & 1.8e-02 & 5.9e-06 & 5.3e-04 & 2.5e-05 & 8.4e-04 & 9.4e-04 & 2.5e-04 & 3.2e-04 & 1.6e-02 & 3.2e-05 & 1.7e-02 & 1.3e-02 & 1.4e-03 & 1.4e-03 & 1.8e-02 & 1.8e-02 & 2.9e-03 & 4.5e-03 & 2.3e-03 & 5.1e-03 & 8.9e-03 & 1.3e-02 & 9.8e-03 & 9.4e-03 & 1.1e-02 & 4.1e-03 & 1.2e-02 & 5.0e-06 & 2.9e-05 & 1.6e-05 & 8.5e-06 & 3.2e-06 & 1.8e-05 & 2.0e-05 & 6.0e-06 & 5.9e-06 & 3.9e-04 & 5.0e-05 & 1.1e-05 & 1.6e-05 & 5.5e-06 & 5.3e-06 & 2.2e-05 & 4.2e-04 & 6.8e-06 & 6.4e-06 & 4.5e-06 & 1.6e-05 & 4.0e-05 & 1.7e-05 & 2.3e-04 & 2.2e-05 & 5.5e-04 \\ \hline
+        \textbf{space} & 5.5e-02 & 4.2e-03 & 9.0e-03 & 2.5e-05 & 8.4e-04 & 9.4e-04 & 2.5e-04 & 3.2e-04 & 1.6e-02 & 6.5e-04 & 3.3e-02 & 1.5e-01 & 8.1e-01 & 1.4e-03 & 3.6e-02 & 1.8e-02 & 7.9e-01 & 1.3e-02 & 5.1e-01 & 1.2e-01 & 1.3e-01 & 1.3e-01 & 6.9e-02 & 1.7e-01 & 7.7e-02 & 3.7e-02 & 7.1e-02 & 3.1e-01 & 6.6e-01 & 3.1e-01 & 1.3e-01 & 3.3e-02 & 3.5e-01 & 1.7e-01 & 2.7e-01 & 1.6e-01 & 5.7e-01 & 1.8e-01 & 1.3e-01 & 2.9e-01 & 7.3e-02 & 1.7e-01 & 3.6e-01 & 5.4e-01 & 9.2e-02 & 2.4e-01 & 3.6e-01 & 7.9e-02 & 1.4e-01 & 6.3e-01 & 1.0e-01 & 1.3e-01 & 7.0e-02 \\ \hline
+        \textbf{-} & 1.8e-02 & 3.3e-05 & 3.2e-03 & 2.5e-05 & 8.4e-04 & 9.4e-04 & 2.5e-04 & 3.2e-04 & 1.6e-02 & 3.2e-05 & 1.7e-02 & 1.3e-02 & 1.4e-03 & 1.4e-03 & 1.8e-02 & 1.8e-02 & 2.9e-03 & 4.5e-03 & 4.5e-03 & 5.1e-03 & 8.9e-03 & 1.3e-02 & 9.8e-03 & 2.8e-02 & 1.1e-02 & 8.2e-03 & 1.2e-02 & 2.5e-04 & 3.4e-03 & 2.9e-03 & 1.3e-03 & 2.5e-04 & 2.4e-03 & 7.1e-04 & 4.4e-04 & 5.3e-04 & 7.7e-04 & 1.8e-03 & 1.3e-03 & 1.1e-03 & 3.7e-04 & 5.0e-04 & 1.2e-03 & 8.5e-04 & 3.7e-04 & 1.1e-03 & 6.0e-04 & 4.8e-04 & 2.8e-04 & 7.6e-04 & 2.3e-04 & 7.3e-04 & 7.1e-03 \\ \hline
+        \textbf{","} & 1.8e-02 & 6.7e-02 & 5.3e-04 & 2.5e-05 & 8.4e-04 & 9.4e-04 & 2.5e-04 & 3.2e-04 & 1.6e-02 & 6.5e-05 & 1.7e-02 & 2.7e-02 & 4.2e-03 & 1.4e-03 & 1.8e-02 & 1.8e-02 & 5.9e-03 & 4.5e-02 & 2.3e-03 & 5.1e-03 & 8.9e-03 & 1.3e-02 & 9.8e-03 & 1.9e-02 & 1.1e-02 & 4.1e-03 & 1.2e-02 & 3.0e-03 & 5.8e-03 & 1.9e-03 & 6.5e-04 & 2.3e-04 & 1.5e-03 & 1.2e-03 & 8.6e-04 & 6.4e-04 & 4.6e-03 & 1.0e-03 & 6.3e-04 & 1.1e-03 & 3.4e-04 & 3.8e-04 & 2.4e-03 & 3.4e-03 & 6.2e-04 & 1.3e-03 & 1.1e-03 & 3.4e-04 & 5.5e-04 & 4.7e-03 & 2.3e-04 & 5.3e-04 & 1.1e-03 \\ \hline
+        \textbf{;} & 1.8e-02 & 2.1e-03 & 5.3e-04 & 2.5e-05 & 8.4e-04 & 9.4e-04 & 2.5e-04 & 3.2e-04 & 1.6e-02 & 3.2e-05 & 1.7e-02 & 1.3e-02 & 1.4e-03 & 1.4e-03 & 1.8e-02 & 1.8e-02 & 2.9e-03 & 4.5e-03 & 2.3e-03 & 5.1e-03 & 8.9e-03 & 1.3e-02 & 9.8e-03 & 9.4e-03 & 1.1e-02 & 4.1e-03 & 1.2e-02 & 6.0e-05 & 2.6e-04 & 3.3e-05 & 8.5e-06 & 9.6e-06 & 3.7e-05 & 3.9e-05 & 6.6e-05 & 5.3e-05 & 3.9e-04 & 5.0e-05 & 2.1e-05 & 3.3e-05 & 1.1e-05 & 2.6e-05 & 6.6e-05 & 4.2e-04 & 6.8e-06 & 5.1e-05 & 9.4e-05 & 1.6e-05 & 4.0e-05 & 6.9e-05 & 2.3e-04 & 6.6e-05 & 5.5e-04 \\ \hline
+        \textbf{:} & 1.8e-02 & 1.5e-03 & 5.3e-04 & 2.5e-05 & 8.4e-04 & 9.4e-04 & 2.5e-04 & 3.2e-04 & 1.6e-02 & 6.5e-05 & 1.7e-02 & 1.3e-02 & 4.2e-03 & 1.4e-03 & 1.8e-02 & 1.8e-02 & 8.8e-03 & 4.5e-03 & 6.8e-03 & 1.0e-02 & 8.9e-03 & 1.3e-02 & 9.8e-03 & 9.4e-03 & 1.1e-02 & 4.1e-03 & 1.2e-02 & 4.0e-05 & 5.9e-05 & 3.3e-05 & 3.4e-05 & 9.6e-06 & 3.7e-05 & 2.0e-05 & 5.4e-05 & 4.1e-05 & 3.9e-04 & 5.0e-05 & 2.1e-05 & 6.5e-05 & 5.5e-06 & 1.6e-05 & 1.1e-04 & 4.2e-04 & 6.8e-06 & 5.1e-05 & 5.8e-05 & 1.6e-05 & 4.0e-05 & 6.9e-05 & 2.3e-04 & 2.2e-05 & 5.5e-04 \\ \hline
+        \textbf{!} & 1.8e-02 & 3.0e-03 & 5.3e-04 & 2.5e-05 & 8.4e-04 & 9.4e-04 & 2.5e-04 & 3.2e-04 & 1.6e-02 & 5.7e-03 & 1.7e-02 & 1.3e-02 & 1.4e-03 & 5.7e-03 & 1.8e-02 & 1.8e-02 & 5.9e-03 & 4.5e-03 & 2.3e-03 & 5.1e-03 & 8.9e-03 & 1.3e-02 & 9.8e-03 & 9.4e-03 & 1.1e-02 & 4.1e-03 & 1.2e-02 & 6.5e-05 & 1.2e-04 & 3.3e-05 & 2.5e-05 & 1.3e-05 & 3.7e-05 & 5.9e-05 & 4.2e-05 & 5.3e-05 & 7.7e-04 & 2.5e-04 & 4.2e-05 & 9.8e-05 & 2.8e-05 & 1.6e-05 & 4.4e-05 & 4.2e-04 & 6.8e-06 & 2.6e-05 & 6.3e-05 & 1.6e-05 & 1.2e-04 & 2.3e-04 & 2.3e-04 & 1.3e-04 & 5.5e-04 \\ \hline
+        \textbf{?} & 1.8e-02 & 2.3e-03 & 5.3e-04 & 2.5e-05 & 8.4e-04 & 9.4e-04 & 2.5e-04 & 3.2e-04 & 1.6e-02 & 3.8e-03 & 1.7e-02 & 1.3e-02 & 4.2e-03 & 4.3e-03 & 1.8e-02 & 1.8e-02 & 2.9e-03 & 4.5e-03 & 2.3e-03 & 5.1e-03 & 8.9e-03 & 1.3e-02 & 9.8e-03 & 9.4e-03 & 1.1e-02 & 4.1e-03 & 1.2e-02 & 3.0e-05 & 1.8e-04 & 4.9e-05 & 3.4e-05 & 9.6e-06 & 5.5e-05 & 2.0e-05 & 8.4e-05 & 1.1e-04 & 3.9e-04 & 5.0e-05 & 2.1e-05 & 3.3e-05 & 4.4e-05 & 2.1e-05 & 4.4e-05 & 4.2e-04 & 6.8e-06 & 1.9e-05 & 7.6e-05 & 1.6e-05 & 4.0e-05 & 2.9e-04 & 2.3e-04 & 1.1e-04 & 5.5e-04 \\ \hline
+        \textbf{/} & 1.8e-02 & 2.0e-06 & 5.3e-04 & 2.5e-05 & 8.4e-04 & 9.4e-04 & 2.5e-04 & 3.2e-04 & 1.6e-02 & 3.2e-05 & 1.7e-02 & 1.3e-02 & 1.4e-03 & 1.4e-03 & 1.8e-02 & 1.8e-02 & 2.9e-03 & 4.5e-03 & 2.3e-03 & 5.1e-03 & 8.9e-03 & 2.6e-02 & 9.8e-03 & 9.4e-03 & 1.1e-02 & 4.1e-03 & 1.2e-02 & 5.0e-06 & 2.9e-05 & 4.9e-05 & 2.5e-05 & 3.2e-06 & 1.8e-05 & 2.0e-05 & 1.2e-05 & 5.9e-06 & 3.9e-04 & 5.0e-05 & 2.1e-05 & 1.6e-05 & 5.5e-06 & 5.3e-06 & 2.2e-05 & 4.2e-04 & 6.8e-06 & 6.4e-06 & 9.0e-06 & 1.6e-05 & 4.0e-05 & 1.7e-05 & 2.3e-04 & 4.4e-05 & 5.5e-04 \\ \hline
+        \textbf{.} & 1.8e-02 & 2.9e-02 & 5.3e-04 & 2.3e-04 & 8.4e-04 & 9.4e-04 & 3.0e-03 & 5.4e-03 & 1.6e-02 & 1.3e-01 & 1.7e-02 & 2.7e-02 & 2.4e-02 & 1.0e-01 & 1.8e-02 & 1.8e-02 & 1.5e-02 & 4.5e-03 & 5.9e-02 & 2.0e-02 & 7.1e-02 & 3.9e-02 & 2.9e-02 & 2.8e-02 & 5.5e-02 & 2.1e-02 & 4.8e-02 & 3.2e-03 & 9.4e-03 & 6.0e-03 & 1.1e-03 & 2.7e-04 & 2.3e-03 & 5.7e-04 & 2.3e-03 & 2.0e-03 & 1.4e-02 & 2.6e-03 & 4.3e-04 & 1.3e-03 & 1.5e-03 & 1.2e-03 & 8.8e-03 & 1.7e-03 & 6.4e-04 & 1.9e-03 & 5.2e-03 & 1.7e-04 & 2.4e-04 & 5.2e-03 & 2.3e-04 & 5.5e-04 & 3.8e-03 \\ \hline
+        \textbf{'} & 1.8e-02 & 2.0e-06 & 5.3e-04 & 5.0e-05 & 8.4e-04 & 9.4e-04 & 2.5e-04 & 3.2e-04 & 1.6e-02 & 3.2e-05 & 1.7e-02 & 1.3e-02 & 1.4e-03 & 1.4e-03 & 1.8e-02 & 1.8e-02 & 2.9e-03 & 4.5e-03 & 2.3e-03 & 5.1e-03 & 8.9e-03 & 1.3e-02 & 9.8e-03 & 9.4e-03 & 1.1e-02 & 4.1e-03 & 1.2e-02 & 9.9e-06 & 2.9e-05 & 1.6e-05 & 8.5e-06 & 3.2e-06 & 1.8e-05 & 2.0e-05 & 6.0e-06 & 5.9e-06 & 3.9e-04 & 5.0e-05 & 1.1e-05 & 1.6e-05 & 5.5e-06 & 5.3e-06 & 2.2e-05 & 4.2e-04 & 6.8e-06 & 3.9e-05 & 4.5e-06 & 1.6e-05 & 4.0e-05 & 1.7e-05 & 2.3e-04 & 2.2e-05 & 5.5e-04 \\ \hline
+        \textbf{double quotes} & 1.8e-02 & 2.0e-05 & 5.3e-04 & 2.5e-05 & 8.4e-04 & 9.4e-04 & 2.5e-04 & 3.2e-04 & 1.6e-02 & 3.2e-05 & 1.7e-02 & 1.3e-02 & 1.4e-03 & 2.9e-03 & 1.8e-02 & 1.8e-02 & 5.9e-03 & 4.5e-03 & 2.3e-03 & 5.1e-03 & 8.9e-03 & 1.3e-02 & 9.8e-03 & 9.4e-03 & 1.1e-02 & 4.1e-03 & 1.2e-02 & 5.0e-06 & 2.9e-05 & 1.6e-05 & 1.7e-05 & 3.2e-06 & 1.8e-05 & 2.0e-05 & 6.0e-06 & 1.2e-05 & 3.9e-04 & 5.0e-05 & 1.1e-05 & 1.6e-05 & 5.5e-06 & 5.3e-06 & 1.8e-04 & 4.2e-04 & 1.4e-05 & 6.4e-06 & 9.0e-06 & 1.6e-05 & 4.0e-05 & 1.7e-05 & 2.3e-04 & 2.2e-05 & 5.5e-04 \\ \hline
+        \textbf{(} & 1.8e-02 & 2.0e-06 & 5.3e-04 & 2.5e-05 & 8.4e-04 & 9.4e-04 & 2.5e-04 & 3.2e-04 & 1.6e-02 & 3.2e-05 & 1.7e-02 & 2.7e-02 & 1.4e-03 & 1.4e-03 & 1.8e-02 & 1.8e-02 & 2.9e-03 & 4.5e-03 & 4.1e-02 & 2.0e-01 & 1.4e-01 & 5.3e-02 & 2.0e-02 & 9.4e-03 & 1.1e-02 & 8.2e-03 & 1.2e-02 & 4.1e-04 & 5.0e-04 & 9.8e-05 & 6.8e-05 & 2.6e-05 & 2.7e-04 & 9.8e-05 & 4.6e-04 & 2.6e-04 & 7.7e-04 & 3.0e-04 & 7.4e-05 & 2.1e-04 & 9.9e-05 & 1.8e-04 & 5.3e-04 & 8.5e-04 & 3.4e-05 & 2.9e-04 & 4.5e-04 & 4.7e-05 & 1.6e-04 & 1.4e-03 & 2.3e-04 & 4.4e-05 & 5.5e-04 \\ \hline
+        \textbf{)} & 1.8e-02 & 7.8e-04 & 5.3e-04 & 4.3e-03 & 5.1e-03 & 1.9e-03 & 2.5e-04 & 3.2e-04 & 1.6e-02 & 1.2e-03 & 1.7e-02 & 1.3e-02 & 2.8e-03 & 1.4e-03 & 1.8e-02 & 1.8e-02 & 2.9e-03 & 4.5e-03 & 2.3e-03 & 5.1e-03 & 8.9e-03 & 1.3e-02 & 9.8e-03 & 9.4e-03 & 1.1e-02 & 4.1e-03 & 1.2e-02 & 4.0e-05 & 8.8e-05 & 1.6e-05 & 8.5e-06 & 3.2e-06 & 3.7e-05 & 2.0e-05 & 1.8e-05 & 2.3e-05 & 3.9e-04 & 5.0e-05 & 1.1e-05 & 3.3e-05 & 5.5e-06 & 2.1e-05 & 6.6e-05 & 4.2e-04 & 6.8e-06 & 1.3e-05 & 3.6e-05 & 1.6e-05 & 4.0e-05 & 6.9e-05 & 2.3e-04 & 2.2e-05 & 5.5e-04 \\ \hline
+        \textbf{[} & 1.8e-02 & 2.0e-06 & 5.3e-04 & 2.5e-05 & 8.4e-04 & 9.4e-04 & 2.5e-04 & 3.2e-04 & 1.6e-02 & 3.2e-05 & 1.7e-02 & 1.3e-02 & 1.4e-03 & 1.4e-03 & 1.8e-02 & 1.8e-02 & 2.9e-03 & 4.5e-03 & 2.3e-03 & 5.1e-03 & 8.9e-03 & 1.3e-02 & 9.8e-03 & 9.4e-03 & 1.1e-02 & 4.1e-03 & 1.2e-02 & 5.0e-06 & 2.9e-05 & 1.6e-05 & 8.5e-06 & 6.4e-06 & 1.8e-05 & 2.0e-05 & 6.0e-06 & 5.9e-06 & 3.9e-04 & 5.0e-05 & 1.1e-05 & 3.3e-05 & 5.5e-06 & 5.3e-06 & 2.2e-05 & 4.2e-04 & 6.8e-06 & 6.4e-06 & 4.5e-06 & 1.6e-05 & 4.0e-05 & 1.7e-05 & 2.3e-04 & 2.2e-05 & 5.5e-04 \\ \hline
+        \textbf{]} & 1.8e-02 & 2.0e-06 & 5.3e-04 & 2.5e-05 & 8.4e-04 & 9.4e-04 & 2.5e-04 & 3.2e-04 & 1.6e-02 & 3.2e-05 & 1.7e-02 & 1.3e-02 & 1.4e-03 & 1.4e-03 & 3.6e-02 & 1.8e-02 & 2.9e-03 & 4.5e-03 & 2.3e-03 & 5.1e-03 & 8.9e-03 & 1.3e-02 & 9.8e-03 & 9.4e-03 & 1.1e-02 & 4.1e-03 & 1.2e-02 & 5.0e-06 & 2.9e-05 & 1.6e-05 & 8.5e-06 & 3.2e-06 & 1.8e-05 & 2.0e-05 & 6.0e-06 & 5.9e-06 & 3.9e-04 & 5.0e-05 & 2.1e-05 & 1.6e-05 & 5.5e-06 & 5.3e-06 & 2.2e-05 & 4.2e-04 & 6.8e-06 & 6.4e-06 & 4.5e-06 & 1.6e-05 & 4.0e-05 & 1.7e-05 & 2.3e-04 & 2.2e-05 & 5.5e-04 \\ \hline
+        \textbf{*} & 1.8e-02 & 4.9e-04 & 5.3e-04 & 2.5e-05 & 8.4e-04 & 9.4e-04 & 2.5e-04 & 3.2e-04 & 1.6e-02 & 3.2e-05 & 1.7e-02 & 1.3e-02 & 2.4e-02 & 1.4e-03 & 1.8e-02 & 1.8e-02 & 2.6e-02 & 4.5e-03 & 2.3e-03 & 5.1e-03 & 8.9e-03 & 1.3e-02 & 9.8e-03 & 9.4e-03 & 1.1e-02 & 4.1e-03 & 1.2e-02 & 1.5e-05 & 2.9e-05 & 3.3e-05 & 1.7e-05 & 3.2e-06 & 1.8e-05 & 2.0e-05 & 6.0e-06 & 1.2e-05 & 3.9e-04 & 1.0e-04 & 1.1e-05 & 1.6e-05 & 5.5e-06 & 5.3e-06 & 4.4e-05 & 4.2e-04 & 6.8e-06 & 6.4e-06 & 4.5e-06 & 3.1e-05 & 4.0e-05 & 5.2e-05 & 2.3e-04 & 4.4e-05 & 5.5e-04 \\ \hline
+        \textbf{0} & 1.8e-02 & 8.2e-05 & 5.3e-04 & 3.0e-04 & 8.4e-04 & 9.4e-04 & 2.5e-04 & 3.2e-04 & 1.6e-02 & 6.5e-05 & 1.7e-02 & 1.3e-02 & 1.4e-03 & 2.9e-03 & 1.8e-02 & 3.6e-02 & 2.9e-03 & 1.3e-01 & 1.1e-02 & 2.0e-02 & 8.9e-03 & 1.3e-02 & 2.7e-01 & 1.2e-01 & 1.8e-01 & 2.5e-02 & 2.1e-01 & 5.0e-06 & 2.9e-05 & 4.9e-05 & 8.5e-06 & 3.2e-06 & 1.8e-05 & 2.0e-05 & 6.0e-06 & 5.9e-06 & 3.9e-04 & 5.0e-05 & 1.1e-05 & 1.6e-05 & 5.5e-06 & 5.3e-06 & 2.2e-05 & 4.2e-04 & 6.8e-06 & 6.4e-06 & 4.5e-06 & 1.6e-05 & 4.0e-05 & 3.5e-05 & 2.3e-04 & 2.2e-05 & 5.5e-04 \\ \hline
+        \textbf{1} & 1.8e-02 & 3.5e-05 & 5.3e-04 & 1.3e-04 & 8.4e-04 & 9.4e-04 & 2.5e-04 & 3.2e-04 & 1.6e-02 & 1.6e-03 & 1.7e-02 & 1.3e-02 & 2.8e-03 & 2.7e-02 & 1.8e-02 & 1.8e-02 & 2.9e-03 & 7.2e-02 & 3.8e-02 & 3.2e-01 & 1.9e-01 & 3.9e-02 & 9.8e-02 & 3.8e-02 & 9.9e-02 & 6.9e-01 & 1.2e-02 & 5.0e-06 & 2.9e-05 & 3.3e-05 & 8.5e-06 & 3.2e-06 & 1.8e-05 & 2.0e-05 & 6.0e-06 & 5.9e-06 & 3.9e-04 & 5.0e-05 & 1.1e-05 & 1.6e-05 & 5.5e-06 & 5.3e-06 & 2.2e-05 & 4.2e-04 & 6.8e-06 & 1.3e-05 & 4.5e-06 & 1.6e-05 & 4.0e-05 & 1.7e-05 & 2.3e-04 & 2.2e-05 & 5.5e-04 \\ \hline
+        \textbf{2} & 1.8e-02 & 6.7e-05 & 5.3e-04 & 3.5e-04 & 8.4e-04 & 9.4e-04 & 2.5e-04 & 6.3e-04 & 1.6e-02 & 4.5e-04 & 1.7e-02 & 1.3e-02 & 1.4e-03 & 5.7e-02 & 1.8e-02 & 3.6e-02 & 2.9e-03 & 5.8e-02 & 6.8e-03 & 3.1e-02 & 2.7e-02 & 5.3e-02 & 3.9e-02 & 2.8e-02 & 3.3e-02 & 4.1e-03 & 1.2e-02 & 9.9e-06 & 2.9e-05 & 1.3e-04 & 8.5e-06 & 3.2e-06 & 1.8e-05 & 2.0e-05 & 6.0e-06 & 5.9e-06 & 3.9e-04 & 1.0e-04 & 1.1e-05 & 1.6e-05 & 1.1e-05 & 5.3e-06 & 2.2e-05 & 4.2e-04 & 6.8e-06 & 6.4e-06 & 4.5e-06 & 1.6e-05 & 4.0e-05 & 1.7e-05 & 2.3e-04 & 2.2e-05 & 5.5e-04 \\ \hline
+        \textbf{3} & 1.8e-02 & 2.5e-05 & 5.3e-04 & 1.8e-04 & 8.4e-04 & 9.4e-04 & 2.5e-04 & 3.2e-04 & 1.6e-02 & 2.6e-04 & 1.7e-02 & 1.3e-02 & 1.4e-03 & 2.4e-02 & 1.8e-02 & 1.8e-02 & 2.9e-03 & 3.6e-02 & 4.5e-03 & 5.1e-03 & 8.9e-03 & 1.3e-02 & 9.8e-03 & 9.4e-03 & 1.1e-02 & 4.1e-03 & 1.2e-02 & 5.0e-06 & 2.9e-05 & 3.3e-05 & 8.5e-06 & 3.2e-06 & 1.8e-05 & 2.0e-05 & 6.0e-06 & 5.9e-06 & 3.9e-04 & 5.0e-05 & 1.1e-05 & 1.6e-05 & 5.5e-06 & 5.3e-06 & 2.2e-05 & 4.2e-04 & 4.1e-05 & 6.4e-06 & 1.8e-05 & 1.6e-05 & 4.0e-05 & 1.7e-05 & 2.3e-04 & 2.2e-05 & 5.5e-04 \\ \hline
+        \textbf{4} & 1.8e-02 & 9.8e-06 & 1.1e-03 & 7.5e-05 & 8.4e-04 & 9.4e-04 & 2.5e-04 & 3.2e-04 & 1.6e-02 & 2.3e-04 & 1.7e-02 & 1.3e-02 & 1.4e-03 & 5.7e-03 & 1.8e-02 & 1.8e-02 & 2.9e-03 & 1.3e-02 & 6.8e-03 & 5.1e-03 & 8.9e-03 & 1.3e-02 & 9.8e-03 & 9.4e-03 & 1.1e-02 & 4.1e-03 & 1.2e-02 & 5.0e-06 & 2.9e-05 & 1.6e-05 & 8.5e-06 & 3.2e-06 & 1.8e-05 & 2.0e-05 & 6.0e-06 & 5.9e-06 & 3.9e-04 & 5.0e-05 & 1.1e-05 & 1.6e-05 & 5.5e-06 & 5.3e-06 & 2.2e-05 & 4.2e-04 & 6.8e-06 & 6.4e-06 & 1.3e-05 & 1.6e-05 & 4.0e-05 & 1.7e-05 & 4.5e-04 & 2.2e-05 & 5.5e-04 \\ \hline
+        \textbf{5} & 1.8e-02 & 1.8e-05 & 1.1e-03 & 4.3e-04 & 1.7e-03 & 1.9e-03 & 2.5e-04 & 3.2e-04 & 3.2e-02 & 2.9e-04 & 1.7e-02 & 1.3e-02 & 1.4e-03 & 2.9e-03 & 1.8e-02 & 1.8e-02 & 2.9e-03 & 3.1e-02 & 2.3e-03 & 5.1e-03 & 8.9e-03 & 2.6e-02 & 9.8e-03 & 9.4e-03 & 1.1e-02 & 4.1e-03 & 2.4e-02 & 5.0e-06 & 2.9e-05 & 6.5e-05 & 8.5e-06 & 3.2e-06 & 1.8e-05 & 2.0e-05 & 6.0e-06 & 5.9e-06 & 3.9e-04 & 5.0e-05 & 1.1e-05 & 1.6e-05 & 5.5e-06 & 5.3e-06 & 2.2e-05 & 4.2e-04 & 6.8e-06 & 6.4e-06 & 9.0e-06 & 1.6e-05 & 4.0e-05 & 1.7e-05 & 2.3e-04 & 4.4e-05 & 5.5e-04 \\ \hline
+        \textbf{6} & 1.8e-02 & 2.2e-05 & 1.6e-03 & 2.3e-04 & 1.7e-03 & 9.4e-04 & 2.5e-04 & 6.3e-04 & 1.6e-02 & 1.9e-04 & 1.7e-02 & 1.3e-02 & 1.4e-03 & 1.4e-03 & 1.8e-02 & 1.8e-02 & 2.9e-03 & 2.7e-02 & 2.3e-03 & 1.5e-02 & 8.9e-03 & 2.6e-02 & 9.8e-03 & 1.2e-01 & 2.2e-02 & 4.1e-03 & 1.2e-02 & 5.0e-06 & 2.9e-05 & 3.3e-05 & 8.5e-06 & 3.2e-06 & 1.8e-05 & 2.0e-05 & 6.0e-06 & 5.9e-06 & 3.9e-04 & 5.0e-05 & 1.1e-05 & 1.6e-05 & 5.5e-06 & 5.3e-06 & 2.2e-05 & 4.2e-04 & 6.8e-06 & 6.4e-06 & 1.8e-05 & 1.6e-05 & 4.0e-05 & 1.7e-05 & 2.3e-04 & 2.2e-05 & 5.5e-04 \\ \hline
+        \textbf{7} & 1.8e-02 & 2.7e-05 & 5.3e-04 & 2.0e-04 & 8.4e-04 & 9.4e-04 & 5.0e-04 & 3.2e-04 & 1.6e-02 & 1.6e-04 & 1.7e-02 & 1.3e-02 & 1.4e-03 & 1.4e-03 & 1.8e-02 & 1.8e-02 & 2.9e-03 & 1.8e-02 & 4.5e-03 & 5.1e-03 & 8.9e-03 & 1.3e-02 & 9.8e-03 & 9.4e-03 & 1.1e-02 & 2.1e-02 & 2.4e-02 & 5.0e-06 & 2.9e-05 & 3.3e-05 & 8.5e-06 & 3.2e-06 & 1.8e-05 & 2.0e-05 & 6.0e-06 & 5.9e-06 & 3.9e-04 & 5.0e-05 & 1.1e-05 & 1.6e-05 & 5.5e-06 & 5.3e-06 & 2.2e-05 & 4.2e-04 & 6.8e-06 & 6.4e-06 & 1.8e-05 & 1.6e-05 & 4.0e-05 & 1.7e-05 & 2.3e-04 & 2.2e-05 & 5.5e-04 \\ \hline
+        \textbf{8} & 1.8e-02 & 1.6e-05 & 5.3e-04 & 7.5e-05 & 8.4e-04 & 9.4e-04 & 2.5e-04 & 3.2e-04 & 1.6e-02 & 1.6e-04 & 1.7e-02 & 1.3e-02 & 1.4e-03 & 1.4e-03 & 1.8e-02 & 1.8e-02 & 2.9e-03 & 3.6e-01 & 1.9e-01 & 2.6e-02 & 8.9e-03 & 2.6e-02 & 9.8e-03 & 1.9e-02 & 2.2e-02 & 8.2e-03 & 4.8e-02 & 5.0e-06 & 2.9e-05 & 1.6e-05 & 8.5e-06 & 3.2e-06 & 1.8e-05 & 2.0e-05 & 6.0e-06 & 5.9e-06 & 3.9e-04 & 5.0e-05 & 1.1e-05 & 1.6e-05 & 5.5e-06 & 1.1e-05 & 4.4e-05 & 4.2e-04 & 6.8e-06 & 6.4e-06 & 1.3e-05 & 1.6e-05 & 4.0e-05 & 1.7e-05 & 2.3e-04 & 2.2e-05 & 5.5e-04 \\ \hline
+        \textbf{9} & 1.8e-02 & 2.7e-05 & 5.3e-04 & 1.5e-04 & 8.4e-04 & 1.9e-03 & 2.5e-04 & 3.2e-04 & 1.6e-02 & 1.9e-04 & 1.7e-02 & 1.3e-02 & 1.4e-03 & 1.4e-03 & 1.8e-02 & 1.8e-02 & 2.9e-03 & 1.8e-02 & 2.3e-03 & 5.1e-03 & 8.9e-03 & 1.3e-02 & 9.8e-03 & 1.9e-02 & 2.2e-02 & 4.1e-03 & 1.2e-02 & 5.0e-06 & 5.9e-05 & 1.6e-05 & 8.5e-06 & 3.2e-06 & 1.8e-05 & 2.0e-05 & 6.0e-06 & 5.9e-06 & 3.9e-04 & 5.0e-05 & 1.1e-05 & 1.6e-05 & 5.5e-06 & 5.3e-06 & 2.2e-05 & 4.2e-04 & 6.8e-06 & 6.4e-06 & 9.0e-06 & 1.6e-05 & 4.0e-05 & 1.7e-05 & 2.3e-04 & 2.2e-05 & 5.5e-04 \\ \hline
+        \textbf{a} & 1.8e-02 & 2.4e-02 & 2.0e-02 & 2.4e-02 & 1.7e-02 & 6.6e-03 & 3.3e-02 & 1.7e-02 & 1.6e-02 & 1.6e-02 & 1.7e-02 & 1.3e-02 & 4.2e-03 & 1.4e-02 & 1.8e-02 & 1.8e-02 & 2.9e-03 & 4.5e-03 & 2.3e-03 & 5.1e-03 & 8.9e-03 & 1.3e-02 & 9.8e-03 & 9.4e-03 & 1.1e-02 & 4.1e-03 & 1.2e-02 & 1.3e-04 & 1.0e-01 & 1.2e-01 & 9.6e-02 & 2.4e-04 & 3.1e-02 & 6.7e-02 & 2.1e-03 & 5.1e-02 & 6.8e-02 & 1.3e-01 & 1.5e-01 & 7.7e-02 & 2.5e-01 & 2.5e-04 & 1.0e-01 & 3.8e-03 & 1.2e-01 & 1.3e-01 & 1.3e-01 & 3.9e-02 & 1.7e-01 & 3.8e-02 & 1.2e-02 & 1.2e-01 & 1.9e-01 \\ \hline
+        \textbf{b} & 1.8e-02 & 2.2e-04 & 3.2e-03 & 9.0e-04 & 8.4e-04 & 9.4e-04 & 5.0e-04 & 9.5e-04 & 1.6e-02 & 6.5e-04 & 1.7e-02 & 1.3e-02 & 1.4e-03 & 4.3e-03 & 1.8e-02 & 1.8e-02 & 2.9e-03 & 4.5e-03 & 2.3e-03 & 5.1e-03 & 8.9e-03 & 1.3e-02 & 9.8e-03 & 9.4e-03 & 1.1e-02 & 4.1e-03 & 1.2e-02 & 1.6e-02 & 6.8e-03 & 6.5e-05 & 1.6e-04 & 3.6e-02 & 1.8e-05 & 2.0e-05 & 1.8e-05 & 5.9e-03 & 6.8e-02 & 5.0e-05 & 3.8e-02 & 1.1e-03 & 5.0e-05 & 2.2e-02 & 2.2e-05 & 4.2e-04 & 1.3e-02 & 3.1e-03 & 1.2e-03 & 8.2e-02 & 1.9e-03 & 3.8e-04 & 2.3e-04 & 5.7e-02 & 5.5e-04 \\ \hline
+        \textbf{c} & 1.8e-02 & 9.2e-04 & 1.6e-03 & 9.8e-04 & 3.4e-03 & 1.9e-03 & 1.0e-03 & 6.3e-04 & 1.6e-02 & 1.0e-03 & 1.7e-02 & 1.3e-02 & 1.4e-03 & 5.7e-03 & 1.8e-02 & 1.8e-02 & 2.9e-03 & 4.5e-03 & 2.3e-03 & 5.1e-03 & 8.9e-03 & 1.3e-02 & 9.8e-03 & 9.4e-03 & 1.1e-02 & 4.1e-03 & 1.2e-02 & 3.3e-02 & 8.8e-05 & 1.7e-02 & 9.3e-05 & 4.2e-02 & 7.3e-05 & 9.8e-05 & 7.0e-02 & 1.5e-02 & 3.9e-04 & 1.3e-01 & 2.1e-02 & 9.8e-05 & 1.1e-05 & 6.8e-02 & 6.6e-05 & 5.0e-02 & 1.4e-02 & 6.8e-04 & 1.9e-02 & 2.6e-02 & 4.0e-05 & 1.7e-04 & 2.3e-04 & 6.9e-03 & 3.3e-03 \\ \hline
+        \textbf{d} & 1.8e-02 & 1.2e-01 & 1.2e-01 & 1.1e-01 & 1.0e-01 & 1.8e-01 & 8.0e-02 & 5.9e-02 & 1.6e-02 & 9.1e-02 & 1.7e-02 & 2.7e-02 & 7.1e-03 & 6.8e-02 & 1.8e-02 & 1.8e-02 & 2.9e-03 & 4.5e-03 & 2.3e-03 & 5.1e-03 & 8.9e-03 & 1.3e-02 & 9.8e-03 & 9.4e-03 & 1.1e-02 & 4.1e-03 & 1.2e-02 & 1.5e-02 & 9.1e-03 & 3.3e-03 & 1.3e-02 & 4.4e-02 & 4.6e-03 & 1.0e-02 & 2.7e-03 & 4.7e-02 & 1.0e-01 & 5.1e-03 & 1.4e-02 & 7.9e-03 & 2.9e-03 & 2.7e-02 & 5.0e-03 & 2.6e-02 & 2.5e-02 & 1.7e-02 & 3.7e-03 & 1.8e-02 & 1.4e-02 & 5.2e-03 & 2.3e-04 & 2.7e-02 & 3.3e-03 \\ \hline
+        \textbf{e} & 1.8e-02 & 1.8e-01 & 1.5e-01 & 1.5e-01 & 1.3e-01 & 1.5e-01 & 1.7e-01 & 1.8e-01 & 1.6e-02 & 1.4e-01 & 3.3e-02 & 1.3e-02 & 8.5e-03 & 1.1e-01 & 1.8e-02 & 1.8e-02 & 2.9e-03 & 4.5e-03 & 2.3e-03 & 1.0e-02 & 8.9e-03 & 1.3e-02 & 9.8e-03 & 9.4e-03 & 1.1e-02 & 4.1e-03 & 1.2e-02 & 7.0e-02 & 1.7e-02 & 9.6e-02 & 2.5e-01 & 2.6e-02 & 6.5e-02 & 4.5e-02 & 7.2e-03 & 2.2e-02 & 3.8e-02 & 1.6e-02 & 1.1e-01 & 1.2e-01 & 1.5e-01 & 1.1e-02 & 8.4e-02 & 1.3e-01 & 3.0e-01 & 1.3e-01 & 3.6e-02 & 6.2e-03 & 2.1e-01 & 6.5e-02 & 7.1e-01 & 8.3e-02 & 9.5e-02 \\ \hline
+        \textbf{f} & 1.8e-02 & 3.3e-02 & 9.0e-02 & 1.2e-02 & 8.4e-03 & 1.9e-02 & 1.3e-02 & 9.8e-03 & 1.6e-02 & 1.4e-02 & 1.7e-02 & 1.3e-02 & 2.8e-03 & 1.9e-02 & 1.8e-02 & 1.8e-02 & 2.9e-03 & 4.5e-03 & 1.1e-02 & 5.1e-03 & 8.9e-03 & 1.3e-02 & 9.8e-03 & 9.4e-03 & 1.1e-02 & 4.1e-03 & 1.2e-02 & 2.0e-02 & 2.2e-03 & 9.1e-04 & 2.4e-04 & 1.5e-02 & 5.2e-02 & 5.9e-04 & 9.1e-04 & 2.8e-02 & 2.3e-03 & 8.5e-04 & 1.3e-02 & 9.2e-04 & 1.3e-04 & 4.3e-02 & 1.1e-03 & 1.3e-03 & 3.9e-02 & 1.4e-03 & 1.1e-02 & 2.9e-02 & 5.9e-04 & 1.5e-03 & 2.3e-04 & 2.3e-03 & 5.5e-04 \\ \hline
+        \textbf{g} & 1.8e-02 & 3.5e-02 & 5.5e-02 & 3.7e-02 & 3.7e-02 & 5.8e-02 & 3.5e-02 & 4.4e-02 & 8.1e-02 & 3.4e-02 & 1.7e-02 & 8.0e-02 & 1.4e-03 & 3.1e-02 & 1.8e-02 & 1.8e-02 & 2.9e-03 & 4.5e-03 & 2.3e-03 & 5.1e-03 & 8.9e-03 & 1.3e-02 & 9.8e-03 & 9.4e-03 & 1.1e-02 & 4.1e-03 & 1.2e-02 & 1.8e-02 & 1.8e-03 & 6.7e-04 & 7.1e-04 & 1.9e-02 & 1.1e-03 & 9.1e-03 & 3.7e-02 & 1.5e-02 & 2.3e-03 & 1.5e-04 & 1.6e-02 & 9.6e-04 & 5.3e-03 & 1.7e-02 & 1.1e-03 & 8.5e-04 & 1.9e-02 & 6.1e-03 & 2.0e-03 & 2.1e-02 & 5.1e-04 & 1.8e-03 & 2.3e-04 & 2.2e-03 & 1.6e-03 \\ \hline
+        \textbf{h} & 1.8e-02 & 2.4e-02 & 2.0e-02 & 3.2e-02 & 2.6e-02 & 1.4e-02 & 5.0e-02 & 3.5e-02 & 1.6e-02 & 1.8e-02 & 1.7e-02 & 1.3e-02 & 5.6e-03 & 1.7e-02 & 1.8e-02 & 1.8e-02 & 2.9e-03 & 4.5e-03 & 4.5e-03 & 5.1e-03 & 8.9e-03 & 1.3e-02 & 9.8e-03 & 9.4e-03 & 1.1e-02 & 4.1e-03 & 1.2e-02 & 1.4e-01 & 2.6e-03 & 1.7e-03 & 8.5e-04 & 2.4e-01 & 1.7e-03 & 4.7e-04 & 7.3e-04 & 1.5e-01 & 7.7e-04 & 5.0e-03 & 1.7e-03 & 6.7e-03 & 7.8e-04 & 7.2e-02 & 1.1e-03 & 4.2e-04 & 8.9e-03 & 1.5e-03 & 2.0e-02 & 2.5e-02 & 5.1e-04 & 1.8e-03 & 2.3e-04 & 2.0e-02 & 5.5e-04 \\ \hline
+        \textbf{i} & 1.8e-02 & 8.9e-03 & 1.2e-02 & 5.6e-03 & 3.4e-03 & 1.9e-03 & 5.5e-03 & 1.4e-02 & 1.6e-02 & 4.2e-03 & 1.7e-02 & 4.0e-02 & 2.8e-03 & 1.4e-03 & 1.8e-02 & 1.8e-02 & 2.9e-03 & 4.5e-03 & 2.3e-03 & 5.1e-03 & 8.9e-03 & 1.3e-02 & 9.8e-03 & 9.4e-03 & 1.1e-02 & 4.1e-03 & 1.2e-02 & 1.5e-02 & 3.8e-02 & 1.4e-01 & 7.4e-02 & 2.7e-02 & 6.9e-02 & 8.4e-02 & 2.4e-04 & 2.9e-03 & 3.9e-04 & 6.0e-02 & 8.4e-02 & 1.6e-01 & 2.7e-01 & 3.8e-02 & 1.9e-02 & 1.2e-02 & 4.0e-02 & 1.4e-01 & 9.5e-02 & 2.1e-03 & 1.4e-01 & 8.7e-04 & 7.3e-02 & 4.4e-05 & 3.2e-01 \\ \hline
+        \textbf{j} & 1.8e-02 & 2.0e-06 & 5.3e-04 & 2.5e-05 & 8.4e-04 & 9.4e-04 & 2.5e-04 & 3.2e-04 & 1.6e-02 & 3.2e-05 & 1.7e-02 & 1.3e-02 & 1.4e-03 & 1.4e-03 & 1.8e-02 & 1.8e-02 & 2.9e-03 & 4.5e-03 & 2.3e-03 & 5.1e-03 & 8.9e-03 & 1.3e-02 & 9.8e-03 & 9.4e-03 & 1.1e-02 & 4.1e-03 & 1.2e-02 & 4.8e-04 & 2.9e-05 & 1.6e-05 & 8.5e-06 & 1.8e-03 & 1.8e-05 & 2.0e-05 & 6.0e-06 & 7.0e-05 & 3.9e-04 & 5.0e-05 & 1.1e-05 & 1.6e-05 & 5.5e-06 & 3.8e-03 & 2.2e-05 & 4.2e-04 & 6.8e-06 & 6.4e-06 & 4.5e-06 & 1.8e-02 & 4.0e-05 & 1.7e-05 & 2.3e-04 & 2.2e-05 & 5.5e-04 \\ \hline
+        \textbf{k} & 1.8e-02 & 7.3e-03 & 1.8e-02 & 1.4e-02 & 1.4e-02 & 1.0e-02 & 1.6e-02 & 1.7e-02 & 3.2e-02 & 1.3e-02 & 1.7e-02 & 1.3e-02 & 2.8e-03 & 1.0e-02 & 1.8e-02 & 1.8e-02 & 2.9e-03 & 4.5e-03 & 2.3e-03 & 5.1e-03 & 8.9e-03 & 1.3e-02 & 9.8e-03 & 9.4e-03 & 1.1e-02 & 4.1e-03 & 1.2e-02 & 2.7e-03 & 5.3e-04 & 9.3e-04 & 6.8e-05 & 1.8e-02 & 4.7e-04 & 2.7e-04 & 5.0e-03 & 2.0e-02 & 7.7e-04 & 5.0e-05 & 3.6e-03 & 6.0e-04 & 1.1e-02 & 1.5e-03 & 2.4e-04 & 4.2e-04 & 6.8e-04 & 4.9e-03 & 2.0e-04 & 1.0e-02 & 5.5e-04 & 1.7e-03 & 2.3e-04 & 2.3e-03 & 5.5e-04 \\ \hline
+        \textbf{l} & 1.8e-02 & 1.9e-02 & 6.0e-02 & 3.5e-02 & 2.9e-02 & 2.1e-02 & 3.6e-02 & 2.3e-02 & 1.6e-02 & 2.0e-02 & 1.7e-02 & 1.3e-02 & 2.8e-03 & 2.0e-02 & 1.8e-02 & 1.8e-02 & 2.9e-03 & 4.5e-03 & 2.3e-03 & 5.1e-03 & 8.9e-03 & 1.3e-02 & 9.8e-03 & 9.4e-03 & 1.1e-02 & 4.1e-03 & 1.2e-02 & 3.8e-02 & 3.0e-03 & 1.8e-03 & 5.9e-02 & 5.4e-02 & 4.3e-02 & 2.1e-03 & 3.2e-04 & 5.5e-02 & 3.9e-04 & 5.2e-02 & 1.4e-01 & 7.0e-03 & 7.5e-04 & 4.4e-02 & 9.3e-03 & 2.5e-03 & 3.4e-03 & 1.1e-02 & 7.9e-03 & 2.3e-02 & 1.7e-02 & 8.8e-03 & 2.3e-04 & 2.2e-01 & 1.4e-02 \\ \hline
+        \textbf{m} & 1.8e-02 & 1.7e-02 & 4.3e-03 & 3.6e-02 & 5.4e-02 & 4.3e-02 & 3.0e-02 & 3.3e-02 & 1.6e-02 & 5.3e-02 & 3.3e-02 & 1.3e-02 & 1.4e-03 & 3.9e-02 & 1.8e-02 & 1.8e-02 & 2.9e-03 & 4.5e-03 & 2.3e-03 & 5.1e-03 & 8.9e-03 & 1.3e-02 & 9.8e-03 & 9.4e-03 & 1.1e-02 & 4.1e-03 & 1.2e-02 & 4.6e-02 & 3.3e-02 & 9.0e-04 & 8.5e-05 & 4.8e-02 & 2.9e-03 & 2.4e-04 & 3.2e-04 & 2.8e-02 & 1.5e-03 & 5.0e-04 & 1.8e-03 & 2.3e-02 & 1.3e-03 & 3.5e-02 & 7.7e-02 & 4.2e-04 & 6.4e-04 & 1.3e-02 & 8.7e-04 & 2.6e-02 & 2.4e-04 & 9.2e-04 & 2.3e-04 & 5.4e-02 & 5.5e-04 \\ \hline
+        \textbf{n} & 1.8e-02 & 6.4e-02 & 8.3e-02 & 7.6e-02 & 7.6e-02 & 6.6e-02 & 6.6e-02 & 6.3e-02 & 1.6e-02 & 6.5e-02 & 6.7e-02 & 2.7e-02 & 5.6e-03 & 5.7e-02 & 1.8e-02 & 1.8e-02 & 2.9e-03 & 4.5e-03 & 9.0e-03 & 5.1e-03 & 8.9e-03 & 1.3e-02 & 9.8e-03 & 9.4e-03 & 1.1e-02 & 4.1e-03 & 1.2e-02 & 2.9e-02 & 7.2e-03 & 1.5e-01 & 2.9e-01 & 4.8e-02 & 1.8e-02 & 4.9e-01 & 2.1e-03 & 3.4e-02 & 4.8e-02 & 7.5e-02 & 2.3e-02 & 3.5e-03 & 1.1e-02 & 6.7e-02 & 4.4e-03 & 1.1e-01 & 1.0e-03 & 4.1e-02 & 7.4e-02 & 1.5e-02 & 3.0e-02 & 4.7e-03 & 1.9e-02 & 4.6e-02 & 1.3e-02 \\ \hline
+        \textbf{o} & 1.8e-02 & 4.6e-02 & 4.3e-02 & 2.1e-02 & 1.3e-02 & 1.4e-02 & 3.6e-02 & 3.9e-02 & 1.6e-02 & 1.2e-02 & 1.7e-02 & 1.3e-02 & 2.8e-03 & 1.4e-02 & 1.8e-02 & 1.8e-02 & 2.9e-03 & 4.5e-03 & 2.3e-03 & 5.1e-03 & 8.9e-03 & 1.3e-02 & 9.8e-03 & 9.4e-03 & 1.1e-02 & 4.1e-03 & 1.2e-02 & 7.5e-03 & 2.9e-02 & 2.4e-02 & 2.7e-02 & 1.7e-03 & 3.1e-01 & 1.6e-02 & 3.3e-03 & 1.3e-02 & 4.8e-02 & 1.7e-01 & 7.8e-02 & 1.8e-01 & 1.5e-01 & 3.8e-02 & 7.0e-02 & 1.4e-02 & 1.3e-01 & 4.4e-02 & 4.6e-02 & 3.7e-01 & 2.3e-01 & 1.7e-01 & 3.2e-02 & 1.6e-02 & 6.3e-02 \\ \hline
+        \textbf{p} & 1.8e-02 & 4.4e-03 & 4.3e-03 & 9.6e-03 & 7.6e-03 & 5.7e-03 & 9.8e-03 & 5.4e-03 & 1.6e-02 & 8.8e-03 & 1.7e-02 & 1.3e-02 & 2.8e-03 & 7.1e-03 & 1.8e-02 & 1.8e-02 & 2.9e-03 & 4.5e-03 & 2.3e-03 & 5.1e-03 & 8.9e-03 & 1.3e-02 & 9.8e-03 & 9.4e-03 & 1.1e-02 & 4.1e-03 & 1.2e-02 & 2.2e-02 & 5.9e-04 & 2.2e-03 & 4.2e-05 & 2.6e-02 & 1.9e-03 & 5.9e-05 & 2.6e-03 & 2.5e-02 & 7.7e-04 & 1.0e-03 & 4.4e-02 & 5.1e-04 & 5.5e-05 & 2.5e-02 & 5.8e-02 & 4.2e-04 & 5.3e-02 & 6.9e-03 & 9.9e-03 & 1.8e-02 & 7.9e-05 & 6.8e-04 & 2.3e-04 & 7.5e-03 & 5.5e-04 \\ \hline
+        \textbf{q} & 1.8e-02 & 3.9e-06 & 5.3e-04 & 5.0e-05 & 8.4e-04 & 9.4e-04 & 2.5e-04 & 3.2e-04 & 1.6e-02 & 3.2e-05 & 1.7e-02 & 1.3e-02 & 1.4e-03 & 1.4e-03 & 1.8e-02 & 1.8e-02 & 2.9e-03 & 4.5e-03 & 2.3e-03 & 5.1e-03 & 8.9e-03 & 1.3e-02 & 9.8e-03 & 9.4e-03 & 1.1e-02 & 4.1e-03 & 1.2e-02 & 5.0e-06 & 2.9e-05 & 1.6e-05 & 8.5e-06 & 3.2e-06 & 1.8e-05 & 2.0e-05 & 6.0e-06 & 5.9e-06 & 3.9e-04 & 5.0e-05 & 1.1e-05 & 1.6e-05 & 5.5e-06 & 5.3e-06 & 2.2e-05 & 4.2e-04 & 6.8e-06 & 6.4e-06 & 4.5e-06 & 3.6e-02 & 4.0e-05 & 1.7e-05 & 2.3e-04 & 2.2e-05 & 5.5e-04 \\ \hline
+        \textbf{r} & 1.8e-02 & 4.7e-02 & 9.5e-02 & 7.1e-02 & 8.2e-02 & 5.8e-02 & 6.6e-02 & 8.3e-02 & 1.6e-02 & 6.8e-02 & 1.7e-02 & 1.3e-02 & 7.1e-03 & 5.1e-02 & 1.8e-02 & 1.8e-02 & 5.9e-03 & 4.5e-03 & 2.3e-03 & 5.1e-03 & 8.9e-03 & 1.3e-02 & 9.8e-03 & 9.4e-03 & 1.1e-02 & 4.1e-03 & 2.4e-02 & 4.1e-02 & 9.2e-03 & 2.6e-02 & 3.8e-02 & 1.1e-01 & 1.1e-02 & 2.8e-02 & 2.2e-03 & 7.7e-02 & 3.1e-03 & 4.7e-02 & 1.4e-02 & 4.9e-02 & 1.4e-02 & 7.8e-02 & 1.4e-02 & 6.8e-03 & 3.4e-02 & 5.3e-02 & 2.2e-02 & 4.5e-02 & 2.9e-02 & 6.8e-03 & 2.3e-04 & 1.3e-01 & 4.6e-02 \\ \hline
+        \textbf{s} & 1.8e-02 & 9.4e-02 & 2.7e-02 & 1.5e-01 & 1.5e-01 & 1.3e-01 & 1.3e-01 & 8.4e-02 & 3.2e-02 & 1.1e-01 & 3.3e-02 & 1.3e-02 & 1.1e-02 & 1.1e-01 & 1.8e-02 & 1.8e-02 & 5.9e-03 & 4.5e-03 & 4.5e-03 & 5.1e-03 & 8.9e-03 & 1.3e-02 & 9.8e-03 & 9.4e-03 & 1.1e-02 & 4.1e-03 & 1.2e-02 & 4.4e-02 & 1.7e-02 & 4.0e-02 & 2.0e-03 & 5.6e-02 & 9.4e-03 & 2.7e-03 & 6.6e-02 & 5.4e-02 & 7.4e-03 & 1.1e-01 & 1.4e-02 & 3.0e-02 & 3.1e-03 & 4.6e-02 & 7.6e-02 & 7.8e-02 & 1.1e-03 & 6.4e-02 & 8.1e-02 & 6.2e-02 & 4.2e-03 & 1.7e-02 & 6.8e-04 & 8.6e-03 & 2.7e-03 \\ \hline
+        \textbf{t} & 1.8e-02 & 9.4e-02 & 5.6e-02 & 8.1e-02 & 9.4e-02 & 5.8e-02 & 8.9e-02 & 1.4e-01 & 1.6e-02 & 8.0e-02 & 1.7e-02 & 1.3e-02 & 5.6e-03 & 6.4e-02 & 1.8e-02 & 1.8e-02 & 2.9e-03 & 4.5e-03 & 2.3e-03 & 5.1e-03 & 1.8e-02 & 1.3e-02 & 2.0e-02 & 9.4e-03 & 1.1e-02 & 4.1e-03 & 1.2e-02 & 3.9e-02 & 4.4e-03 & 1.3e-02 & 9.8e-04 & 6.1e-02 & 5.6e-03 & 1.4e-03 & 4.4e-01 & 8.3e-02 & 2.7e-03 & 2.4e-03 & 3.2e-02 & 6.5e-03 & 1.3e-03 & 1.3e-01 & 3.8e-03 & 5.1e-03 & 3.9e-02 & 2.7e-02 & 2.2e-02 & 4.7e-02 & 1.5e-03 & 2.5e-02 & 2.3e-04 & 7.0e-02 & 7.6e-02 \\ \hline
+        \textbf{u} & 1.8e-02 & 5.8e-03 & 2.1e-03 & 7.8e-03 & 8.4e-03 & 5.7e-03 & 2.7e-02 & 3.1e-02 & 1.6e-02 & 6.4e-03 & 1.7e-02 & 1.3e-02 & 1.4e-03 & 5.7e-03 & 1.8e-02 & 1.8e-02 & 2.9e-03 & 4.5e-03 & 2.3e-03 & 5.1e-03 & 8.9e-03 & 1.3e-02 & 9.8e-03 & 9.4e-03 & 1.1e-02 & 4.1e-03 & 1.2e-02 & 6.6e-03 & 2.6e-02 & 3.4e-02 & 1.2e-02 & 6.1e-03 & 6.2e-03 & 6.4e-02 & 3.4e-04 & 1.0e-02 & 7.7e-04 & 5.5e-03 & 6.6e-02 & 2.2e-02 & 4.8e-02 & 8.2e-04 & 6.2e-02 & 1.3e-03 & 5.4e-02 & 5.7e-02 & 4.9e-02 & 6.2e-05 & 3.2e-03 & 3.6e-04 & 9.9e-03 & 4.6e-04 & 4.1e-02 \\ \hline
+        \textbf{v} & 1.8e-02 & 3.1e-03 & 6.4e-03 & 1.4e-02 & 1.0e-02 & 6.6e-03 & 7.3e-03 & 8.2e-03 & 1.6e-02 & 9.1e-03 & 1.7e-02 & 1.3e-02 & 2.8e-03 & 1.0e-02 & 1.8e-02 & 1.8e-02 & 2.9e-03 & 4.5e-03 & 2.3e-03 & 5.1e-03 & 8.9e-03 & 1.3e-02 & 9.8e-03 & 9.4e-03 & 1.1e-02 & 4.1e-03 & 1.2e-02 & 8.5e-03 & 1.8e-04 & 3.3e-05 & 2.5e-05 & 4.6e-02 & 1.5e-04 & 2.2e-04 & 7.8e-05 & 2.8e-02 & 1.5e-03 & 4.5e-04 & 2.3e-03 & 9.8e-05 & 3.0e-03 & 7.8e-03 & 1.1e-04 & 4.2e-04 & 6.4e-04 & 1.7e-03 & 1.2e-04 & 3.3e-04 & 7.9e-05 & 3.8e-04 & 2.3e-04 & 2.3e-03 & 5.5e-04 \\ \hline
+        \textbf{w} & 1.8e-02 & 1.2e-02 & 1.2e-02 & 2.4e-02 & 2.3e-02 & 1.5e-02 & 2.3e-02 & 2.7e-02 & 1.6e-02 & 1.7e-02 & 1.7e-02 & 1.3e-02 & 2.8e-03 & 2.6e-02 & 1.8e-02 & 1.8e-02 & 2.9e-03 & 4.5e-03 & 2.3e-03 & 5.1e-03 & 8.9e-03 & 1.3e-02 & 9.8e-03 & 9.4e-03 & 1.1e-02 & 4.1e-03 & 1.2e-02 & 6.1e-02 & 4.4e-04 & 5.7e-04 & 3.2e-03 & 2.5e-02 & 1.2e-03 & 1.8e-04 & 7.2e-02 & 5.9e-02 & 7.7e-04 & 3.4e-03 & 2.2e-03 & 2.9e-04 & 1.2e-02 & 2.4e-02 & 2.7e-04 & 8.5e-04 & 3.6e-03 & 4.9e-03 & 3.3e-04 & 4.0e-04 & 1.2e-04 & 1.1e-03 & 2.3e-04 & 9.7e-04 & 5.5e-04 \\ \hline
+        \textbf{x} & 1.8e-02 & 4.7e-04 & 4.8e-03 & 9.5e-04 & 8.4e-04 & 3.8e-03 & 7.6e-04 & 3.2e-04 & 3.2e-02 & 6.5e-04 & 1.7e-02 & 1.3e-02 & 1.4e-03 & 1.4e-03 & 1.8e-02 & 1.8e-02 & 2.9e-03 & 4.5e-03 & 2.3e-03 & 5.1e-03 & 8.9e-03 & 1.3e-02 & 9.8e-03 & 9.4e-03 & 1.1e-02 & 4.1e-03 & 1.2e-02 & 1.7e-03 & 8.8e-05 & 9.9e-03 & 1.7e-05 & 8.6e-04 & 7.3e-05 & 2.0e-05 & 3.8e-04 & 3.0e-03 & 3.9e-04 & 1.0e-04 & 3.2e-05 & 3.3e-05 & 1.7e-05 & 7.9e-05 & 3.0e-02 & 2.1e-03 & 2.7e-05 & 3.2e-05 & 2.2e-03 & 3.0e-04 & 6.2e-03 & 6.9e-05 & 3.6e-02 & 1.1e-04 & 5.5e-04 \\ \hline
+        \textbf{y} & 1.8e-02 & 4.6e-02 & 8.0e-02 & 7.9e-02 & 7.8e-02 & 9.7e-02 & 6.7e-02 & 6.9e-02 & 3.2e-02 & 7.0e-02 & 1.7e-02 & 1.3e-02 & 4.2e-03 & 5.4e-02 & 1.8e-02 & 1.8e-02 & 2.9e-03 & 4.5e-03 & 2.3e-03 & 5.1e-03 & 8.9e-03 & 1.3e-02 & 9.8e-03 & 9.4e-03 & 1.1e-02 & 4.1e-03 & 1.2e-02 & 7.0e-03 & 7.3e-03 & 4.3e-03 & 9.3e-04 & 8.4e-03 & 3.4e-03 & 1.0e-03 & 8.1e-04 & 6.8e-03 & 4.3e-03 & 2.5e-03 & 2.6e-03 & 3.7e-03 & 4.5e-04 & 3.2e-02 & 2.8e-03 & 2.1e-03 & 1.0e-03 & 9.3e-03 & 4.4e-03 & 1.2e-03 & 5.1e-04 & 3.8e-03 & 2.3e-04 & 2.2e-04 & 6.6e-03 \\ \hline
+        \textbf{z} & 1.8e-02 & 1.4e-04 & 3.2e-03 & 8.0e-04 & 1.7e-03 & 1.9e-03 & 1.0e-03 & 9.5e-04 & 1.6e-02 & 4.9e-04 & 1.7e-02 & 1.3e-02 & 1.4e-03 & 2.9e-03 & 1.8e-02 & 1.8e-02 & 2.9e-03 & 4.5e-03 & 2.3e-03 & 5.1e-03 & 8.9e-03 & 1.3e-02 & 9.8e-03 & 9.4e-03 & 1.1e-02 & 4.1e-03 & 1.2e-02 & 5.0e-04 & 2.9e-05 & 3.3e-05 & 2.5e-04 & 2.3e-03 & 1.8e-05 & 2.0e-05 & 9.3e-04 & 1.4e-03 & 3.9e-04 & 2.5e-04 & 4.7e-04 & 1.0e-03 & 9.9e-05 & 3.3e-03 & 2.2e-05 & 4.2e-04 & 6.8e-06 & 1.9e-05 & 4.5e-06 & 4.2e-04 & 4.0e-05 & 6.9e-05 & 2.3e-04 & 3.3e-04 & 2.6e-02 \\ \hline
+    \end{tabular}
+}
+
+(Apologies for the tiny font, latex was being difficult)
+
+\newpage
+
+The invariant distribution $\phi$:
+
+\begin{center}
+\scalebox{0.5} {\begin{tabular}{c|c}%
+ \bfseries $Symbol$ & $Probability$% specify table head
+\csvreader[head to column names,
+/csv/separator=pipe,
+]{outputs/q5/q5a-invariant.csv}{}% use head of csv as column names
+{\\\hline\csvcoli&\csvcolii}% specify your coloumns here
+\end{tabular}
+}
+\end{center}
+
+\newpage
 \item[(b)] The latent variables $\sigma(s)$ for different symbols $s$ are not independent.
@@
 
-$$P(e_1, e_2,...,e_n|\sigma) = \psi(\gamma=\sigma^{-1}(e_1))\prod_{i=2}^n \psi(\alpha=\sigma^{-1}(e_i), \beta=\sigma^{-1}(e_{i-1}))$$
+$$P(e_1, e_2,...,e_n|\sigma) = \phi(\gamma=\sigma^{-1}(e_1))\prod_{i=2}^n \psi(\alpha=\sigma^{-1}(e_i), \beta=\sigma^{-1}(e_{i-1}))$$
 
-because $\sigma$ is the encoding function, mapping to a symbol $s$ into the encoded text as $e$, we require $\sigma^{-1}$ the decoding function mapping the encoded symbol $e$ back to $s$.
+because $\sigma$ is the encoding function, mapping a symbol $s$ into the encoded symbol $e$, we require $\sigma^{-1}$ the decoding function mapping the encoded symbol $e$ back to $s$.
 
-\item[(c)] The proposal probability $S(\sigma \rightarrow \sigma')$ depends on the permutations of $\sigma$ and $\sigma'$ because we only choose a proposal $\sigma'$ that differs at \textit{two} spots:
-
+\item[(c)] The proposal probability $S(\sigma \rightarrow \sigma')$ depends on the permutations of $\sigma$ and $\sigma'$.
+Our proposal generating process restricts us to choose a proposal $\sigma'$ that differs from $\sigma$ only at \textit{two} spots:
 $$\sigma'(s^i) = \sigma(s^j)$$
 $$\sigma'(s^j) = \sigma(s^i)$$
-
 for any two symbols $s^i$ and $s^j$ of the 53 possible symbols ($s^i \neq s^j$).
 
-Therefore, if the above doesn't hold for $\sigma'$, $S(\sigma \rightarrow \sigma')=0$, because with our method of choosing proposals, it is not possible to choose $\sigma'$.
-At $\sigma$ there are ${53 \choose 2}$ possible proposal $\sigma'$'s with the above property.
+Therefore, if the above doesn't hold for $\sigma'$, $S(\sigma \rightarrow \sigma')=0$.
+From $\sigma$ there are ${53 \choose 2}$ possible proposal $\sigma'$'s with the above property.
 Because we are assuming a uniform prior distribution over $\sigma$'s, the transition probability of a $\sigma'$ that satisfies the above property is $S(\sigma \rightarrow \sigma')=\frac{1}{{53 \choose 2}}$.
@@
 The MH acceptance probability is given as:
-
 $$A(\sigma \rightarrow \sigma'| \mathcal{D}) = \min \{1, \frac{S(\sigma' \rightarrow \sigma)P(\sigma'|\mathcal{D})}{S(\sigma \rightarrow \sigma')P(\sigma|\mathcal{D})})\}$$
-
-where $S(\sigma \rightarrow \sigma')$ is the conditional transition probability of $\sigma'$ given $\sigma$ and $\mathcal{D}$ is our encrypted text $e_1, e_2,...,e_n$.
+because $S(\sigma \rightarrow \sigma')$ is the conditional transition probability of $\sigma'$ given $\sigma$ and $\mathcal{D}$ is our encrypted text $e_1, e_2,...,e_n$.
 
 $S(\sigma \rightarrow \sigma') = S(\sigma' \rightarrow \sigma)$ for all $\sigma$ and $\sigma'$ that differ only at two spots because the probability in this case will always be $\frac{1}{{53 \choose 2}}$, we can simplify:
-
 $$A(\sigma \rightarrow \sigma'| \mathcal{D}) = \min \{1, \frac{P(\sigma'|\mathcal{D})}{P(\sigma|\mathcal{D})})\}$$
-
 From Bayes' Theorem:
-
 $$P(\sigma|\mathcal{D}) = \frac{P(\mathcal{D}|\sigma)P(\sigma)}{\sum_{\sigma'}P(\mathcal{D}|\sigma')P(\sigma')}$$
-
 We are assuming a uniform prior for $\sigma$, so $P(\sigma)$ is a constant and we can simplify further:
-
 $$A(\sigma \rightarrow \sigma'| \mathcal{D}) = \min \{1, \frac{P(\mathcal{D}|\sigma')}{P(\mathcal{D}|\sigma)})\}$$
-
 This is the acceptance probability for a given proposal $\sigma'$.
@@
 \newpage
-\item[(d)] The Python code for the MH sampler:
+\item[(d)] Reporting the current decryption of the first 60 symbols after every 100 iterations:
+
+\begin{center}
+%\begin {table}[h]
+\scalebox{0.4} {\begin{tabular}{c|l}%
+ \bfseries \textbf{MH Iteration} & \textbf{Current Decryption}% specify table head
+\csvreader[head to column names,
+/csv/separator=pipe,
+]{outputs/q5/q5d-iterations.csv}{}% use head of csv as column names
+{\\\hline\csvcoli&\csvcolii}% specify your coloumns here
+\end{tabular}
+}
+%\end{table}
+\end{center}
+
+\newpage
+The corresponding $\sigma$:
+\begin{center}
+\scalebox{0.5} {\begin{tabular}{c|c}%
+ \bfseries $\textbf{s}$ & $\textbf{\sigma(s)}$% specify table head
+\csvreader[head to column names,
+/csv/separator=pipe,
+]{outputs/q5/q5d-decrypter.csv}{}% use head of csv as column names
+{\\\hline\csvcoli&\csvcolii}% specify your coloumns here
+\end{tabular}
+}
+\end{center}
+
+To help with chain initialisation, 10000 different $\sigma$'s were randomly and independently sampled.
+The $\sigma$ providing the best log-likelihood was chosen as the starting point for the MH chain and algorithm was then run for 10000 iterations.
+Moreover, ten different trials were performed, where the trial with the best log-likelihood is displayed.
+
+\newpage
+
+The Python code for the MH sampler:
 \lstinputlisting[language=Python]{src/solutions/q5.py}
 \newpage
-Implement the MH sampler, and run it on the provided encrypted text.
-Report the current decryption of the first 60 symbols after every 100 iterations.
-Your Markov chain should converge to give you a fairly sensible message.
-(Hint: it may help to initialize your chain intelligently and to try multiple times; in any case, please describe what you did).
-[30 marks]
 
-TODO
+\newpage
 
-\item[(e)] Note that some $\Psi(\alpha, \beta)$ values may be zero.
-Does this affect the ergodicity of the chain?
-If the chain remains ergodic, give a proof; if not, explain and describe how you can restore ergodicity.
-[5 marks]
-
-TODO
+\item[(e)] When some values of $\Psi(\alpha, \beta) = 0$, this affects the ergodicity of the chain.
+An ergodic chain is one that is irreducible (i.e.
+all possible transitions between symbols have probability greater than zero).
+If $\Psi(\alpha, \beta) = 0$, this means that there is zero probability that $\beta$ will transition to $\alpha$, breaking our definition.
+To restore ergodicity, we can add a small transition probability between all symbols of the chain.
+This essentially acts as a prior, stating that the probability of a symbol to transition to any other symbol (including itself) should never be zero.
 
@@
 
-TODO
+If we were to use symbol probabilities alone for decoding, the joint probability would be:
+$$P(e_1, e_2,...,e_n|\sigma) = \prod_{i=1}^n P(\sigma^{-1}(e_i))$$
+the product of the likelihoods of the decoded letters.
+In this case, the optimal decoding would simply replace the most frequent symbols in the encrypted message with the most frequent symbols in the training text.
+This is much more difficult because each letter is assumed to be independent of its neighbours.
+For a first order Markov chairn, we exploit the structure of language by considering pairs of letters.
+Assuming that as the training text size approaches infinity and the size of the encrypted message also approaches infinity, that the two will have the same symbol frequency and that the probability of each symbol is unique, (i.e.
+two different decodings can't have the same likelihood), then using symbol probabilities alone should theoretically work.
+However, in practise we would unlikely to be able to make these assumptions about symbol frequencies from the size of our training set and encrypted message.
+
+A second-order chain should also work in theory.
+However, this approach is probably practically  more difficult for finding a suitable decoding.
+This is because our transition matrix would contain $N^3$, where $N$ is the number of symbols, to account for all possible second order transitions.
+Our training text would need to increase quadratically to maintain the same ratio of possible transitions to example transitions (number of second order transitions in a text of length $N$ is $N-2$ and third order its $N-3$).
+
+For an encryption scheme where two symbols map to the same encrypted value:
+$$\exists \alpha, \beta, \sigma(\alpha) = \sigma(\beta), \alpha \neq \beta$$
+this approach can become much more complicated.
+Our $\sigma^{-1}(e)$ is ill-defined, and therefore how we computing the joint probability of the encrypted text is no longer immediately clear.
+Moreover, generating proposal encodings is not as simple as swapping the encryption for two symbols.
+This is because we do not know which two symbols map to the same encrypted symbol and simply swapping would preserve the same collision mapping of the current encoding.
+Overall, many changes would need to be made to the approach to accommodate for these complications.
+It is not immediately obvious how current approach could work for this case.
+
+If we used this approach for Chinese with $\geq 10000$ symbols, we would be attempting to solve the same problem but with $N\geq10000$ instead of $N=53$.
+Similar to the second order Markov chain, although this is theoretically possible, it would require a transition matrix of size $\geq10000^2$ which is quite impractical.
+An alternative set up  could be with using Chinese phonetics, for which there are likely much fewer than $10000$, however this would require a mapping from a phonetic to an encrypted phonetic.
 
@@
 \item[(a)] To find the local extrema of the function $f(x,y) = x + 2y$ subject to the constraint $y^2 + xy = 1$, first we define $g(x, y)$:
-
 $$g(x, y) = y^2 + xy - 1$$
-
 where $g(x, y)=0$ is an equivalent representation of the given constraint.
@@
 We can therefore construct the optimisation problem:
-
 $$\min_{\textbf{x}} f(\textbf{x})$$
-
 such that $g(\textbf{x})=\textbf{0}$ and $\textbf{x} := [x, y]^T$.
@@
 We can calculate $\nabla f(\textbf{x})$:
-
 $$\nabla f(\textbf{x}) = [\frac{\partial}{\partial x} (x+2y), \frac{\partial}{\partial y} (x+2y)]^T$$
 $$\nabla f(\textbf{x}) = [1, 2]^T$$
-
 and calculating $\nabla g(\textbf{x})$:
-
 $$\nabla g(\textbf{x}) = [\frac{\partial}{\partial x} (y^2 + xy - 1), \frac{\partial}{\partial y} (y^2 + xy - 1)]^T$$
 $$\nabla g(\textbf{x}) = [y, 2y+x]^T$$
-
 Solving the constraint optimisation problem with Lagrange multipliers, we set up the equations:
-
 $$\nabla f(\textbf{x})+\lambda \nabla g(\textbf{x}) = \textbf{0}$$
-
 and
-
 $$g(\textbf{x}) = 0$$
-
 Giving us the three equations:
-
 $$1+\lambda y = 0$$
@@
 $$y^2 + xy - 1 = 0$$
-
-Substituting $\lambda y = -1$ from the first equation into the second equation:
-
-$$2+2(-1) + x = 0$$
-
+Substituting $ y = \frac{-1}{\lambda}$ from the first equation into the second equation:
+$$2+\frac{-1}{\lambda}(2y+x) = 0$$
+$$\frac{-x}{y} = 0$$
 We see that $x=0$.
 Solving for $y$ in our third equation with $x=0$:
-
 $$y^2-1=0$$
-
 We see that $y=\pm 1$ and from the first equation $\lambda\mp 1$.
@@
 \item[(b)] _
-
 \begin{enumerate}
@@
              $$x_{n+1} = x_{n} - \frac{f(x_n)}{f'(x_n)}$$
-             where $f(x_n) = f(x_n, a) = \exp(x_n) - a$
+             where $f(x_n) = \exp(x_n) - a$
 

### 2022-11-10 main.py

@@ -2,8 +2,14 @@ import os
 
 import numpy as np
 
-from src.constants import BINARY_DIGITS_FILE_PATH, OUTPUTS_FOLDER
-from src.solutions import q1, q2, q3
+from src.constants import (
+    BINARY_DIGITS_FILE_PATH,
+    MESSAGE_FILE_PATH,
+    OUTPUTS_FOLDER,
+    SYMBOLS_FILE_PATH,
+    TRAINING_TEXT_FILE_PATH,
+)
+from src.solutions import q1, q2, q3, q5
 
 if __name__ == "__main__":
 
@@ -47,4 +53,38 @@ if __name__ == "__main__":
         max_number_of_steps=int(1e2),
         figure_path=os.path.join(Q3_OUTPUT_FOLDER, "q3e"),
         figure_title="Q3e",
+    )
+
+    # Question 5
+    Q5_OUTPUT_FOLDER = os.path.join(OUTPUTS_FOLDER, "q5")
+    if not os.path.exists(Q5_OUTPUT_FOLDER):
+        os.makedirs(Q5_OUTPUT_FOLDER)
+
+    with open(TRAINING_TEXT_FILE_PATH) as fp:
+        training_text = fp.read().replace("\n", "").lower()
+    with open(SYMBOLS_FILE_PATH) as fp:
+        symbols = fp.read().split("\n")
+    with open(MESSAGE_FILE_PATH) as fp:
+        encrypted_message = fp.read()
+
+    q5.a(
+        symbols,
+        training_text,
+        transition_matrix_path=os.path.join(Q5_OUTPUT_FOLDER, "q5a-transition.csv"),
+        invariant_distribution_path=os.path.join(Q5_OUTPUT_FOLDER, "q5a-invariant.csv"),
+    )
+
+    q5.d(
+        encrypted_message,
+        symbols,
+        training_text,
+        number_trials=10,
+        number_of_mh_loops=int(1e4),
+        number_start_attempts=int(1e4),
+        check_decryption_interval=100,
+        check_decryption_size=60,
+        decryptor_table_path=os.path.join(Q5_OUTPUT_FOLDER, "q5d-decrypter.csv"),
+        decrypted_message_iterations_table_path=os.path.join(
+            Q5_OUTPUT_FOLDER, "q5d-iterations.csv"
+        ),
     )

### 2022-11-10 src/solutions/q3.py

@@ -51,15 +51,13 @@ class Theta:
         return np.repeat(self.log_pi, n, axis=0)
 
 
-def _init_params(k: int, d: int, seed: int = DEFAULT_SEED) -> Theta:
+def _init_params(k: int, d: int) -> Theta:
     """
     Random initialisation of theta parameters (log_pi and log_p_matrix)
     :param k: Number of components
     :param d: Image dimension (number of pixels in a single image)
-    :param seed: seed initialisation for random methods
     :return: theta: the parameters of the model
     """
-    np.random.seed(seed)
     return Theta(
         log_pi=np.log(np.random.dirichlet(np.ones(k), size=1)),
         log_p_matrix=np.log(np.random.uniform(low=0, high=1, size=(d, k))),
@@ -146,36 +144,20 @@ def _compute_log_p_matrix_hat(
         log_responsibility[:, np.newaxis, :], d, axis=1
     )  # (n, d, k)
 
-    log_p_matrix_unnormalised_likelihood = logsumexp(
-        log_responsibility_repeated, b=x_repeated, axis=0
-    )  # (d, k)
-    log_p_matrix_normaliser_likelihood = np.array(
-        logsumexp(log_responsibility_repeated, axis=0)
-    )  # (d, k)
-
     alpha = 2
     beta = 2
+
     log_p_matrix_unnormalised_posterior = logsumexp(
-        np.stack(
-            (
-                (alpha - 1) * np.ones(log_p_matrix_unnormalised_likelihood.shape),
-                log_p_matrix_unnormalised_likelihood,
-            ),
-            axis=0,
-        ),
-        axis=0,
-    )
+        log_responsibility_repeated, b=(x_repeated + alpha - 1), axis=0
+    )  # (d, k)
+
     log_p_matrix_normaliser_posterior = logsumexp(
-        np.stack(
-            (
-                (alpha + beta - 2) * np.ones(log_p_matrix_normaliser_likelihood.shape),
-                log_p_matrix_normaliser_likelihood,
-            ),
-            axis=0,
-        ),
-        axis=0,
+        log_responsibility_repeated, b=(alpha + beta - 1), axis=0
+    )  # (d, k)
+
+    log_p_matrix_normalised_posterior = (
+        log_p_matrix_unnormalised_posterior - log_p_matrix_normaliser_posterior
     )
-    log_p_matrix_normalised_posterior = log_p_matrix_unnormalised_posterior - log_p_matrix_normaliser_posterior
[…]

### 2022-11-10 src/solutions/q5.py

@@ -1,0 +1,286 @@
+from typing import Dict, List, Tuple
+
+import numpy as np
+import pandas as pd
+from sklearn.preprocessing import normalize
+
+from src.constants import DEFAULT_SEED
+
+
+class Decrypter:
+    def __init__(self, decryption_dict):
+        self.decryption_dict = decryption_dict
+
+    def decrypt(self, encrypted_message):
+        return "".join([self.decryption_dict[x] for x in encrypted_message])
+
+
+class Statistics:
+    def __init__(
+        self,
+        training_text: str,
+        symbols: List[str],
+        invariant_stopping_epsilon: float = 5e-20,
+    ):
+        self.training_text = training_text
+        self.symbols = symbols
+        self.num_symbols = len(symbols)
+        self.symbols_dict = {k: v for v, k in enumerate(symbols)}
+        self.text_numbers = [
+            self.symbols_dict[symbol]
+            for symbol in list(training_text)
+            if symbol in self.symbols_dict
+        ]
+        self.transition_matrix = self._construct_transition_matrix(
+            training_text, self.symbols_dict
+        )
+        self.invariant_distribution = self._approximate_invariant_distribution(
+            invariant_stopping_epsilon
+        )
+        self.log_transition_matrix = np.log(self.transition_matrix)
+        self.log_invariant_distribution = np.log(self.invariant_distribution)
+
+    def _construct_transition_matrix(
+        self, training_text: str, symbols_dict: Dict[str, int]
+    ) -> np.ndarray:
+
+        # initialise with ones to ensure ergodicity
+        transition_matrix = np.ones((self.num_symbols, self.num_symbols))
+        for i in range(1, len(training_text)):
+            # check symbols are valid
+            if (
+                training_text[i] in symbols_dict
+                and training_text[i - 1] in symbols_dict
+            ):
+                transition_matrix[
+                    symbols_dict[training_text[i - 1]], symbols_dict[training_text[i]]
+                ] += 1
+        # normalise to get transition probabilities
+        transition_matrix = normalize(transition_matrix, axis=0, norm="l1")
+        return transition_matrix
+
+    def _approximate_invariant_distribution(
+        self, invariant_stopping_epsilon: float
+    ) -> np.ndarray:
+        invariant_distribution = np.zeros((self.num_symbols, 1))
+        previous_invariant_distribution = invariant_distribution.copy()
+        invariant_distribution[0] = 1
+
[…]

### 2022-11-11 submission.tex
comment code and refactor

@@
 \end{enumerate}
+\end{enumerate}
+
+\newpage
+\section*{Question 8}
+\begin{enumerate}
+\item[(a)] For:
+$$\sup_{\{\textbf{x}\in \mathbb{R}^n\}} R_A(\textbf{x})$$
+where $R_A(\textbf{x}) = \frac{\textbf{x}^T\textbf{Ax}}{\|\textbf{x}\|^2}$, we want to show that a maximum is attained.
+
+To do this, we will show the above optimisation can be equivalently formulated:
+$$\sup_{\{\textbf{x}\in \mathbb{R}^n | \| \textbf{x}\|=1\}} R_A(\textbf{x})$$
+We begin by considering any $\textbf{w} \in \mathbb{R}^n$ and let $\textbf{x} = \frac{\textbf{w}}{\|\textbf{w}\|}$. Because $\|\textbf{x}\| = 1$ we can substitute:
+$$\sup_{\{\frac{\textbf{w}}{\|\textbf{w}\|}\in \mathbb{R}^n | \| \frac{\textbf{w}}{\|\textbf{w}\|}\|=1\}} R_A(\textbf{x}) = \sup_{\textbf{x}\in \mathbb{R}^n | \| \textbf{x}\|=1} \frac{\textbf{w}^T\textbf{Aw} \|\textbf{w}\|^2}{\|\textbf{w}\|^2 \textbf{w}^T\textbf{w}}$$
+where $\|\textbf{x}\|^2 = \textbf{x}^T\textbf{x}$.
+
+The set $\{\frac{\textbf{w}}{\|\textbf{w}\|}\in \mathbb{R}^n | \| \frac{\textbf{w}}{\|\textbf{w}\|}\|=1\}$ holds for all $\textbf{w} \in \mathbb{R}^n$ so we can rewrite:
+$$\sup_{\{\textbf{w}\in \mathbb{R}^n\}} \frac{\textbf{w}^T\textbf{Aw} \|\textbf{w}\|^2}{\|\textbf{w}\|^2 \textbf{w}^T\textbf{w}}$$
+We can simplify the expression:
+$$\sup_{\{\textbf{w}\in \mathbb{R}^n\}} \frac{\textbf{w}^T\textbf{Aw} }{\textbf{w}^T\textbf{w}}$$
+$$\sup_{\{\textbf{w}\in \mathbb{R}^n\}} R_A(\textbf{w})$$
+
+and recover our original optimisation problem by letting $\textbf{x}=\textbf{w}$, showing that it is equivalent to the supremum over the unit sphere.
+Assuming the set containing the unit sphere is compact, the extreme value theory of calculus states that $\sup_{\{\textbf{x}\in \mathbb{R}^n | \| \textbf{x}\|=1\}} R_A(\textbf{x})$ is attained so equivalently $\sup_{\{\textbf{x}\in \mathbb{R}^n\}} R_A(\textbf{x})$ is attained as required.
+
+\item[(b)] We can now reformulate the optimisation as:
+$$\sup_{\{\textbf{x}\in \mathbb{R}^n | \| \textbf{x}\|=1\}} \frac{\textbf{x}^T\textbf{Ax}}{\|\textbf{x}\|^2}$$
+But because $\| \textbf{x}\|=1$, we can equivalently write:
+$$\sup_{\{\textbf{x}\in \mathbb{R}^n | \| \textbf{x}\|=1\}} \textbf{x}^T\textbf{Ax}}$$
+Thus, showing $\textbf{x}^T\textbf{Ax}} \leq \lambda_1$ will be equivalent to showing $R_A(\textbf{x}) \leq \lambda_1$ for $\|\textbf{x}\|=1$.
+We know that for all $\textbf{x} \in \mathbb{R}^n$:
+$$\textbf{x} = \sum_{i=1}^n (\xi_i^T \textbf{x}) \xi_i$$
+so we can write:
+$$\textbf{x}^T\textbf{Ax} =\bigg(\sum_{i=1}^n (\xi_i^T \textbf{x}) \xi_i^T\bigg)\textbf{A}\bigg(\sum_{i=1}^n (\xi_i^T \textbf{x}) \xi_i\bigg)$$
+Given that $\xi_i$ are eigenvectors of $\textbf{A}$ corresponding to eigenvalues $\lambda_i$:
+$$\textbf{x}^T\textbf{Ax} =\bigg(\sum_{i=1}^n (\xi_i^T \textbf{x}) \xi_i^T\bigg)\bigg(\sum_{i=1}^n \lambda_i(\xi_i^T \textbf{x}) \xi_i\bigg)$$
+Given that the eigenvectors $\xi_i$ form an orthonormal basis, we know that $\xi_i^T\xi_j = 0$ when $i\neq j$ and $\xi_i^T\xi_j = 1$ when $i=j$, so:
+$$\textbf{x}^T\textbf{Ax} =\sum_{i=1}^n \lambda_i(\xi_i^T \textbf{x})^2$$
+We know that $\|\textbf{x}\|^2 = 1$ so $\|\textbf{x}\|^2 = \sum_{j=1}^{n} x_i^2 = \sum_{j=1}^{n} (\xi_j \textbf{x})^2 = 1$.
+Thus the quantity $\sum_{i=1}^n \lambda_i(\xi_i^T \textbf{x})^2$ is simply a weighted average of $\lambda_i$'s and we can write:
+$$\textbf{x}^T\textbf{Ax} =\sum_{i=1}^n \lambda_i(\xi_i^T \textbf{x})^2 \leq \lambda_1$$
+where $\lambda_1$ is the largest eigenvalue of eigenvalues $\lambda_i$. Therefore, $R_A(\textbf{x}) \leq \lambda_1$ as required.
+
+\item[(c)] Given that $\lambda_j < \lambda_1 \forall j > k$ and $\textbf{x} \in span\{\xi_{k+1}, ..., xi_{n}\}$, we can rewrite $\textbf{x}$:
+$$\textbf{x} = \sum_{i=k+1}^n (\xi_i^T \textbf{x}) \xi_i$$
+From the same argument as (b) we can bound $\textbf{x}^T\textbf{Ax}}$:
+$$\textbf{x}^T\textbf{Ax}} = \sum_{i=k+1}^n \lambda_i(\xi_i^T \textbf{x})^2 \leq \max\{\lambda_{k+1}, ..., \lambda_n\}$$
+But given that the maximum eigenvalue $\lambda_1$ is not contained in $\{\lambda_{k+1}, ..., \lambda_n\}$:
+$$\max\{\lambda_{k+1}, ..., \lambda_n\} < \lambda_1$$
+and therefore,
+$$R_A(\textbf{x}) < \lambda_1$$
+as required.
 

### 2022-11-11 main.py

@@ -12,16 +12,14 @@ from src.constants import (
 from src.solutions import q1, q2, q3, q5
 
 if __name__ == "__main__":
-
     if not os.path.exists(OUTPUTS_FOLDER):
         os.makedirs(OUTPUTS_FOLDER)
-
     x = np.loadtxt(BINARY_DIGITS_FILE_PATH)
+
     # Question 1
     Q1_OUTPUT_FOLDER = os.path.join(OUTPUTS_FOLDER, "q1")
     if not os.path.exists(Q1_OUTPUT_FOLDER):
         os.makedirs(Q1_OUTPUT_FOLDER)
-
     q1.d(
         x,
         figure_path=os.path.join(Q1_OUTPUT_FOLDER, "q1d.png"),
@@ -59,21 +57,18 @@ if __name__ == "__main__":
     Q5_OUTPUT_FOLDER = os.path.join(OUTPUTS_FOLDER, "q5")
     if not os.path.exists(Q5_OUTPUT_FOLDER):
         os.makedirs(Q5_OUTPUT_FOLDER)
-
     with open(TRAINING_TEXT_FILE_PATH) as fp:
         training_text = fp.read().replace("\n", "").lower()
     with open(SYMBOLS_FILE_PATH) as fp:
         symbols = fp.read().split("\n")
     with open(MESSAGE_FILE_PATH) as fp:
         encrypted_message = fp.read()
-
     q5.a(
         symbols,
         training_text,
         transition_matrix_path=os.path.join(Q5_OUTPUT_FOLDER, "q5a-transition.csv"),
         invariant_distribution_path=os.path.join(Q5_OUTPUT_FOLDER, "q5a-invariant.csv"),
     )
-
     q5.d(
         encrypted_message,
         symbols,
@@ -81,8 +76,8 @@ if __name__ == "__main__":
         number_trials=10,
         number_of_mh_loops=int(1e4),
         number_start_attempts=int(1e4),
-        check_decryption_interval=100,
-        check_decryption_size=60,
+        log_decryption_interval=100,
+        log_decryption_size=60,
         decryptor_table_path=os.path.join(Q5_OUTPUT_FOLDER, "q5d-decrypter.csv"),
         decrypted_message_iterations_table_path=os.path.join(
             Q5_OUTPUT_FOLDER, "q5d-iterations.csv"

### 2022-11-11 src/solutions/q1.py

@@ -4,7 +4,9 @@ import numpy as np
 
 def _compute_maximum_likelihood_estimate(x: np.ndarray) -> np.ndarray:
     """
-    X: numpy array of shape (N, D)
+    Calculates MLE of images
+    :param x: numpy array of shape (N, D)
+    :return: MLE estimate
     """
     return np.mean(x, axis=0)
 
@@ -13,16 +15,25 @@ def _compute_maximum_a_priori_estimate(
     x: np.ndarray, alpha: float, beta: float
 ) -> np.ndarray:
     """
-    X: numpy array of shape (N, D)
-    alpha: param of prior distribution
-    beta: param of prior distribution
+    Calculates MAP estimate of images
+    :param x: numpy array of shape (N, D)
+    :param alpha: param of prior distribution
+    :param beta: param of prior distribution
+    :return: MAP estimate
     """
 
     n, _ = x.shape
     return (alpha - 1 + np.sum(x, axis=0)) / (n + alpha + beta - 2)
 
 
-def d(x, figure_path, figure_title):
+def d(x: np.ndarray, figure_path: str, figure_title: str) -> None:
+    """
+    Produces answers for question 1d
+    :param x: numpy array of shape (N, D)
+    :param figure_path: path to store figure
+    :param figure_title: figure title
+    :return:
+    """
     maximum_likelihood = _compute_maximum_likelihood_estimate(x)
     plt.figure()
     plt.imshow(
@@ -35,7 +46,18 @@ def d(x, figure_path, figure_title):
     plt.savefig(figure_path)
 
 
-def e(x, alpha, beta, figure_path, figure_title):
+def e(
+    x: np.ndarray, alpha: float, beta: float, figure_path: str, figure_title: str
+) -> None:
+    """
+    Produces answers for question 1e
+    :param x: numpy array of shape (N, D)
+    :param alpha: param of prior distribution
+    :param beta: param of prior distribution
+    :param figure_path: path to store figure
+    :param figure_title: figure title
+    :return:
+    """
     maximum_a_priori = _compute_maximum_a_priori_estimate(x, alpha, beta)
     plt.figure()
     plt.imshow(

### 2022-11-11 src/solutions/q2.py

@@ -3,24 +3,44 @@ import pandas as pd
 from scipy.special import betaln, logsumexp
 
 
-def _log_p_d_given_m1(x):
+def _log_p_d_given_m1(x: np.ndarray) -> float:
+    """
+    Calculates log likelihood of model 1
+    :param x: numpy array of shape (N, D)
+    :return: log likelihood
+    """
     n, d = x.shape
     return n * d * np.log(0.5)
 
 
-def _log_p_d_given_m2(x):
+def _log_p_d_given_m2(x: np.ndarray):
+    """
+    Calculates log likelihood of model 2
+    :param x: numpy array of shape (N, D)
+    :return: log likelihood
+    """
     n, d = x.shape
     k = np.sum(x, axis=0).astype(int)
     return betaln(np.sum(k) + 1, n * d - np.sum(k) + 1)
 
 
-def _log_p_d_given_m3(x):
+def _log_p_d_given_m3(x: np.ndarray):
+    """
+    Calculates log likelihood of model 3
+    :param x: numpy array of shape (N, D)
+    :return: log likelihood
+    """
     n, _ = x.shape
     k = np.sum(x, axis=0).astype(int)
     return logsumexp(betaln(k + 1, n - k + 1))
 
 
-def c(x, table_path):
+def _log_p_model_given_data(x) -> np.ndarray:
+    """
+    Calculates posterior log likelihood of models given image data
+    :param x: numpy array of shape (N, D)
+    :return: posterior log likelihood
+    """
     log_p_d_given_m = np.array(
         [
             _log_p_d_given_m1(x),
@@ -28,14 +48,25 @@ def c(x, table_path):
             _log_p_d_given_m3(x),
         ]
     )
-    log_p_m_given_d = log_p_d_given_m - logsumexp(log_p_d_given_m)
+    log_p_m_given_data = log_p_d_given_m - logsumexp(log_p_d_given_m)
+    return log_p_m_given_data
+
+
+def c(x: np.ndarray, table_path: str) -> None:
+    """
+    Produces answers for question 2c
+    :param x: numpy array of shape (N, D)
+    :param table_path: path to store table posterior likelihoods
+    :return:
+    """
+    log_p_m_given_data = _log_p_model_given_data(x)
     df = pd.DataFrame(
         data=np.array(
             [
-                np.arange(len(log_p_m_given_d)).astype(int) + 1,
[…]

### 2022-11-11 src/solutions/q3.py

@@ -12,6 +12,7 @@ from src.constants import DEFAULT_SEED
 @dataclass
 class Theta:
     """
+    Data class containing the model parameters
     log_pi: the logarithm of the mixing proportions (1, k)
     log_p_matrix: the logarithm of the probability where the (i,j)th element is the probability that
                   pixel j takes value 1 under mixture component i (d, k)
@@ -21,11 +22,19 @@ class Theta:
     log_p_matrix: np.ndarray
 
     @property
-    def pi(self):
+    def pi(self) -> np.ndarray:
+        """
+        Calculates the mixing proportions
+        :return: vector of mixing proportions (1, k)
+        """
         return np.exp(self.log_pi)
 
     @property
-    def p_matrix(self):
+    def p_matrix(self) -> np.ndarray:
+        """
+        Calculates the Bernoulli parameters
+        :return: matrix Bernoulli parameters (d, k)
+        """
         d, k = self.log_p_matrix.shape
         image_dimension = int(np.sqrt(d))
         return np.exp(self.log_p_matrix).reshape(image_dimension, image_dimension, -1)
@@ -201,9 +210,17 @@ def _run_expectation_maximisation(
     return theta, log_responsibility, log_likelihoods
 
 
-def _plot_p_matrix(
+def _visualise_p_matrix(
     thetas: List[Theta], ks: List[int], figure_title: str, figure_path: str
-):
+) -> None:
+    """
+    Visualises the P matrix for different thetas and ks
+    :param thetas: list of Theta instances
+    :param ks: list of k values used for each Theta
+    :param figure_title: name of figure
+    :param figure_path: path to store figure
+    :return:
+    """
     n = len(ks)
     m = np.max(ks)
     fig = plt.figure()
@@ -237,12 +254,20 @@ def _plot_p_matrix(
     plt.savefig(figure_path)
 
 
-def _plot_tsne_responsibility_clusters(
+def _visualise_responsibility_clusters(
     log_responsibilities: List[np.ndarray],
     ks: List[int],
     figure_title: str,
     figure_path: str,
-):
+) -> None:
+    """
+    Visualise responsibility vectors of images using TSNE for different k values
+    :param log_responsibilities: list of log responsibilities for different ks
+    :param ks: list of k values used for each Theta
+    :param figure_title: name of figure
+    :param figure_path: path to store figure
+    :return:
+    """
     n = len(ks)
     fig = plt.figure()
     fig.set_figwidth(5 * n)
@@ -272,6 +297,15 @@ def _plot_log_posteriors(
[…]

### 2022-11-11 src/solutions/q5.py

@@ -7,12 +7,43 @@ from sklearn.preprocessing import normalize
 from src.constants import DEFAULT_SEED
 
 
+def _convert_to_scientific_notation(x: float) -> str:
+    """
+    Convert value to string in scientific notation
+    :param x: value to convert
+    :return: string of x in scientific notation
+    """
+    return "{:.1e}".format(float(x))
+
+
 class Decrypter:
-    def __init__(self, decryption_dict):
+    def __init__(self, decryption_dict: Dict[str, str]) -> None:
+        """
+        Decrypter containing the mapping a symbol to its encrypted symbol
+        :param decryption_dict:
+        """
         self.decryption_dict = decryption_dict
 
-    def decrypt(self, encrypted_message):
+    def decrypt(self, encrypted_message: str) -> str:
+        """
+        Decrypts an encrypted message using the decryption dictionary
+        :param encrypted_message: the encrypted message to decrypt
+        :return: decrypted message
+        """
         return "".join([self.decryption_dict[x] for x in encrypted_message])
+
+    @property
+    def table(self) -> pd.DataFrame:
+        """
+        Generate table containing symbol decryptions
+        :return: pandas table of decryptions
+        """
+        decrpyter_table = pd.DataFrame(
+            self.decryption_dict.items(), columns=["s", "sigma(s)"]
+        )
+        decrpyter_table[decrpyter_table == " "] = "space"
+        decrpyter_table[decrpyter_table == '"'] = "double quotes"
+        return decrpyter_table.set_index("s")
 
 
 class Statistics:
@@ -21,16 +52,17 @@ class Statistics:
         training_text: str,
         symbols: List[str],
         invariant_stopping_epsilon: float = 5e-20,
-    ):
+    ) -> None:
+        """
+        Statistics for text
+        :param training_text: training text for calculating transition and invariant probability
+        :param symbols: symbols in the training text
+        :param invariant_stopping_epsilon: stopping condition for constructing the invariant distribution
+        """
         self.training_text = training_text
         self.symbols = symbols
         self.num_symbols = len(symbols)
-        self.symbols_dict = {k: v for v, k in enumerate(symbols)}
-        self.text_numbers = [
-            self.symbols_dict[symbol]
-            for symbol in list(training_text)
-            if symbol in self.symbols_dict
-        ]
+        self.symbols_dict = self._construct_symbols_dictionary(symbols)
         self.transition_matrix = self._construct_transition_matrix(
[…]

### 2022-11-13 submission.tex
add constants.py to submission
edit submission and comment solutions

--- 2022-11-10 submission.tex
@@
 
-\item[(a)] Our sample space for images is $\{0, 1\}^D$, where each of our D dimensions can only take binary values, D being the number of pixels in the image.
+\item[(a)] Our sample space for images is $\{0, 1\}^D$, where each of our D dimensions can only take binary values (D being the number of pixels in the image).
 The exponential family best suited on this sample space is the D-dimensional multivariate Bernoulli distribution because it shares the same sample space.
 On the other hand, a D-dimensional multivariate Gaussian has the sample space $\mathbb{R}^D$, which does not match the sample space of our data.
-To match our data sample space, we might have to define an additional mapping between our data and model spaces, adding unnecessary complexity.
+%To match our data sample space, we might have to define an additional mapping between our data and model spaces, adding unnecessary complexity.
+It is not immediately clear how the likelihood of an image of binary (discrete) values would be calculated under the continuous distribution of a multivariate Gaussian.
 Thus it would be inappropriate to model this dataset of images with a multivariate Gaussian.
 
-\item[(b)] For $\{\textbf{x}^{(n)}\}_{n=1}^N$, a data set of N images, the joint likelihood (assuming images are independently and identically distributed) is the product of N D-dimensional multivariate Bernoulli distributions, one for each image:
+\item[(b)] For $\{\textbf{x}^{(n)}\}_{n=1}^N$, a data set of N images, the joint likelihood (assuming images are independently and identically distributed) is the product of N, D-dimensional multivariate Bernoulli distributions:
 $$P(\{\textbf{x}^{(n)}\}_{n=1}^N|\textbf{p}) = \prod_{n=1}^N P(\textbf{x}^{(n)} | \textbf{p})$$
@@
 $$P(\{\textbf{x}^{(n)}\}_{n=1}^N|\textbf{p}) = \prod_{n=1}^{N}\prod_{d=1}^{D} p_d^{x_d^{(n)}} (1-p_d)^{1-x_d^{(n)}}$$
-Taking the logarithm of this, we get the log likelihood:
+Taking the logarithm, we get the log likelihood:
 $$\mathcal{L}(\{\textbf{x}^{(n)}\}_{n=1}^N|\textbf{p}) = \sum_{n=1}^{N}\sum_{d=1}^{D} [x_d^{(n)}\log(p_d) +  (1-x_d^{(n)})\log(1-p_d)]$$
@@
 $$\frac{\partial\mathcal{L}(\{\textbf{x}^{(n)}\}_{n=1}^N|\textbf{p})}{\partial p_d} = \frac{\sum_{n=1}^{N} x_d^{(n)}}{p_d} -  \frac{\sum_{n=1}^{N} (1-x_d^{(n)})}{1-p_d}$$
-and set the derivative to zero and solve for $\hat p_d$:
+and set the derivative to zero to solve for $\hat p_d$:
 $$\frac{\sum_{n=1}^{N} x_d^{(n)}}{\hat p_d} -  \frac{\sum_{n=1}^{N} (1-x_d^{(n)})}{1-\hat p_d} = 0$$
@@
 $$  \hat p_d = \frac{1}{N}\sum_{n=1}^{N} x_d^{(n)}$$
-Moreover, the second derivative with respect to $p_d$:
-$$\frac{\partial\mathcal{L}(\{\textbf{x}^{(n)}\}_{n=1}^N|\textbf{p})}{\partial p_d^2} = \frac{-\sum_{n=1}^{N} x_d^{(n)}}{p_d^2} +  \frac{-\sum_{n=1}^{N} (1-x_d^{(n)})}{(1-p_d)^2}$$
-For a maximum, we need to show that the second derivative is negative.
-Since $x_d^{(n)} \in \{0, 1\}$, in the worst case, of $N=1$, the single pixel $x_d^{(1)}$ must either be white ($\sum_{n=1}^{N} x_d^{(n)}> 0$) or black ($\sum_{n=1}^{N} 1-x_d^{(n)} > 0$) with the other being zero,  $\frac{\partial\mathcal{L}(\{\textbf{x}^{(n)}\}_{n=1}^N|\textbf{p})}{\partial p_d^2} < 0$ will be guaranteed and $\hat p_d$ is a maximum as required for the maximum likelihood estimate.
+%Moreover, the second derivative with respect to $p_d$:
+%$$\frac{\partial\mathcal{L}(\{\textbf{x}^{(n)}\}_{n=1}^N|\textbf{p})}{\partial p_d^2} = \frac{-\sum_{n=1}^{N} x_d^{(n)}}{p_d^2} +  \frac{-\sum_{n=1}^{N} (1-x_d^{(n)})}{(1-p_d)^2}$$
+%For a maximum, we need to show that the second derivative is negative.
+%Since $x_d^{(n)} \in \{0, 1\}$, in the worst case, of $N=1$, the single pixel $x_d^{(1)}$ must either be white ($\sum_{n=1}^{N} x_d^{(n)}> 0$) or black ($\sum_{n=1}^{N} 1-x_d^{(n)} > 0$) with the other being zero, so  $\frac{\partial\mathcal{L}(\{\textbf{x}^{(n)}\}_{n=1}^N|\textbf{p})}{\partial p_d^2} < 0$ will be guaranteed and $\hat p_d$ is a maximum as required for the maximum likelihood estimate.
 
@@
 $$\frac{\partial\mathcal{L}(\{\textbf{x}^{(n)}\}_{n=1}^N|\textbf{p})}{\partial p_d} = \frac{\sum_{n=1}^{N} x_d^{(n)}}{p_d} -  \frac{\sum_{n=1}^{N} (1-x_d^{(n)})}{1-p_d}$$
-For the second term $\frac{\partial\mathcal{L}(\textbf{p})}{\partial p_d}$, we start with $P(\textbf{p}))$, assuming each pixel to have an independent prior:
+For the second term $\frac{\partial\mathcal{L}(\textbf{p})}{\partial p_d}$, we start with $P(\textbf{p})$, assuming each pixel to have an independent prior:
 $$P(\textbf{p}) = \prod_{d=1}^D P(p_d)$$
-Assuming a Beta prior on each $p_d$:
+and assuming a Beta prior on each $p_d$:
 $$P(\textbf{p}) = \prod_{d=1}^D \frac{1}{B(\alpha, \beta)} p^{\alpha-1}_d (1-p_d)^{\beta-1}$$
@@
 $$\hat{p_d} =  \frac{\alpha - 1 +\sum_{n=1}^{N} x_d^{(n)}}{(N+\alpha+\beta-2)}$$
-To show that this is a maximum, the second derivative is:
-$$\frac{\partial^2\mathcal{L}(\textbf{p}|\{\textbf{x}^{(n)}\}_{n=1}^N)}{(\partial p_d)^2} = \frac{(1-\alpha) - \sum_{n=1}^{N} x_d^{(n)}}{(p_d)^2} +  \frac{(1-\beta) - \sum_{n=1}^{N} (1-x_d^{(n)})}{(1-p_d)^2}$$.
-For a maximum, we need $\frac{\partial^2\mathcal{L}(\textbf{p}|\{\textbf{x}^{(n)}\}_{n=1}^N)}{(\partial p_d)^2} < 0$ meaning that we need at least one of the strict inequalities $\alpha < 1- \sum_{n=1}^{N} x_d^{(n)}$ or $\beta < 1-\sum_{n=1}^{N} (1-x_d^{(n)})$ to be satisified, where the other can be $\leq$.
+%To show that this is a maximum, the second derivative is:
+%$$\frac{\partial^2\mathcal{L}(\textbf{p}|\{\textbf{x}^{(n)}\}_{n=1}^N)}{(\partial p_d)^2} = \frac{(1-\alpha) - \sum_{n=1}^{N} x_d^{(n)}}{(p_d)^2} +  \frac{(1-\beta) - \sum_{n=1}^{N} (1-x_d^{(n)})}{(1-p_d)^2}$$.
+%For a maximum, we need $\frac{\partial^2\mathcal{L}(\textbf{p}|\{\textbf{x}^{(n)}\}_{n=1}^N)}{(\partial p_d)^2} < 0$ meaning that we need at least one of the strict inequalities $\alpha < 1- \sum_{n=1}^{N} x_d^{(n)}$ or $\beta < 1-\sum_{n=1}^{N} (1-x_d^{(n)})$ to be satisified, where the other can be $\leq$.
 The Beta distribution requires $\alpha > 0$ and $\beta > 0$ so this requirement will always be satisfied (in the worst case of a single image, either $x_d^{(1)}=1$ or $1-x_d^{(1)}=1$).
@@
     $$\hat{\textbf{p}}^{MAP}  =  \frac{\alpha - 1 +\sum_{n=1}^{N} \textbf{x}^n}{(N+\alpha+\beta-2)}$$
-As the number of data points increases, $\hat{\textbf{p}}^{MAP}$ approaches $\frac{1}{N}\sum_{n=1}^{N} \textbf{x}^{(n)}$, $\hat{\textbf{p}}^{MLE}$.
-This makes sense because as our data set gets bigger, we are less reliant on our prior.
+As the number of data points increases, $\hat{\textbf{p}}^{MAP}$ approaches $\frac{1}{N}\sum_{n=1}^{N} \textbf{x}^{(n)}$, the $\hat{\textbf{p}}^{MLE}$.
+This makes sense because as our data set gets bigger, the effect of the prior diminishes.
 However, if a specific pixel in all of the images of our data set are white or all black, the MLE for that pixel would either be 1 or 0.
@@
 Priors can also help ensure numerical stability during calculations.
-The logarithm of zero is negative infinity, so having if the MLE is zero it can be problematic for log-likelihoods calculations whereas MAP can ensure non-zero probabilities.
+The logarithm of zero is negative infinity, so having if the MLE is zero it can be problematic for log-likelihood calculations whereas MAP can ensure non-zero probabilities.
 Interestingly, when $\alpha=\beta=1$, $\hat{\textbf{p}}^{MLE} = \hat{\textbf{p}}^{MAP}$.
@@
 
-On the other hand, a mis-specified prior can be problematic, as the estimated parameters might be skewed by the prior and not properly represent the underlying data generating process, this can result in parameter estimates that are worse than using the MLE if our data set is limited.
+On the other hand, a mis-specified prior can be problematic, as the estimated parameters might be skewed by the prior and not properly represent the underlying data generating process, this can result in parameter estimates that are `worse' than using the MLE if our data set is limited in size.
 %In this example, we're using images of different digits as part of our data set.
@@
 
-When all D components are generated from a Bernoulli distribution with $p_d = 0.5$, we have the likelihood function for model $M_1$:
-$$P(\textbf{x}^{(n)|\textbf{p}^{(1)}}=[0.5, 0.5,...,0.5]^T, M_1) = \prod_{n=1}^{N}\prod_{d=1}^{D} (0.5)^{x_d^{(n)}} (0.5)^{1-x_d^{(n)}}$$
+\item[] When all D components are generated from a Bernoulli distribution with $p_d = 0.5$, we have the likelihood function for model $M_1$:
+$$P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\textbf{p}^{(1)}}=[0.5, 0.5,...,0.5]^T, M_1) = \prod_{n=1}^{N}\prod_{d=1}^{D} (0.5)^{x_d^{(n)}} (0.5)^{1-x_d^{(n)}}$$
+Knowing that either $x_d^{(n)}$ or $1-x_d^{(n)}$ will be 1 and the other zero, we can simplify $(0.5)^{x_d^{(n)}} (1-0.5)^{1-{x_d}^{(n)}}$ to $0.5$:
+$$P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\textbf{p}^{(1)}}=[0.5, 0.5,...,0.5]^T, M_1) = \prod_{n=1}^{N} \prod_{d=1}^D (0.5)$$
+$$P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\textbf{p}^{(1)}}=[0.5, 0.5,...,0.5]^T, M_1) = 0.5^{N\cdot D}$$
 When all D components are generated from Bernoulli distributions with unknown, but identical, $p_d$, we have the likelihood function for model $M_2$:
-$$P(\textbf{x}^{(n)}|\textbf{p}^{(2)}}=[p_d, p_d, ..., p_d]^T, M_2) = \prod_{n=1}^{N}\prod_{d'=1}^{D} p_d^{x_{d'}^{(n)}} (1-p_d)^{1-x_{d'}^{(n)}}$$
+$$P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\textbf{p}^{(2)}}=[p_d, p_d, ..., p_d]^T, M_2) = \prod_{n=1}^{N}\prod_{d'=1}^{D} p_d^{x_{d'}^{(n)}} (1-p_d)^{1-x_{d'}^{(n)}}$$
 When each component is Bernoulli distributed with separate, unknown $p_d$, we have the likelihood function for model $M_3$:
-$$P(\textbf{x}^{(n)}|\textbf{p}^{(3)}}=[p_1, p_2, ..., p_D]^T, M_3) = \prod_{n=1}^{N}\prod_{d=1}^{D} p_d^{x_d^{(n)}} (1-p_d)^{1-x_d^{(n)}}$$
+$$P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\textbf{p}^{(3)}}=[p_1, p_2, ..., p_D]^T, M_3) = \prod_{n=1}^{N}\prod_{d=1}^{D} p_d^{x_d^{(n)}} (1-p_d)^{1-x_d^{(n)}}$$
 For each model $M_i$, we can marginalise out $\textbf{p}^{(i)}}$ to get $P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_i)$:
-$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_i) = \int_0^1...\int_0^1 P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|p_d, M_i)P(p_d|M_i) dp_1...dp_D$$
-where $d=1,...,D$ and $ \{\textbf{x}^{(n)}\}_{n=1}^{N}$ is our data set.
-
-Given that the prior of any unknown probabilities is uniform, i.e. $P(p_d | M_i) = 1$. We can simplify:
-$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_i) = \int_0^1...\int_0^1 P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|p_d, M_i) dp_1...dp_D$$
-For $M_1$, we have that all pixels have probability 0.5:
-$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_1) = \int_0^1...\int_0^1 \prod_{n=1}^{N} \prod_{d=1}^{D} (0.5)^{x_d^{(n)}} (1-0.5)^{1-{x_d}^{(n)}} d\theta_1...d\theta_D$$
-We can remove the integrals and knowing that either $x_d^{(n)}$ or $1-x_d^{(n)}$ will be 1 and the other zero, we can simplify $(0.5)^{x_d^{(n)}} (1-0.5)^{1-{x_d}^{(n)}}$ to $0.5$:
-$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_1) = \prod_{n=1}^{N} \prod_{d=1}^D (0.5)$$
+$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_i) = \int_0^1...\int_0^1 P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|\textbf{p}^{(i)}, M_i)P(\textbf{p}^{(i)}|M_i) dp_1...dp_D$$
+Given that the prior of any unknown probabilities is uniform, i.e. $P(\textbf{p}^{(i)} | M_i) = 1$. We can simplify:
+$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_i) = \int_0^1...\int_0^1 P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\textbf{p}^{(i)}, M_i) dp_1...dp_D$$
+For $M_1$:
+$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_1) = \int_0^1...\int_0^1 0.5^{N\cdot D} d\theta_1...d\theta_D$$
+We can remove the integrals:
 $$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_1) = (0.5)^{N\cdot D}$$
@@
 
-This integral is the beta function:
-$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_2) = \frac{K! (N\cdot D-k)!}{(N\cdot D+1)!}$$
+\item[]This integral is the beta function:
+$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_2) = \frac{K! (N\cdot D-K)!}{(N\cdot D+1)!}$$
 For $M_3$, we need an integral for each $p_d$:
@@
 Changing the products to sums:
-$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_3) = \prod_{d=1}^D \Bigg( \int_0^1   p_d^{\sum_{n=1}^{N} x_d^{(n)}} (1- p_d)^{\sum_{n=1}^{N} 1-{x_d}^{(j)}} d p_d \Bigg)$$
+$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_3) = \prod_{d=1}^D \Bigg( \int_0^1   p_d^{\sum_{n=1}^{N} x_d^{(n)}} (1- p_d)^{\sum_{n=1}^{N} 1-{x_d}^{(n)}} d p_d \Bigg)$$
 In this case, we have the product of integrals where each evaluates to a beta function:
@@
 
-The posterior probability of a model $M_i$ can be expressed:
+\item[]The posterior probability of a model $M_i$ can be expressed:
 $$P(M_i |  \{\textbf{x}^{(n)}\}_{n=1}^{N}) = \frac{P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_i)P(M_i)}{P( \{\textbf{x}^{(n)}\}_{n=1}^{N})}$$
@@
 $$P(M_i| \{\textbf{x}^{(n)}\}_{n=1}^{N}) = \frac{P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_i)}{\sum_{i\in \{1,2,3\}} P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_i)}$$
-Calculating the posterior probabilities of each of the three models having generated the data in binarydigits.txt using python, we can show the values in the Table $\ref{table2}$:
+\newpage
+Calculating the posterior probabilities of each of the three models having generated the data in binarydigits.txt using Python, we can show the values in the Table $\ref{table2}$.
 
@@
 
-We can see that for models specified to have the same parameter value for all pixels like $M_1$ is very unlikely with the given data set.
+We can see that for models specified to have the same parameter value for all pixels, like $M_1$, is very unlikely with the given data set.
 This makes sense because it is specifying models where the image is essentially blank (a uniform shade), which is not reflective of our digit images.
 Moreover, $M_1$ specifies a specific value of 0.5 for all the parameters whereas $M_2$ specifies any value for all the parameters as long as it's the same.
-So the model $M_1$ is a subset of the models specified in $M_2$ and we can see this reflected in our probabilities when $P(M_2|\{\textbf{x}^{(n)}\}_{n=1}^{N}) > P(M_1|\{\textbf{x}^{(n)}\}_{n=1}^{N})$.
+So the model $M_1$ is just one possible model specified in $M_2$ and we can see this reflected in our probabilities when $P(M_2|\{\textbf{x}^{(n)}\}_{n=1}^{N}) > P(M_1|\{\textbf{x}^{(n)}\}_{n=1}^{N})$.
 
@@
 \item[(a)] The likelihood for a model consisting of a mixture of K multivariate Bernoulli distributions can be expressed as the product across $N$ data points:
-$$P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\theta) = \prod_{i=1}^{N}P(x_i|\theta)$$
-where $\{\textbf{x}^{(n)}\}_{n=1}^{N}$ is our data set with $\textbf{x}^{(n)} \in \mathbb{R}^{D \times 1}$ and $\theta = \{\pi, \mathbf{P}\}$, $\pi = [\pi_1, .
+$$P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\theta) = \prod_{i=1}^{N}P(\textbf{x}^{(n)}|\theta)$$
+where $\{\textbf{x}^{(n)}\}_{n=1}^{N}$ is our data set with $\textbf{x}^{(n)} \in \mathbb{R}^{D \times 1}$ and $\theta = \{\pi, \mathbf{P}\}$ are our parameters, $\pi = [\pi_1, .
 .
 .
-, \pi_K] \in \mathbb{R}^{K \times 1}$ our mixing proportions $(0 \leq \pi_k \leq 1; \sum_k \pi_k = 1)$ and $\mathbf{P} \in \mathbb{R}^{D \times K}$ the K Bernoulli parameter vectors with elements $p_{kd}$ denoting the probability that pixel d takes value 1 under mixture component k.
+, \pi_K] \in \mathbb{R}^{K \times 1}$ our K mixing proportions $(0 \leq \pi_k \leq 1; \sum_k \pi_k = 1)$ and $\mathbf{P} \in \mathbb{R}^{D \times K}$ the K Bernoulli parameter vectors with elements $p_{kd}$ denoting the probability that pixel d takes value 1 given mixture component k.
 We also assume the images are iid and that the pixels are independent of each other within each component distribution.
@@
 For each $P(\textbf{x}^{(n)}|\theta)$:
-$$P(\textbf{x}^{(n)}|\theta) = \sum_{k=1}^{K} \pi_k \prod_{d=1}^D (p_{kd})^{\textbf{x}^{(n)}_{d}} (1-p_{kd})^{1-\textbf{x}^{(n)}_{d}}$$
-The log-likelihood $\mathcal{L}(\textbf{x}^{(n)}|\theta)$ can be expressed in matrix form:
+$$P(\textbf{x}^{(n)}|\theta) = \sum_{k=1}^{K} \pi_k \prod_{d=1}^D (p_{kd})^{x^{(n)}_{d}} (1-p_{kd})^{1-x^{(n)}_{d}}$$
+The log-likelihood $\mathcal{L}(\textbf{x}^{(n)}|\theta)$ can be expressed in vector form:
 $$\mathcal{L}(\textbf{x}^{(n)}|\theta) = \log \sum_{k=1}^{K}  \pi_k \exp\bigg(\textbf{x}^{(n)}\log(\mathbf{P}_{k}) + (1-\textbf{x}^{(n)})\log(1-\mathbf{P}_{k})\bigg) $$
@@
 where $s^{(n)} \in \{1,...,K\}$ a discrete hidden variable with $P(s^{(n)} = k|\textbf{x}^{(n)}, \pi) = \pi_k$.
-Note that $P(s^{(n)} = k|\textbf{x}^{(n)} , \pi) = P(s^{(n)} = k|\textbf{x}^{(n)} , \pi, \mathbf{P})$ as $s^{(n)$ isn't dependent on $\mathbf{P}$.
+Note that $P(s^{(n)} = k|\pi, \mathbf{P}) = P(s^{(n)} = k| \pi)$ as $s^{(n)$ isn't dependent on $\mathbf{P}$.
 
-Let $P(s^{(n)} = k|\textbf{x}^{(n)}, \pi, \mathbf{P}) \propto P(s^{(n)}$ be the unnormalised responsibility $\tilde{r}_{nk}$.
+Let $\tilde{r}_{nk}$ be the unnormalised responsibility $ P(\textbf{x}^{(n)} | s^{(n)} = k, \pi, \mathbf{P}) P(s^{(n)} = k|\pi, \mathbf{P})$.
 Using the mixture for component $k$, $\pi_k$ and the likelihood function of component $k$:
-$$\tilde{r}_{nk} = \pi_k \prod_{d=1}^D (p_{kd})^{\textbf{x}^{(n)}_{d}} (1-p_{kd})^{1-\textbf{x}^{(n)}_{d}}$$
+$$\tilde{r}_{nk} = \pi_k \prod_{d=1}^D (p_{kd})^{x^{(n)}_{d}} (1-p_{kd})^{1-x^{(n)}_{d}}$$
 Normalising across the components:
@@
 Moreover,
-$$ \log \tilde{r}_{nk} = \log \pi_k + \sum_{d=1}^D \bigg( \textbf{x}^{(n)}_{d}\log(p_{kd}) + (1-\textbf{x}^{(n)}_{d}) \log(1-\exp(\log(p_{kd}))) \bigg)$$
+$$ \log \tilde{r}_{nk} = \log \pi_k + \sum_{d=1}^D \bigg( x^{(n)}_{d}\log(p_{kd}) + (1-x^{(n)}_{d}) \log(1-\exp(\log(p_{kd}))) \bigg)$$
 and
 $$\log r_{nk} = \log\tilde{r}_{nk} - \log\sum_{j=1}^K\exp(\log\tilde{r}_{nj})$$
-which can be vectorised as $\log \textbf{r}_{n}$ calculated with $\log\pi$ and $\log \textbf{P}$ using Python scipy's $logsumexp$ operation.
+which can be vectorised as $\log \textbf{r}$ calculated with $\log\pi$ and $\log \textbf{P}$ using Python scipy's $logsumexp$ operation.
 
@@
 $$\Bigg\langle \sum_n \log P(\textbf{x}^{(n)}, s^{(n)}| \pi, \mathbf{P})\Bigg\rangle_{q(\{s^{(n)}\})} = \sum_{n=1}^N q(s^{(n)}) \log P(\textbf{x}^{(n)}, s^{(n)}| \pi, \mathbf{P})$$
-Let this quantity be $E$. Each term of $E$ can be expressed:
-$$q(s^{(n)}) = \textbf{r}_{n}$$
+Let this quantity be $E$. For each term of $E$:
+$$q(s^{(n)}) = \textbf{r}_{n}^T$$
 and
@@
 $$ E = \sum_n \textbf{r}_{n}^T[\log(\pi)+\log(\mathbf{P})^T \textbf{x}^{(n)} + \log(1-\mathbf{P})^T(1-\textbf{x}^{(n)})]$$
-To maximise with respect to $\pi$ and $\mathbf{P}$ for the M step, we want to take the derivative, set to zero, and solve for $\hat\pi$ and $\hat P$.
+To maximise with respect to $\pi$ and $\mathbf{P}$ for the M step, we want to take the derivative, set to zero, and solve for $\hat\pi$ and $\hat \textbf{P}$.
 
@@
 $$\frac{\partial E}{\partial \pi_k} = \sum_n r_{nk} \frac{1}{\pi_k}$$
-The second derivative:
-$$\frac{\partial E}{(\partial \pi_k)^2} = \sum_n r_{nk} \frac{-1}{(\pi_k)^2}$$
-is always negative because $r_{nk}\geq 0$, $\sum_{n}r_{nk}=1$, \pi_{k}\geq 0$, and $\sum_{n}\pi_{k}=1$, ensuring a maximum in the next step.
+%The second derivative:
+%$$\frac{\partial E}{(\partial \pi_k)^2} = \sum_n r_{nk} \frac{-1}{(\pi_k)^2}$$
+%is always negative because $r_{nk}\geq 0$, $\sum_{n}r_{nk}=1$, \pi_{k}\geq 0$, and $\sum_{n}\pi_{k}=1$, ensuring a maximum in the next step.
 
@@
 For the $dk^{th}$ element of $\mathbf{P}$:
-$$\frac{\partial E}{\partial \mathbf{P}_{dk}} = \sum_n r_{nk}\frac{\partial}{\partial \mathbf{P}_{dk}}[\textbf{x}^{(n)}_{d}\log \mathbf{P}_{dk}+(1-\textbf{x}^{(n)}_{d}) \log(1-\mathbf{P}_{dk})]$$
+$$\frac{\partial E}{\partial \mathbf{P}_{dk}} = \sum_n r_{nk}\frac{\partial}{\partial \mathbf{P}_{dk}}[x^{(n)}_{d}\log \mathbf{P}_{dk}+(1-x^{(n)}_{d}) \log(1-\mathbf{P}_{dk})]$$
 Simplifying:
-$$\frac{\partial E}{\partial \mathbf{P}_{dk}} = \sum_n r_{nk}(\frac{\textbf{x}^{(n)}_{d}}{ \mathbf{P}_{dk}}-\frac{1-\textbf{x}^{(n)}_{d}}{1-\mathbf{P}_{dk}})$$
-Similar to Question 1, we can see that taking second derivative, the term in the brackets will always be less than zero and with $r_{nk}\geq 0$ and $\sum_{n}r_{nk}=1$, the second derivative will always be negative.
+$$\frac{\partial E}{\partial \mathbf{P}_{dk}} = \sum_n r_{nk}(\frac{x^{(n)}_{d}}{ \mathbf{P}_{dk}}-\frac{1-x^{(n)}_{d}}{1-\mathbf{P}_{dk}})$$
+%Similar to Question 1, we can see that taking second derivative, the term in the brackets will always be less than zero and with $r_{nk}\geq 0$ and $\sum_{n}r_{nk}=1$, the second derivative will always be negative.
 This ensures that we have a maximum in the next step.
@@
 Setting the derivative to zero:
-$$\frac{\sum_n  \textbf{x}^{(n)}_{d}r_{nk}}{ \mathbf{P}_{dk}}-\frac{\sum_n r_{nk}-\sum_n  \textbf{x}^{(n)}_{d}r_{nk}}{1-\mathbf{P}_{dk}} = 0$$
+$$\frac{\sum_n  x^{(n)}_{d}r_{nk}}{\hat{\mathbf{P}}_{dk}}-\frac{\sum_n r_{nk}-\sum_n  x^{(n)}_{d}r_{nk}}{1-\hat{\mathbf{P}}_{dk}} = 0$$
 Solving for $\hat{\mathbf{P}}_{dk}$:
-$$\hat{\mathbf{P}}_{dk}\sum_n r_{nk}-\hat{\mathbf{P}}_{dk}\sum_n \textbf{x}^{(n)}_{d}r_{nk} = \sum_n  \textbf{x}^{(n)}_{d}r_{nk}-\hat{\mathbf{P}}_{dk}\sum_n  \textbf{x}^{(n)}_{d}r_{nk}$$
+$$\hat{\mathbf{P}}_{dk}\sum_n r_{nk}-\hat{\mathbf{P}}_{dk}\sum_n x^{(n)}_{d}r_{nk} = \sum_n  x^{(n)}_{d}r_{nk}-\hat{\mathbf{P}}_{dk}\sum_n  x^{(n)}_{d}r_{nk}$$
 Thus,
-$$\hat{\mathbf{P}}_{dk} = \frac{\sum_n \textbf{x}^{(n)}_{d}r_{nk}}{\sum_n r_{nk}}$$
+$$\hat{\mathbf{P}}_{dk} = \frac{\sum_n x^{(n)}_{d}r_{nk}}{\sum_n r_{nk}}$$
 We have the maximizing parameters for the expected log-joint
@@
 thus obtaining an iterative update for the parameters $\pi$ and $\mathbf{P}$ in the M-step of EM.
-For numerical stability, we can compute the maximisation step for the MAP of $\mathbf{P}$,$\hat{\mathbf{P}}_{dk}^{MAP}$ by solving:
+
+\item[]For numerical stability, we can compute the maximisation step for the MAP of $\mathbf{P}$, by solving for $\hat{\mathbf{P}}_{dk}^{MAP}$ with:
 $$\frac{\partial E'}{\partial \mathbf{P}_{dk}} = 0$$
@@
 Substituting the appropriate expressions:
-$$\frac{\partial E'}{\partial \mathbf{P}_{dk}} =  \sum_n \Bigg(r_{nk} \bigg(\frac{\textbf{x}^{(n)}_{d}}{ \mathbf{P}_{dk}}-\frac{1-\textbf{x}^{(n)}_{d}}{1-\mathbf{P}_{dk}} + \frac{(\alpha-1)}{\mathbf{P}_{dk}} -  \frac{(\beta-1)}{1-\mathbf{P}_{dk}} \bigg) \Bigg)$$
+$$\frac{\partial E'}{\partial \mathbf{P}_{dk}} =  \sum_n \Bigg(r_{nk} \bigg(\frac{x^{(n)}_{d}}{ \mathbf{P}_{dk}}-\frac{1-x^{(n)}_{d}}{1-\mathbf{P}_{dk}} + \frac{(\alpha-1)}{\mathbf{P}_{dk}} -  \frac{(\beta-1)}{1-\mathbf{P}_{dk}} \bigg) \Bigg)$$
 Simplifying:
-$$\frac{\partial E'}{\partial \mathbf{P}_{dk}} =  \frac{\sum_n r_{nk}(\alpha-1 + \textbf{x}^{(n)}_{d})}{ \mathbf{P}_{dk}}-\frac{\sum_n r_{nk}(\beta-\textbf{x}^{(n)}_{d})}{1-\mathbf{P}_{dk}}$$
-For a maximum, we see that we need $\alpha > \textbf{x}^{(n)}_{d}-1$  or $\beta < \textbf{x}^{(n)}_{d}$, both of which are satisfied knowing that $\alpha > 0$ and $\beta > 0$.
+$$\frac{\partial E'}{\partial \mathbf{P}_{dk}} =  \frac{\sum_n r_{nk}(\alpha-1 + x^{(n)}_{d})}{ \mathbf{P}_{dk}}-\frac{\sum_n r_{nk}(\beta-x^{(n)}_{d})}{1-\mathbf{P}_{dk}}$$
+%For a maximum, we see that we need $\alpha > x^{(n)}_{d}-1$  or $\beta < x^{(n)}_{d}$, both of which are satisfied knowing that $\alpha > 0$ and $\beta > 0$.
 Setting $\frac{\partial E'}{\partial \mathbf{P}_{dk}}=0$ we can calculate $\hat{\mathbf{P}}_{dk}^{MAP}$:
-$$\sum_n r_{nk}(\alpha-1 + \textbf{x}^{(n)}_{d})} - \hat{\mathbf{P}}_{dk}\sum_n r_{nk}(\alpha-1 + \textbf{x}^{(n)}_{d})} = \hat{\mathbf{P}}_{dk}\sum_n r_{nk}(\beta-\textbf{x}^{(n)}_{d})$$
-$$ \hat{\mathbf{P}}_{dk}^{MAP} = \frac{\sum_n r_{nk}(\textbf{x}^{(n)}_{d}+\alpha-1)}{(\alpha+\beta-1)( \sum_n r_{nk})}$$
+$$\sum_n r_{nk}(\alpha-1 + x^{(n)}_{d})} - \hat{\mathbf{P}}_{dk}\sum_n r_{nk}(\alpha-1 + x^{(n)}_{d})} = \hat{\mathbf{P}}_{dk}\sum_n r_{nk}(\beta-x^{(n)}_{d})$$
+$$ \hat{\mathbf{P}}_{dk}^{MAP} = \frac{\sum_n r_{nk}(x^{(n)}_{d}+\alpha-1)}{(\alpha+\beta-1)( \sum_n r_{nk})}$$
 As a sense check, we can see when setting $\alpha=1$ and $\beta=1$ we recover $ \hat{\mathbf{P}}_{dk}^{MLE}$ as we would expect.
 \newpage
-\item[(d)] Plotting the posterior likelihood as a function of the iteration number:
+\item[(d)] Plotting the posterior likelihood as a function of the iteration number for different k values:
 
@@
 
-where $epsilon$ is the stopping condition for the posterior posterior converges.
+where $epsilon$ is the stopping condition for when the log posterior converges sufficiently.
 
@@
 
-Displaying the parameters found for $K$ in $\{2, 3, 4, 7, 10\}$:
+Displaying the parameters found for $K \in \{2, 3, 4, 7, 10\}$:
 \begin{figure}[h]
@@
 
-For smaller k, we can visually see that we obtain very similar solutions (a 7 and a 0 for $k=2$).
+For smaller k, we can visually see that we obtain very similar solutions (a seven and a zero for $k=2$).
+For $k=3$, we get one each of zero, seven, and five, but in different orderings for different trials.
 However for higher K, we see that this may not always be the case.
-For Trial 2 of $k=10$, we have three 5's whereas in Trial 4 we have two 5's.
+For Trial 1 of $k=10$, we have three 5's whereas in Trial 3 we have four 5's.
 Interestingly, different clusters of the same digits can be different, representing different variants of the written digit (i.e.
@@
 However for $k=7$ and especially $k=10$, we can see some components have very small or zero probability (i.e.
-$\pi_2$ of trial 2).
+$\pi_3$ of trial 2).
 It will be unlikely for those components to represent very distinct clusters (i.e.
-the parameters for $\pi_2$ and $\pi_9$ are very similar in trial 2) This can be verified when we perform a TSNE visualisation of the responsibility vector for each of the images (Note that for $k=2$, the responsibility vector is displayed).
-We can see that for large k, qualitatively the number of clusters no longer matches the k value, indicating that some clusters are redundant.
-For example for $k=7$ and $k=10$ we can only qualitatively see four or five clusters with TSNE.
+the parameters for $\pi_2$ and $\pi_3$ are very similar in trial 2) This can be verified when we perform a TSNE visualisation of the responsibility vector for each of the images (Note that for $k=2$, just the responsibility vector is plotted because it is two dimensional).
+We can see that for large k, qualitatively the number of clusters no longer matches the k value, indicating that some mixtures are redundant.
+For example for $k=7$ and $k=10$ we can only qualitatively see three to five clusters with TSNE.
 
@@
 
-Improvements to the model could include searching for an optimal $k$ by maximising the log posterior with regularisation on the magnitude of $k$ to balancing maximising log posterior with minimising model complexity.
-Additionally, adding a prior on the responsibility components can be helpful to ensure non-zero mixing components unlike the components visualised here.
+Improvements to the model could include searching for an optimal $k$ by maximising the log posterior with regularisation on the magnitude of $k$ to balance maximising log posterior with minimising model complexity.
+Additionally, adding a prior on the responsibility components can be helpful to ensure more even mixture components unlike the components visualised here.
 This could help promote more meaningful clusters as $k$ increases.
@@
 \end{enumerate}
-
-[BONUS] Express the log-likelihoods obtained in bits and relate these numbers to the length
-of the naive encoding of these binary data. How does your number compare to gzip (or
-another compression algorithm)? Why the difference? [5 marks]
-
- [BONUS] Consider the total cost of encoding both the model parameters and the data given
-the model. How does this total cost compare to gzip (or similar)? How does it depend on
-K ? What might this tell you? [5 marks]
 
@@
 \item[(a)] The formulae for the ML estimates of $P(s_i = \alpha |s_{i-1} = \beta) = \Psi(\alpha, \beta)$:
-$$\Psi(\alpha, \beta) = \frac{N_{s_i, s_{i-1}}}{N_{s_{i-1}}}$$
-where $N_{s_i, s_{i-1}}$ is the count of the number of occurrences of the pair $(s_i, s_{i-1})$, where $s__{i-1}$ is followed by $s_{i}$ and $N_{s__{i-1}}$ is the number of occurrences of $s_{i-1}$.
+$$\Psi(\alpha, \beta) = \frac{N_{\alpha, \beta}}{N_{\beta}}$$
+where $N_{\alpha, \beta}$ is the count of the number of occurrences of the pair $(\alpha, \beta)$, where $\beta$ is followed by $s\alpha$ in the text and $N_{\beta}$ is the number of occurrences of $\beta$.
+Moreover to ensure ergodicity, a one was added to each $N_{\alpha, \beta}$.
+This was also taken into account for the normaliser $N_{\beta}$.
 
@@
 \begin{enumerate}
-  \item[(i)] Initialise any $\phi_0 \in \mathbb{R}^{53 \times 1}$
-  \item[(ii)] Repeat $\phi_{i+1} = \Psi \phi_{i}$
-  \item[(iii)] Terminate when $\phi_{i+1} - \phi_{i} < \epsilon$
+  \item[(i)] Initialise any $\phi^{(0)} \in \mathbb{R}^{53 \times 1}$ and $\sum_i \phi^{(0)}_i = 1$
+  \item[(ii)] Repeat $\phi^{(i+1)} = \Psi \phi^{(i)}$
+  \item[(iii)] Terminate when $\phi^{(i+1) - \phi^{(i)} < \epsilon$
 \end{enumerate}
 
-where $\Psi \in \mathbf{R}^{53\times53}$ containing the transition probabilities, $\Psi_{i, j} = P(\alpha_j | \alpha_i)$ where $\alpha_i$ is the $i^{th}$ symbol and $\alpha_j$ is the $j^{th}$ symbol, and $\epsilon$ is some small number indicating sufficient convergence of the distribution to be considered stationary.
-The function $\phi(\gamma)$ is simply the index of $\gamma$ in the vector $\phi$.
+where $\Psi \in \mathbf{R}^{53\times53}$ containing the transition probabilities, $\Psi_{i, j} = P(s_j | s_i)$ where $s_i$ is the $i^{th}$ symbol and $s_j$ is the $j^{th}$ symbol, and $\epsilon$ is some small number indicating sufficient convergence of the distribution to be considered stationary.
+The function $\phi(\gamma)$ is simply the index of symbol $\gamma$ in the vector $\phi$.
 
@@
 We have 53 symbols but only 52 degrees of freedom, because once we have defined the encoding for 52 symbols, the encoding for the $53^{rd}$ symbol cannot be chosen.
-Thus, there exists a dependence between the symbols for a given $\sigma$.
+Thus, there exists a dependence between $\sigma(s)$ for different symbols $s$.
 
@@
 %\begin {table}[h]
-\scalebox{0.4} {\begin{tabular}{c|l}%
+\scalebox{0.4} {\begin{tabular}{c|c}%
  \bfseries \textbf{MH Iteration} & \textbf{Current Decryption}% specify table head
@@
 
-To help with chain initialisation, 10000 different $\sigma$'s were randomly and independently sampled.
-The $\sigma$ providing the best log-likelihood was chosen as the starting point for the MH chain and algorithm was then run for 10000 iterations.
-Moreover, ten different trials were performed, where the trial with the best log-likelihood is displayed.
+\newpage
+To help with chain initialisation, 10000 different $\sigma$'s were first randomly and independently sampled.
+The $\sigma$ with the best log-likelihood was chosen as the starting point for the MH chain and the algorithm was then run for 10000 iterations.
+Moreover, ten different trials of this was performed, where the trial with the best log-likelihood was displayed.
+The decrypted message for each of the ten trials:
+
+\begin{center}
+%\begin {table}[h]
+\scalebox{0.4} {\begin{tabular}{c|c}%
+ \bfseries \textbf{Trial} & \textbf{Decryption}% specify table head
+\csvreader[head to column names,
+/csv/separator=pipe,
+]{outputs/q5/q5d-trials.csv}{}% use head of csv as column names
+{\\\hline\csvcoli&\csvcolii}% specify your coloumns here
+\end{tabular}
+}
+%\end{table}
+\end{center}
 
@@
 An ergodic chain is one that is irreducible (i.e.
-all possible transitions between symbols have probability greater than zero).
+all possible transitions between symbols, including to itself, have probability greater than zero).
 If $\Psi(\alpha, \beta) = 0$, this means that there is zero probability that $\beta$ will transition to $\alpha$, breaking our definition.
@@
 
-\item[(f)] Analyse this approach to decoding.
-For instance, would symbol probabilities alone (rather than transitions) be sufficient?
-If we used a second order Markov chain for English text, what problems might we encounter?
-Will it work if the encryption scheme allows two symbols to be mapped to the same encrypted value?
-Would it work for Chinese with $>$ 10000 symbols?
-[13 marks]
-
-If we were to use symbol probabilities alone for decoding, the joint probability would be:
+\item[(f)] If we were to use symbol probabilities alone for decoding, the joint probability would be:
 $$P(e_1, e_2,...,e_n|\sigma) = \prod_{i=1}^n P(\sigma^{-1}(e_i))$$
@@
 In this case, the optimal decoding would simply replace the most frequent symbols in the encrypted message with the most frequent symbols in the training text.
-This is much more difficult because each letter is assumed to be independent of its neighbours.
-For a first order Markov chairn, we exploit the structure of language by considering pairs of letters.
+This decoding approach is much more difficult because each letter is assumed to be independent of its neighbours.
+For a first order Markov chain, we exploit the structure of language by considering pairs of letters.
 Assuming that as the training text size approaches infinity and the size of the encrypted message also approaches infinity, that the two will have the same symbol frequency and that the probability of each symbol is unique, (i.e.
-two different decodings can't have the same likelihood), then using symbol probabilities alone should theoretically work.
-However, in practise we would unlikely to be able to make these assumptions about symbol frequencies from the size of our training set and encrypted message.
+two different symbols can't have the same frequency), then using symbol probabilities alone should theoretically work by matching symbol probabilities.
+However, in practise it would be unlikely to be able to make these assumptions about symbol frequencies, especially with the finite size of our training set and encrypted message.
 
 A second-order chain should also work in theory.
-However, this approach is probably practically  more difficult for finding a suitable decoding.
-This is because our transition matrix would contain $N^3$, where $N$ is the number of symbols, to account for all possible second order transitions.
-Our training text would need to increase quadratically to maintain the same ratio of possible transitions to example transitions (number of second order transitions in a text of length $N$ is $N-2$ and third order its $N-3$).
+However, with this approach it is probably practically more difficult for finding a suitable decoding.
+This is because our transition tensor would contain $N^3$ elements, where $N$ is the number of symbols, to account for all possible second order transitions.
+Our training text would need to increase quadratically to maintain the same ratio of possible transitions to example transitions (number of first order transitions in a text of length $N$ is $N-1$ and second order its $N-2$).
+This can also introduce sparsity (as in small non-zero probabilities in many entries) in our transition matrix despite our ability to maintain ergodicity with small probabilities to prevent non-zero entries.
+However, the log-likelihood of many areas of $\sigma$ space might be very small or all just the same, when all the transition probabilities are just the offset probability (added to maintain ergodicity).
+Navigating this space will be much more difficult for the sampler as the space could be relatively flat.
 
@@
 this approach can become much more complicated.
-Our $\sigma^{-1}(e)$ is ill-defined, and therefore how we computing the joint probability of the encrypted text is no longer immediately clear.
+Our $\sigma$ is no longer as easily inverted and therefore for each duplicate mapping, we would have to integrate out the probability for the two possible decrypted symbols when computing the log-likelihood.
 Moreover, generating proposal encodings is not as simple as swapping the encryption for two symbols.
 This is because we do not know which two symbols map to the same encrypted symbol and simply swapping would preserve the same collision mapping of the current encoding.
-Overall, many changes would need to be made to the approach to accommodate for these complications.
-It is not immediately obvious how current approach could work for this case.
+Moreover, the number of proposal $\sigma'$'s will depend on how many duplicates exist in the current $\sigma$.
+Thus $S(\sigma \rightarrow \sigma')$ would no longer be symmetric, complicating the acceptance probability calculation as it would be dependent on the $\sigma$ and $\sigma'$.
+Overall, this approach could work but would require many changes to accommodate for these complications.
+Integrating out collision mappings in the log-likelihood, non-symmetric proposal probabilities, and a much larger $\sigma$ space because duplicates are allowed, means that it will take much longer for the sampler to find a reasonable $\sigma$.
+
+%If we start at a unique mapping, there are ${53\choose2} + 52(53-M)$ proposal encodings that involve a single change, where ${53\choose2}$ are the number of possible swaps and $53(53-M)$ is the number of proposals for which we add one collision mapping, where $M$ is the number of unique symbols in the encrypted message.
+$52(53-M)$ where we can move the mapping to any one of $53-M$ encoded symbols $e$ that don't show up in the encrypted message and collide it with any one of the other $52$ other encoded symbols.
+If we start at a mapping with duplicates, there are $({53\choose2} - \tilde{C}) + \mathbb{I}(M<53-C)(52-2C)(53-M-C) + 52C$ where $\tilde{C}$ is the current number of duplicates in $\sigma$, $C$ is the current number of duplicates to encoded symbols seen in the encrypted text (duplicates to unseen encoded symbols are like placeholders in this case), and $\mathbb{I}$ is an indicator function.
+$({53}\choose{2} - \tilde{C})$ is the number of swaps possible, where we remove the swapping of collision mappings.
+$\mathbb{I}(M<53-C)(52-2C)(53-C)$ is the number of proposals where we can add another duplicate mapping.
+$\mathbb{I}(M<53-C)$ indicates whether more adding another duplicate is possible, $52-2C$ is number of symbols that aren't currently colliding to an encoded symbol seen in the encoded message and can be remapped to collide, and
 
 If we used this approach for Chinese with $\geq 10000$ symbols, we would be attempting to solve the same problem but with $N\geq10000$ instead of $N=53$.
-Similar to the second order Markov chain, although this is theoretically possible, it would require a transition matrix of size $\geq10000^2$ which is quite impractical.
-An alternative set up  could be with using Chinese phonetics, for which there are likely much fewer than $10000$, however this would require a mapping from a phonetic to an encrypted phonetic.
+Similar to the second order Markov chain, although this is theoretically possible, it would require a transition matrix of size $\geq10000^2$ which is quite impractical and we'd run into similar problems as for second order Markov Chains.
+An alternative set up  could be with using Chinese phonetics, for which there are much fewer than $10000$, however this would require a mapping from a phonetic to an encrypted phonetic.
 
@@
 $$\min_{\textbf{x}} f(\textbf{x})$$
-such that $g(\textbf{x})=\textbf{0}$ and $\textbf{x} := [x, y]^T$.
+such that $g(\textbf{x})=0$ and $\textbf{x} := [x, y]^T$.
 
@@
 Substituting $ y = \frac{-1}{\lambda}$ from the first equation into the second equation:
-$$2+\frac{-1}{\lambda}(2y+x) = 0$$
-$$\frac{-x}{y} = 0$$
-We see that $x=0$.
+$$2+\lambda(2(\frac{-1}{\lambda}) + x) = 0$$
+$$x = 0$$
 Solving for $y$ in our third equation with $x=0$:
@@
 
-The local extrema are $(x=0, y=1)$ when our $\lambda=-1$ and $(x=0, y=-1)$ when our $\lambda=1$.
+The local extrema are $(x=0, y=1)$ when $\lambda=-1$ and $(x=0, y=-1)$ when $\lambda=1$.
 
@@
 \begin{enumerate}
-             \item[(i)] Given that $g(a) = \ln(a)$, we want to transform this to the form $f(x,a)=0$:
+             \item[(i)] Given that $g(a) = \ln(a)$, we want to transform this to the form $f(x,a)=0$ where $x=g(a)$:
              $$x = \ln(a)$$
@@
              $$f'(x) = \frac{\partial f(x, a)}{\partial x}= \exp(x)$$
-             Assuming we can evaluate $\exp(x)$, our update equation:
+             Assuming we can evaluate $\exp(x)$, our update equation is:
              $$x_{n+1} = x_{n} - \frac{\exp(x_n) - a}{\exp(x_n)}$$
@@
              $$x_{n+1} = x_{n} + \frac{a}{\exp(x_n)} - 1$$
+             we have our update equation in Newton's algorithm for this problem.
 \end{enumerate}
+\end{enumerate}
+
+\newpage
+\section*{Question 8}
+\begin{enumerate}
+\item[(a)] For:
+$$\sup_{\{\textbf{x}\in \mathbb{R}^n\}} R_A(\textbf{x})$$
+where $R_A(\textbf{x}) = \frac{\textbf{x}^T\textbf{Ax}}{\|\textbf{x}\|^2}$, we want to show that a maximum is attained.
+
+To do this, we will first show that the above optimisation can be equivalently formulated as:
+$$\sup_{\big\{\textbf{x}\in \mathbb{R}^n \bigm| \| \textbf{x}\|=1\big\}} R_A(\textbf{x})$$
+We begin by considering any $\textbf{w} \in \mathbb{R}^n$ and let $\textbf{x} = \frac{\textbf{w}}{\|\textbf{w}\|}$. Because $\|\textbf{x}\| = 1$ we can substitute:
+$$\sup_{\big\{\textbf{x}\in \mathbb{R}^n \bigm| \| \textbf{x}\|=1\big\}} R_A(\textbf{x}) = \sup_{\bigg\{\frac{\textbf{w}}{\|\textbf{w}\|}\in \mathbb{R}^n \biggm| \| \frac{\textbf{w}}{\|\textbf{w}\|}\|=1\bigg\}} \frac{\textbf{w}^T\textbf{Aw} \|\textbf{w}\|^2}{\|\textbf{w}\|^2 \textbf{w}^T\textbf{w}}$$
+where $\|\textbf{x}\|^2 = \textbf{x}^T\textbf{x}$.
+
+The set $\bigg\{\frac{\textbf{w}}{\|\textbf{w}\|}\in \mathbb{R}^n \biggm| \| \frac{\textbf{w}}{\|\textbf{w}\|}\|=1\bigg\}$ contains all $\textbf{w} \in \mathbb{R}^n$ so we can rewrite:
+$$\sup_{\big\{\textbf{x}\in \mathbb{R}^n \bigm| \| \textbf{x}\|=1\big\}} R_A(\textbf{x})=\sup_{\{\textbf{w}\in \mathbb{R}^n\}} \frac{\textbf{w}^T\textbf{Aw} \|\textbf{w}\|^2}{\|\textbf{w}\|^2 \textbf{w}^T\textbf{w}}$$
+We can simplify the expression:
+$$\sup_{\big\{\textbf{x}\in \mathbb{R}^n \bigm| \| \textbf{x}\|=1\big\}} R_A(\textbf{x})=\sup_{\{\textbf{w}\in \mathbb{R}^n\}} \frac{\textbf{w}^T\textbf{Aw} }{\textbf{w}^T\textbf{w}}$$
+$$\sup_{\big\{\textbf{x}\in \mathbb{R}^n \bigm| \| \textbf{x}\|=1\big\}} R_A(\textbf{x})=\sup_{\{\textbf{w}\in \mathbb{R}^n\}} R_A(\textbf{w})$$
+
+and recover our original optimisation problem by letting $\textbf{x}=\textbf{w}$, showing that it is equivalent to the supremum over the unit sphere.
+Assuming the set containing the unit sphere is compact, the extreme value theory of calculus states that $\sup_{\{\textbf{x}\in \mathbb{R}^n | \| \textbf{x}\|=1\}} R_A(\textbf{x})$ is attained so equivalently $\sup_{\{\textbf{x}\in \mathbb{R}^n\}} R_A(\textbf{x})$ is attained as required.
+
+\item[(b)] We can now reformulate the optimisation as:
+$$\sup_{\big\{\textbf{x}\in \mathbb{R}^n \bigm| \| \textbf{x}\|=1\big\}} \frac{\textbf{x}^T\textbf{Ax}}{\|\textbf{x}\|^2}$$
+Because $\| \textbf{x}\|=1$, we can equivalently write:
+$$\sup_{\big\{\textbf{x}\in \mathbb{R}^n \bigm| \| \textbf{x}\|=1\big\}} \textbf{x}^T\textbf{Ax}}$$
+Thus, showing $R_A(\textbf{x}) \leq \lambda_1$ will be equivalent to showing $\textbf{x}^T\textbf{Ax}} \leq \lambda_1$ for $\|\textbf{x}\|=1$.
+We know that for all $\textbf{x} \in \mathbb{R}^n$:
+$$\textbf{x} = \sum_{i=1}^n (\xi_i^T \textbf{x}) \xi_i$$
+so we can write:
+$$\textbf{x}^T\textbf{Ax} =\bigg(\sum_{i=1}^n (\xi_i^T \textbf{x}) \xi_i^T\bigg)\textbf{A}\bigg(\sum_{i=1}^n (\xi_i^T \textbf{x}) \xi_i\bigg)$$
+Given that $\xi_i$ are eigenvectors of $\textbf{A}$ corresponding to eigenvalues $\lambda_i$:
+$$\textbf{x}^T\textbf{Ax} =\bigg(\sum_{i=1}^n (\xi_i^T \textbf{x}) \xi_i^T\bigg)\bigg(\sum_{i=1}^n \lambda_i(\xi_i^T \textbf{x}) \xi_i\bigg)$$
+Given that the eigenvectors $\xi_i$ form an orthonormal basis, we know that $\xi_i^T\xi_j = 0$ when $i\neq j$ and $\xi_i^T\xi_j = 1$ when $i=j$, so:
+$$\textbf{x}^T\textbf{Ax} =\sum_{i=1}^n \lambda_i(\xi_i^T \textbf{x})^2$$
+From our above reformulation with the unit sphere, we know that $\|\textbf{x}\|^2 = 1$ so $\|\textbf{x}\|^2 = \sum_{j=1}^{n} x_i^2 = \sum_{j=1}^{n} (\xi_j \textbf{x})^2 = 1$.
+Thus the quantity $\sum_{i=1}^n \lambda_i(\xi_i^T \textbf{x})^2$ is a weighted average of $\lambda_i$'s, which is always less than or equal to the largest $\lambda_i$ value so:
+$$\textbf{x}^T\textbf{Ax} =\sum_{i=1}^n \lambda_i(\xi_i^T \textbf{x})^2 \leq \lambda_1$$
+where $\lambda_1$ is the largest eigenvalue of eigenvalues $\lambda_i$. Therefore, $R_A(\textbf{x}) \leq \lambda_1$ as required.
+
+\item[(c)] Given that $\textbf{x} \in span\{\xi_{k+1}, ..., \xi_{n}\}$, we can rewrite $\textbf{x}$:
+$$\textbf{x} = \sum_{i=k+1}^n (\xi_i^T \textbf{x}) \xi_i$$
+Using the same argument as in (b) we can bound $\textbf{x}^T\textbf{Ax}}$:
+$$\textbf{x}^T\textbf{Ax}} = \sum_{i=k+1}^n \lambda_i(\xi_i^T \textbf{x})^2 \leq \max\{\lambda_{k+1}, ..., \lambda_n\}$$
+But given that the maximum eigenvalue $\lambda_1$ is not contained in $\{\lambda_{k+1}, ..., \lambda_n\}$:
+$$\max\{\lambda_{k+1}, ..., \lambda_n\} < \lambda_1$$
+and therefore
+$R_A(\textbf{x}) < \lambda_1$
+as required.
 
@@
 \newpage
-\section*{Appendix: main.py}
+\section*{Appendix 1: constants.py}
+\lstinputlisting[language=Python]{src/constants.py}
+\newpage
+\section*{Appendix 2: main.py}
 \lstinputlisting[language=Python]{main.py}

### 2022-11-13 main.py

@@ -45,6 +45,8 @@ if __name__ == "__main__":
         os.makedirs(Q3_OUTPUT_FOLDER)
     q3.e(
         x,
+        alpha=2,
+        beta=2,
         number_of_trials=4,
         ks=[2, 3, 4, 7, 10],
         epsilon=1e-1,
@@ -78,6 +80,7 @@ if __name__ == "__main__":
         number_start_attempts=int(1e4),
         log_decryption_interval=100,
         log_decryption_size=60,
+        trial_decryptions_table_path=os.path.join(Q5_OUTPUT_FOLDER, "q5d-trials.csv"),
         decryptor_table_path=os.path.join(Q5_OUTPUT_FOLDER, "q5d-decrypter.csv"),
         decrypted_message_iterations_table_path=os.path.join(
             Q5_OUTPUT_FOLDER, "q5d-iterations.csv"

### 2022-11-13 src/solutions/q2.py

@@ -20,8 +20,8 @@ def _log_p_d_given_m2(x: np.ndarray):
     :return: log likelihood
     """
     n, d = x.shape
-    k = np.sum(x, axis=0).astype(int)
-    return betaln(np.sum(k) + 1, n * d - np.sum(k) + 1)
+    k = np.sum(x).astype(int)
+    return betaln(k + 1, n * d - k + 1)
 
 
 def _log_p_d_given_m3(x: np.ndarray):
@@ -31,8 +31,8 @@ def _log_p_d_given_m3(x: np.ndarray):
     :return: log likelihood
     """
     n, _ = x.shape
-    k = np.sum(x, axis=0).astype(int)
-    return logsumexp(betaln(k + 1, n - k + 1))
+    k_d = np.sum(x, axis=0).astype(int)
+    return logsumexp(betaln(k_d + 1, n - k_d + 1))
 
 
 def _log_p_model_given_data(x) -> np.ndarray:

### 2022-11-13 src/solutions/q3.py

@@ -51,7 +51,7 @@ class Theta:
         weights[1] = -1  # scale p matrix by -1 for subtraction
         return np.array(logsumexp(stacked_sum, b=weights, axis=0))
 
-    def log_pi_repeated(self, n: int):
+    def log_pi_repeated(self, n: int) -> np.ndarray:
         """
         Repeats the log_pi vector n times along axis 0
         :param n: number of repetitions
@@ -137,12 +137,17 @@ def _compute_log_pi_hat(log_responsibility: np.ndarray) -> np.ndarray:
 
 
 def _compute_log_p_matrix_hat(
-    x: np.ndarray, log_responsibility: np.ndarray
+    x: np.ndarray,
+    log_responsibility: np.ndarray,
+    alpha: float,
+    beta: float,
 ) -> np.ndarray:
     """
     Compute the log of the maximised pixel probabilities
     :param x: the image data (n, d)
     :param log_responsibility: an array of the log responsibilities of k mixture components for each image (n, k)
+    :param alpha: alpha parameter of the beta prior
+    :param beta: beta parameter of the beta prior
     :return: an array of the maximised pixel probabilities for each component (d, k)
     """
     n, d = x.shape
@@ -152,9 +157,6 @@ def _compute_log_p_matrix_hat(
     log_responsibility_repeated = np.repeat(
         log_responsibility[:, np.newaxis, :], d, axis=1
     )  # (n, d, k)
-
-    alpha = 2
-    beta = 2
 
     log_p_matrix_unnormalised_posterior = logsumexp(
         log_responsibility_repeated, b=(x_repeated + alpha - 1), axis=0
@@ -166,30 +168,41 @@ def _compute_log_p_matrix_hat(
 
     log_p_matrix_normalised_posterior = (
         log_p_matrix_unnormalised_posterior - log_p_matrix_normaliser_posterior
-    )
+    )  # (d, k)
     return log_p_matrix_normalised_posterior
 
 
-def _compute_log_m_step(x: np.ndarray, log_responsibility: np.ndarray) -> Theta:
+def _compute_log_m_step(
+    x: np.ndarray, log_responsibility: np.ndarray, alpha: float, beta: float,
+) -> Theta:
     """
     Compute the m step of expectation maximisation
     :param x: the image data (n, d)
     :param log_responsibility: an array of the log responsibilities of k mixture components for each image (n, k)
+    :param alpha: alpha parameter of the beta prior
+    :param beta: beta parameter of the beta prior
     :return: thetas optimised after maximisation step
     """
[…]

### 2022-11-13 src/solutions/q5.py

@@ -117,7 +117,8 @@ class Statistics:
     @staticmethod
     def _construct_symbols_dictionary(symbols: List[str]) -> Dict[str, int]:
         """
-        Construct a dictionary mapping each symbol to an integer
+        Construct a dictionary mapping each symbol to an integer to index the transition matrix
+        and the invariant distribution
         :param symbols: list of symbols to map
         :return: symbol to integer mapping
         """
@@ -269,8 +270,8 @@ class MetropolisHastingsDecryption:
     def _find_good_starting_decrypter(
         self,
         statistics: Statistics,
-        encrypted_message,
-        number_start_attempts,
+        encrypted_message: str,
+        number_start_attempts: int,
     ) -> Decrypter:
         """
         Find a good starting decrypter for the sampler by choosing the one with the best log likelihood
@@ -331,17 +332,17 @@ class MetropolisHastingsDecryption:
         return decrypter, logged_decryption_message
 
 
-def _construct_logged_decryptions_table(
-    logged_decryption_message, log_decryption_interval
+def _construct_decryptions_table(
+    decryption_messages: List[str], decryption_interval: int, columns: List[str]
 ) -> pd.DataFrame:
     decrypted_message_iterations_table = pd.DataFrame(
         [
-            np.arange(0, len(logged_decryption_message)) * log_decryption_interval,
-            logged_decryption_message,
+            np.arange(0, len(decryption_messages)) * decryption_interval,
+            decryption_messages,
         ]
     ).transpose()
-    decrypted_message_iterations_table.columns = ["MH Iteration", "Current Decryption"]
-    return decrypted_message_iterations_table.set_index("MH Iteration")
+    decrypted_message_iterations_table.columns = columns
+    return decrypted_message_iterations_table.set_index(columns[0])
 
 
 def a(
@@ -375,6 +376,7 @@ def d(
     number_start_attempts: int,
     log_decryption_interval: int,
     log_decryption_size: int,
+    trial_decryptions_table_path: str,
     decryptor_table_path: str,
     decrypted_message_iterations_table_path: str,
 ) -> None:
@@ -388,6 +390,7 @@ def d(
     :param number_start_attempts: number of possible starting decrypters to check
     :param log_decryption_interval: number of samples between logging the decrypted message
     :param log_decryption_size: number of symbols to decrypt when logging the decrypted message
+    :param trial_decryptions_table_path: path to store decryption messages for each trial
[…]

### 2022-11-13 submission.tex
update using log posterior

@@
 $$ \hat{\mathbf{P}}_{dk}^{MAP} = \frac{\sum_n r_{nk}(x^{(n)}_{d}+\alpha-1)}{(\alpha+\beta-1)( \sum_n r_{nk})}$$
-As a sense check, we can see when setting $\alpha=1$ and $\beta=1$ we recover $ \hat{\mathbf{P}}_{dk}^{MLE}$ as we would expect.
+As a sense check, we can see when setting $\alpha=1$ and $\beta=1$ we recover $ \hat{\mathbf{P}}_{dk}^{MLE}$ as we would expect. For the following parts, a Beta(2, 2) prior was used.
 \newpage
-\item[(d)] Plotting the posterior likelihood as a function of the iteration number for different k values:
+\item[(d)] Plotting the unnormalised posterior likelihood as a function of the iteration number for different k values:
 
@@
 
-where $epsilon$ is the stopping condition for when the log posterior converges sufficiently.
+where $epsilon$ is the stopping condition for when the unnormalised log posterior converges sufficiently.
+Note that the normalisation constant for the log posterior $\log P(\textbf{x}^{(n)}, s^{(n)} | \pi)$ is intractable and so only the unnormalised portion $\log P( \textbf{x}^{(n)}, s^{(n)} | \pi, \mathbf{P} ) + \log P(\mathbf{P})$ was computed and reported.
 
@@
 For smaller k, we can visually see that we obtain very similar solutions (a seven and a zero for $k=2$).
-For $k=3$, we get one each of zero, seven, and five, but in different orderings for different trials.
+For $k=3$, we get one each of zero, seven, and five, but in different permutations for different trials.
 However for higher K, we see that this may not always be the case.
@@
 Improvements to the model could include searching for an optimal $k$ by maximising the log posterior with regularisation on the magnitude of $k$ to balance maximising log posterior with minimising model complexity.
-Additionally, adding a prior on the responsibility components can be helpful to ensure more even mixture components unlike the components visualised here.
+Additionally, adding a prior on the responsibility components can be helpful to ensure a more even distribution across mixture components unlike the components visualised here.
 This could help promote more meaningful clusters as $k$ increases.
+Moreover, more experimentation for choosing better priors can be helpful to find better separation between mixtures.
+Finally, increasing the size of our data set (i.e.
+more images) and resolution of our images (i.e.
+more pixels) can help the model better understand the distinguishing nuances of different mixtures and provide better clustering, although the number of images and the resolution should scale together to ensure that the model doesn't learn the noise in the higher resolution images.
+This is assuming we are able to scale our computing resources.
+Finally, given that we know that there are ten digits and our current data set only includes a subset of these digits, we can also expand our data set to include all ten digits.
+Hopefully, for $k=10$, we will then be able to achieve a unique digit for each mixing component, rather than variations of repeated digits as we see now.
 

### 2022-11-13 main.py

@@ -45,8 +45,8 @@ if __name__ == "__main__":
         os.makedirs(Q3_OUTPUT_FOLDER)
     q3.e(
         x,
-        alpha=2,
-        beta=2,
+        alpha_parameter=2,
+        beta_parameter=2,
         number_of_trials=4,
         ks=[2, 3, 4, 7, 10],
         epsilon=1e-1,

### 2022-11-13 src/solutions/q3.py

@@ -3,7 +3,7 @@ from typing import List, Tuple
 
 import matplotlib.pyplot as plt
 import numpy as np
-from scipy.special import logsumexp
+from scipy.special import betaln, logsumexp
 from sklearn.manifold import TSNE
 
 from src.constants import DEFAULT_SEED
@@ -113,6 +113,39 @@ def _compute_log_likelihood(x: np.ndarray, theta: Theta) -> float:
     return np.sum(_compute_log_p_x_i_given_theta(x, theta)).item()
 
 
+def _compute_log_prior(
+    theta: Theta, alpha_parameter: float, beta_parameter: float
+) -> float:
+    """
+    Compute the prior log probability of the P matrix under a Beta prior
+    :param theta: the parameters of the model
+    :param alpha_parameter: alpha parameter of the beta prior
+    :param beta_parameter: beta parameter of the beta prior
+    :return: log_p_of_p_matrix
+    """
+    return np.sum(
+        -betaln(alpha_parameter, beta_parameter)
+        + (alpha_parameter - 1) * theta.log_p_matrix
+        + (beta_parameter - 1) * theta.log_one_minus_p_matrix
+    ).item()
+
+
+def _compute_unnormalised_log_posterior_likelihood(
+    x: np.ndarray, theta: Theta, alpha_parameter: float, beta_parameter: float
+) -> float:
+    """
+    Compute the unnormalised posterior log probability of the P matrix
+    :param x: the image data (n, d)
+    :param theta: the parameters of the model
+    :param alpha_parameter: alpha parameter of the beta prior
+    :param beta_parameter: beta parameter of the beta prior
+    :return: log_p_of_p_matrix
+    """
+    log_likelihood = _compute_log_likelihood(x, theta)
+    log_prior = _compute_log_prior(theta, alpha_parameter, beta_parameter)
+    return log_likelihood + log_prior
+
+
 def _compute_log_e_step(x: np.ndarray, theta: Theta) -> np.ndarray:
     """
     Compute the e step of expectation maximisation
@@ -139,15 +172,15 @@ def _compute_log_pi_hat(log_responsibility: np.ndarray) -> np.ndarray:
 def _compute_log_p_matrix_hat(
     x: np.ndarray,
     log_responsibility: np.ndarray,
-    alpha: float,
-    beta: float,
+    alpha_parameter: float,
+    beta_parameter: float,
 ) -> np.ndarray:
     """
     Compute the log of the maximised pixel probabilities
     :param x: the image data (n, d)
[…]

### 2022-11-14 submission.tex
small fix for figure cpation
reduce beta prior for 3 and add solutions to bonus questions for 3

@@
 $$ \hat{\mathbf{P}}_{dk}^{MAP} = \frac{\sum_n r_{nk}(x^{(n)}_{d}+\alpha-1)}{(\alpha+\beta-1)( \sum_n r_{nk})}$$
-As a sense check, we can see when setting $\alpha=1$ and $\beta=1$ we recover $ \hat{\mathbf{P}}_{dk}^{MLE}$ as we would expect. For the following parts, a Beta(2, 2) prior was used.
+As a sense check, we can see when setting $\alpha=1$ and $\beta=1$ we recover $ \hat{\mathbf{P}}_{dk}^{MLE}$ as we would expect.
+For the following parts, a very weak Beta(1+1e-5, 1+1e-5) prior was used.
 \newpage
@@
 \includegraphics[scale=0.35]{outputs/q3/q3e-0-log-pos}
-\caption{Log Likelihood vs Iteration Number}
+\caption{Unnormalised Log Posterior vs Iteration Number}
 \label{fig:3d-log-like}
@@
 
-For smaller k, we can visually see that we obtain very similar solutions (a seven and a zero for $k=2$).
+For smaller k, we can visually see that we obtain very similar solutions (a seven and a zero for $k=2$, although for trial 2, it looks more like zero and five mixed with seven).
 For $k=3$, we get one each of zero, seven, and five, but in different permutations for different trials.
 However for higher K, we see that this may not always be the case.
-For Trial 1 of $k=10$, we have three 5's whereas in Trial 3 we have four 5's.
+For Trial 1 of $k=10$, we have two 5's whereas in Trial 3 we have four 5's.
 Interestingly, different clusters of the same digits can be different, representing different variants of the written digit (i.e.
@@
 However for $k=7$ and especially $k=10$, we can see some components have very small or zero probability (i.e.
-$\pi_3$ of trial 2).
+$\pi_3$ of trial 0 and $k=10$).
 It will be unlikely for those components to represent very distinct clusters (i.e.
-the parameters for $\pi_2$ and $\pi_3$ are very similar in trial 2) This can be verified when we perform a TSNE visualisation of the responsibility vector for each of the images (Note that for $k=2$, just the responsibility vector is plotted because it is two dimensional).
+the parameters for $\pi_3$ and $\pi_4$ are very similar in trial 0 and $k=10$).
+Moreover, for these small probabilities, we see that the parameters are almost like binary images.
+This shows that there are not many images represented in these components.
+This can be verified when we perform a TSNE visualisation of the responsibility vector for each of the images (Note that for $k=2$, just the responsibility vector is plotted because it is two dimensional).
 We can see that for large k, qualitatively the number of clusters no longer matches the k value, indicating that some mixtures are redundant.
-For example for $k=7$ and $k=10$ we can only qualitatively see three to five clusters with TSNE.
+For example for $k=7$ and $k=10$ we can only qualitatively see three or four clusters with TSNE.
 
@@
 Hopefully, for $k=10$, we will then be able to achieve a unique digit for each mixing component, rather than variations of repeated digits as we see now.
+\newpage
+\item[(f)] The log-likelihood in bits can be expressed as:
+
+$$\log_2(P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\theta))$$
+
+The length of the naive encoding of these binary data is $N \cdot D$, the number of pixels in $\{\textbf{x}^{(n)}\}_{n=1}^{N}$.
+This is because the images are binary so each pixel can be represented with a single bit.
+We can compute the ratio of log-likelihood bits to the length of the naive encoding for each $k$:
+
+$$rate = 1 - \frac{\log_2(P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\theta))}{N \cdot D}$$
+
+Presenting the compression rates for different trials and $k$ values:
+
+\begin{table}[h]
+\begin{center}
+{\begin{tabular}{c|c|c|c|c}%
+ \bfseries k value & Trial 0 & Trial 1& Trial 2 & Trial 3% specify table head
+\csvreader[head to column names
+]{outputs/q3/q3e-compression.csv}{}% use head of csv as column names
+{\\\hline\csvcoli&\csvcolii&\csvcoliii&\csvcoliv&\csvcolv}% specify your coloumns here
+\end{tabular}
+}
+\caption{Compression Rates}
+\end{center}
+\end{table}
+
+As k increases, we can see that our compression rate gets better.
+This is intuitive because with higher $k$ we are specifying a more complex and expressive model (i.e.
+with more parameters) and thus we are able to capture more of the structure of the data in the model.
+Thus, the bit rate, or information provided by a sample decreases with respect to the complexity of the model and thus our compression rate increases.
+From the source coding theorem, lossless compression algorithms are lower bounded by the entropy of the underlying data generating distribution $P(\textbf{x})$.
+This is $-\sum_{x\in \mathcal{X}} P(\textbf{x})\log P(\textbf{x})$.
+On the other hand, with EM we are maximising $\langle \log P(Z, X | \theta) \rangle_{q(X)} + H[q]$ or minimising the negative of this.
+When our proposal distribution $P$ matches $q$, we recover $H[P]$ and we get the optimal model, the data generating model for encoding our data, which makes sense because we would be able to compress our data with the best possible distribution.
+This matches the lower bound of the source coding theorem.
+However, because it is very unlikely that our proposal $q$ will actually match $P$, our compression rate will always be worse than the optimal compression rate of $H[P]$.
+On the other hand, a compression algorithm would compressions on a per image basis, independent of the other images.
+And thus, it is able to attain a better compression rate for that image and is much closer to the source coding theorem lower bound.
+Depending on the data (i.e.
+$H[P]$ of the data), the compression rate of gzip can range from $60\%$ to $88\%$ (https://web.dev/optimizing-content-efficiency-optimize-encoding-and-transfer/#text-compression-with-gzip), much higher than that of our models, as we expected.
+
+\newpage
+\item[(g)] The total cost of encoding with model parameters and data:
+
+$$\log_2(P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\theta)) + M \cdot K \cdot D + M \cdot K$$
+
+Where $M$ is the cost of storing a single float value, in our case we used $float64$ so 64 bits.
+The first term is the log-likelihood as expressed in (f), the second term is the cost of storing $\mathbf{P}$, and the last term is the cost of storing $\pi$.
+The latter two terms scale with the value of k.
+This means that as $k$ increases, our compression rate deteriorates.
+Looking at the total compression ratio $\frac{log_2(P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\theta)) + M \cdot K \cdot D + M \cdot K}{N\cdot D}$ in a table:
+
+\begin{table}[h]
+\begin{center}
+{\begin{tabular}{c|c|c|c|c}%
+ \bfseries k value & Trial 0 & Trial 1& Trial 2 & Trial 3% specify table head
+\csvreader[head to column names
+]{outputs/q3/q3e-compression-total.csv}{}% use head of csv as column names
+{\\\hline\csvcoli&\csvcolii&\csvcoliii&\csvcoliv&\csvcolv}% specify your coloumns here
+\end{tabular}
+}
+\caption{Total Compression Ratios}
+\end{center}
+\end{table}
+
+We can see that the ratio is greater than one, meaning that this is actually worse than the naive encoding.
+This is due to the high cost of each value in the model parameters being 64 bits.
+However, because this remains constant with respect to our data set size, so as $N$ increases, these ratios will approach $\frac{\log_2(P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\theta)}{N\cdot D}$ and we'll recover our compression rates from part (f).
+By increasing $k$, we see that the ratio increases, and thus our compression rate is worse.
+This makes sense because we are essentially slowly storing the data into the model.
+In the extreme example of $k=N$ we can have the parameters for each mixture model as an image in the data.
+Although in this case, we could store our parameters as binary values instead of 64 bit floats.
+Our mixture component is uniform because each mixture is equally likely (all images are equally likely) so there is no need to store any values for $\pi$.
+Thus, we would recover the cost of naive encoding our data.
+Taking into account the model parameters further verifies that the cost of encoding with this approach is much higher than the cost of gzip.
 

### 2022-11-14 main.py

@@ -45,14 +45,15 @@ if __name__ == "__main__":
         os.makedirs(Q3_OUTPUT_FOLDER)
     q3.e(
         x,
-        alpha_parameter=2,
-        beta_parameter=2,
+        alpha_parameter=1 + 1e-5,
+        beta_parameter=1 + 1e-5,
         number_of_trials=4,
         ks=[2, 3, 4, 7, 10],
-        epsilon=1e-1,
+        epsilon=1e-5,
         max_number_of_steps=int(1e2),
         figure_path=os.path.join(Q3_OUTPUT_FOLDER, "q3e"),
         figure_title="Q3e",
+        compression_csv_path=os.path.join(Q3_OUTPUT_FOLDER, "q3e-compression"),
     )
 
     # Question 5

### 2022-11-14 src/solutions/q3.py

@@ -3,6 +3,7 @@ from typing import List, Tuple
 
 import matplotlib.pyplot as plt
 import numpy as np
+import pandas as pd
 from scipy.special import betaln, logsumexp
 from sklearn.manifold import TSNE
 
@@ -378,6 +379,61 @@ def _plot_log_posteriors(
     plt.savefig(figure_path)
 
 
+def _compute_compression_rate(
+    ks: List[int], log_posteriors: List[List[float]], i: int, n: int, d: int
+) -> pd.DataFrame:
+    """
+    Compute the compress rate, not taking into account the cost of storing model parameters
+    :param ks: k values to use for each trial
+    :param log_posteriors: list of vectors, each representing the log posterior during EM for a specific k
+    :param i: trial number
+    :param n: number of data points
+    :param d: number of dimensions per data point
+    :return: dataframe containing the compression rate for this trial
+    """
+    df = pd.DataFrame(
+        data=[
+            [
+                np.round(1 - (-log_posterior[-1] / (np.log(2) * n * d)), 2)
+                for log_posterior in log_posteriors
+            ]
+        ],
+        columns=ks,
+    ).T
+    df = df.reset_index()
+    df.columns = ["k value", f"Trial {i}"]
+    return df.set_index("k value")
+
+
+def _compute_total_compression_ratio(
+    ks: List[int], log_posteriors: List[List[float]], i: int, n: int, d: int
+) -> pd.DataFrame:
+    """
+    Compute the total compress ratio, taking into account the cost of storing model parameters (assuming float64)
+    :param ks: k values to use for each trial
+    :param log_posteriors: list of vectors, each representing the log posterior during EM for a specific k
+    :param i: trial number
+    :param n: number of data points
+    :param d: number of dimensions per data point
+    :return: dataframe containing the total compression ratios for this trial
+    """
+    df = pd.DataFrame(
+        data=[
+            [
+                np.round(
+                    (-log_posterior[-1] + (64 * ks[j] * (d + 1))) / (np.log(2) * n * d),
+                    2,
+                )
+                for j, log_posterior in enumerate(log_posteriors)
+            ]
+        ],
+        columns=ks,
+    ).T
+    df = df.reset_index()
+    df.columns = ["k value", f"Trial {i}"]
+    return df.set_index("k value")
+
+
 def e(
     x: np.ndarray,
     alpha_parameter: float,
[…]

### 2022-11-14 submission.tex
add edits

@@
 We can remove the integrals:
-$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_1) = (0.5)^{N\cdot D}$$
+$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_1) = 0.5^{N\cdot D}$$
 For $M_2$, we have that all pixels share some probability $p_d$ so we only need to integrate over a single variable $p_d$:
@@
 Changing the products to sums:
-$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_2) = \int_0^1  p_d^{\sum_{n=1}^{N} \sum_{d'=1}^D x_{d'}^{(n)}} (1-p_d)^{\sum_{j=1}^{N} \sum_{d'=1}^D 1-{x_{d'}^{(n)}}} d p_d$$
+$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_2) = \int_0^1  p_d^{\sum_{n=1}^{N} \sum_{d'=1}^D x_{d'}^{(n)}} (1-p_d)^{\sum_{n=1}^{N} \sum_{d'=1}^D 1-{x_{d'}^{(n)}}} d p_d$$
 Rewriting:
-$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_2) = \int_0^1  (p_d)^{K} (1- p_{d'=1})^{N\cdot D-K} d p_d$$
+$$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_2) = \int_0^1  (p_d)^{K} (1- p_{d=1})^{N\cdot D-K} d p_d$$
 where $K =\sum_{n=1}^{N} \sum_{d'=1}^D x_{d'}^{(n)}$.
@@
 $$P( \{\textbf{x}^{(n)}\}_{n=1}^{N}|M_3) = \prod_{d=1}^D \frac{K_d! (N-K_d)!}{(N+1)!}$$
-where $K_d = \sum_{n=1}^{N} x_d^{(n}$.
+where $K_d = \sum_{n=1}^{N} x_d^{(n)}$.
 
@@
 
-We can see that for models specified to have the same parameter value for all pixels, like $M_1$, is very unlikely with the given data set.
-This makes sense because it is specifying models where the image is essentially blank (a uniform shade), which is not reflective of our digit images.
+We can see that for models specified to have the same parameter value for all pixels, like $M_1$, are very unlikely with the given data set.
+This makes sense because it is specifying models where the image is essentially a uniform shade, which is not reflective of our digit images.
 Moreover, $M_1$ specifies a specific value of 0.5 for all the parameters whereas $M_2$ specifies any value for all the parameters as long as it's the same.
@@
 The log-likelihood $\mathcal{L}(\textbf{x}^{(n)}|\theta)$ can be expressed in vector form:
-$$\mathcal{L}(\textbf{x}^{(n)}|\theta) = \log \sum_{k=1}^{K}  \pi_k \exp\bigg(\textbf{x}^{(n)}\log(\mathbf{P}_{k}) + (1-\textbf{x}^{(n)})\log(1-\mathbf{P}_{k})\bigg) $$
-which can be further vectorised using Python scipy's $logsumexp$ operation.
+$$\mathcal{L}(\textbf{x}^{(n)}|\theta) = \log \sum_{k=1}^{K}  \pi_k \exp\bigg((\textbf{x}^{(n)})^T \log(\mathbf{P}_{k}) + (1-\textbf{x}^{(n)})^T \log(1-\mathbf{P}_{k})\bigg) $$
+where $\mathbf{P}_{k}$ is the $k^{th}$ column of $\mathbf{P}$. This can can be further vectorised using Python scipy's $logsumexp$ operation.
 
 Moreover, the log-likelihood $\mathcal{L}(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\theta)$ can be expressed:
-$$\mathcal{L}(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\theta) = \sum_{i=1}^{N} \Bigg( \log \sum_{k=1}^{K}  \pi_k \exp\bigg(\textbf{x}^{(n)}\log(\mathbf{P}_{k}) + (1-\textbf{x}^{(n)})\log(1-\mathbf{P}_{k})\bigg) \Bigg)$$
+$$\mathcal{L}(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\theta) = \sum_{i=1}^{N} \Bigg( \log \sum_{k=1}^{K}  \pi_k \exp\bigg((\textbf{x}^{(n)})^T\log(\mathbf{P}_{k}) + (1-\textbf{x}^{(n)})^T\log(1-\mathbf{P}_{k})\bigg) \Bigg)$$
 
@@
 $$P(s^{(n)} = k|\textbf{x}^{(n)}, \pi, \mathbf{P}) \propto  P(\textbf{x}^{(n)} | s^{(n)} = k, \pi, \mathbf{P}) P(s^{(n)} = k|\pi, \mathbf{P})$$
-where $s^{(n)} \in \{1,...,K\}$ a discrete hidden variable with $P(s^{(n)} = k|\textbf{x}^{(n)}, \pi) = \pi_k$.
-Note that $P(s^{(n)} = k|\pi, \mathbf{P}) = P(s^{(n)} = k| \pi)$ as $s^{(n)$ isn't dependent on $\mathbf{P}$.
+where $s^{(n)} \in \{1,...,K\}$ a discrete hidden variable with $P(s^{(n)} = k| \pi) = \pi_k$.
+Note that $P(s^{(n)} = k|\pi, \mathbf{P}) = P(s^{(n)} = k| \pi)$ as $s^{(n)}=k$ isn't dependent on $\mathbf{P}$.
 
@@
 $$\Bigg\langle \sum_n \log P(\textbf{x}^{(n)}, s^{(n)}| \pi, \mathbf{P})\Bigg\rangle_{q(\{s^{(n)}\})} = \sum_{n=1}^N q(s^{(n)}) \log P(\textbf{x}^{(n)}, s^{(n)}| \pi, \mathbf{P})$$
-Let this quantity be $E$. For each term of $E$:
+Let this quantity be $E$. For each term of the summation in $E$:
 $$q(s^{(n)}) = \textbf{r}_{n}^T$$
@@
 For $k=3$, we get one each of zero, seven, and five, but in different permutations for different trials.
-However for higher K, we see that this may not always be the case.
+However for higher k, we see that this may not always be the case.
 For Trial 1 of $k=10$, we have two 5's whereas in Trial 3 we have four 5's.
@@
 Moreover, looking at the responsibilities of each mixture component, we can see that when k is relatively small they are relatively evenly distributed.
-However for $k=7$ and especially $k=10$, we can see some components have very small or zero probability (i.e.
+However for $k=7$ and especially $k=10$, we can see some components have very small probability (i.e.
 $\pi_3$ of trial 0 and $k=10$).
@@
 Moreover, more experimentation for choosing better priors can be helpful to find better separation between mixtures.
-Finally, increasing the size of our data set (i.e.
+Increasing the size of our data set (i.e.
 more images) and resolution of our images (i.e.
@@
 This is because the images are binary so each pixel can be represented with a single bit.
-We can compute the ratio of log-likelihood bits to the length of the naive encoding for each $k$:
+We can compute compression rate with respect to the ratio of log-likelihood bits to the length of the naive encoding for each $k$:
 
@@
 On the other hand, with EM we are maximising $\langle \log P(Z, X | \theta) \rangle_{q(X)} + H[q]$ or minimising the negative of this.
-When our proposal distribution $P$ matches $q$, we recover $H[P]$ and we get the optimal model, the data generating model for encoding our data, which makes sense because we would be able to compress our data with the best possible distribution.
+When our proposal distribution $P$ matches $q$, we recover $H[P]$ and we get the optimal model, the data generating model for encoding our data.
+This makes sense because we would be able to compress our data with the best possible distribution to represent the data.
 This matches the lower bound of the source coding theorem.
@@
 \newpage
-\item[(g)] The total cost of encoding with model parameters and data:
+\item[(g)] The total cost of encoding with model parameters and data is:
 
@@
 Where $M$ is the cost of storing a single float value, in our case we used $float64$ so 64 bits.
-The first term is the log-likelihood as expressed in (f), the second term is the cost of storing $\mathbf{P}$, and the last term is the cost of storing $\pi$.
+The first term is the log-likelihood as expressed in part (f), the second term is the cost of storing $\mathbf{P}$, and the last term is the cost of storing $\pi$.
 The latter two terms scale with the value of k.
@@
 We can see that the ratio is greater than one, meaning that this is actually worse than the naive encoding.
-This is due to the high cost of each value in the model parameters being 64 bits.
-However, because this remains constant with respect to our data set size, so as $N$ increases, these ratios will approach $\frac{\log_2(P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\theta)}{N\cdot D}$ and we'll recover our compression rates from part (f).
+This is due to the high cost of storing each value in the model parameters being 64 bits.
+However, because this remains constant with respect to our data set size, as $N$ increases these ratios will approach $\frac{\log_2(P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\theta)}{N\cdot D}$ and we'll recover our compression rates from part (f).
 By increasing $k$, we see that the ratio increases, and thus our compression rate is worse.
@@
 $$\Psi(\alpha, \beta) = \frac{N_{\alpha, \beta}}{N_{\beta}}$$
-where $N_{\alpha, \beta}$ is the count of the number of occurrences of the pair $(\alpha, \beta)$, where $\beta$ is followed by $s\alpha$ in the text and $N_{\beta}$ is the number of occurrences of $\beta$.
+where $N_{\alpha, \beta}$ is the count of the number of occurrences of the pair $(\alpha, \beta)$, where $\beta$ is before $\alpha$ in the text and $N_{\beta}$ is the number of occurrences of $\beta$.
 Moreover to ensure ergodicity, a one was added to each $N_{\alpha, \beta}$.
@@
   \item[(ii)] Repeat $\phi^{(i+1)} = \Psi \phi^{(i)}$
-  \item[(iii)] Terminate when $\phi^{(i+1) - \phi^{(i)} < \epsilon$
+  \item[(iii)] Terminate when $\phi^{(i+1)} - \phi^{(i)} < \epsilon$
 \end{enumerate}
 
-where $\Psi \in \mathbf{R}^{53\times53}$ containing the transition probabilities, $\Psi_{i, j} = P(s_j | s_i)$ where $s_i$ is the $i^{th}$ symbol and $s_j$ is the $j^{th}$ symbol, and $\epsilon$ is some small number indicating sufficient convergence of the distribution to be considered stationary.
+where $\Psi \in \mathbf{R}^{53\times53}$ containing the transition probabilities, $\Psi_{i, j} = P(s^j | s^i)$ where $s^i$ is the $i^{th}$ symbol and $s^j$ is the $j^{th}$ symbol with respect to the indices of symbols in the $\Psi$ matrix, and $\epsilon$ is some small number indicating sufficient convergence of the distribution to be considered stationary.
 The function $\phi(\gamma)$ is simply the index of symbol $\gamma$ in the vector $\phi$.
@@
 $$A(\sigma \rightarrow \sigma'| \mathcal{D}) = \min \{1, \frac{S(\sigma' \rightarrow \sigma)P(\sigma'|\mathcal{D})}{S(\sigma \rightarrow \sigma')P(\sigma|\mathcal{D})})\}$$
-because $S(\sigma \rightarrow \sigma')$ is the conditional transition probability of $\sigma'$ given $\sigma$ and $\mathcal{D}$ is our encrypted text $e_1, e_2,...,e_n$.
+where $S(\sigma \rightarrow \sigma')$ is the conditional transition probability of $\sigma'$ given $\sigma$ and $\mathcal{D}$ is our encrypted text $e_1, e_2,...,e_n$.
 
-$S(\sigma \rightarrow \sigma') = S(\sigma' \rightarrow \sigma)$ for all $\sigma$ and $\sigma'$ that differ only at two spots because the probability in this case will always be $\frac{1}{{53 \choose 2}}$, we can simplify:
+$S(\sigma \rightarrow \sigma') = S(\sigma' \rightarrow \sigma)$ for all $\sigma$ and $\sigma'$ that differ only at two spots because the probability in this case will always be $\frac{1}{{53 \choose 2}}$, so we can simplify:
 $$A(\sigma \rightarrow \sigma'| \mathcal{D}) = \min \{1, \frac{P(\sigma'|\mathcal{D})}{P(\sigma|\mathcal{D})})\}$$
@@
 However, in practise it would be unlikely to be able to make these assumptions about symbol frequencies, especially with the finite size of our training set and encrypted message.
+Therefore in practise, symbol probabilities alone would not be sufficient.
 
@@
 Our training text would need to increase quadratically to maintain the same ratio of possible transitions to example transitions (number of first order transitions in a text of length $N$ is $N-1$ and second order its $N-2$).
-This can also introduce sparsity (as in small non-zero probabilities in many entries) in our transition matrix despite our ability to maintain ergodicity with small probabilities to prevent non-zero entries.
-However, the log-likelihood of many areas of $\sigma$ space might be very small or all just the same, when all the transition probabilities are just the offset probability (added to maintain ergodicity).
-Navigating this space will be much more difficult for the sampler as the space could be relatively flat.
+This can also introduce sparsity (in this case, small non-zero probabilities because ergodicity is maintained) in our transition tensor.
+Thus, the log-likelihood of many areas of $\sigma$ space might be very small or the same as their neighbours, when the transition probabilities are mostly just the offset probability added to maintain ergodicity.
+Navigating this space will be much more difficult for the sampler.
 
@@
 $$\sup_{\big\{\textbf{x}\in \mathbb{R}^n \bigm| \| \textbf{x}\|=1\big\}} \frac{\textbf{x}^T\textbf{Ax}}{\|\textbf{x}\|^2}$$
-Because $\| \textbf{x}\|=1$, we can equivalently write:
+Because $\| \textbf{x}\|=1$ (i.e. choosing $\textbf{w} \in \mathbb{R}$ and $\textbf{x}=\frac{\textbf{w}}{\|\textbf{w}\|}$ to reformulate the problem over the unit sphere), we can equivalently write:
 $$\sup_{\big\{\textbf{x}\in \mathbb{R}^n \bigm| \| \textbf{x}\|=1\big\}} \textbf{x}^T\textbf{Ax}}$$
@@
 so we can write:
-$$\textbf{x}^T\textbf{Ax} =\bigg(\sum_{i=1}^n (\xi_i^T \textbf{x}) \xi_i^T\bigg)\textbf{A}\bigg(\sum_{i=1}^n (\xi_i^T \textbf{x}) \xi_i\bigg)$$
-Given that $\xi_i$ are eigenvectors of $\textbf{A}$ corresponding to eigenvalues $\lambda_i$:
+$$\textbf{x}^T\texRtbf{Ax} =\bigg(\sum_{i=1}^n (\xi_i^T \textbf{x}) \xi_i^T\bigg)\textbf{A}\bigg(\sum_{i=1}^n (\xi_i^T \textbf{x}) \xi_i\bigg)$$
+Given that $\xi_i$R are eigenvectors of $\textbf{A}$ corresponding to eigenvalues $\lambda_i$:
 $$\textbf{x}^T\textbf{Ax} =\bigg(\sum_{i=1}^n (\xi_i^T \textbf{x}) \xi_i^T\bigg)\bigg(\sum_{i=1}^n \lambda_i(\xi_i^T \textbf{x}) \xi_i\bigg)$$
@@
 $$\textbf{x}^T\textbf{Ax} =\sum_{i=1}^n \lambda_i(\xi_i^T \textbf{x})^2$$
-From our above reformulation with the unit sphere, we know that $\|\textbf{x}\|^2 = 1$ so $\|\textbf{x}\|^2 = \sum_{j=1}^{n} x_i^2 = \sum_{j=1}^{n} (\xi_j \textbf{x})^2 = 1$.
-Thus the quantity $\sum_{i=1}^n \lambda_i(\xi_i^T \textbf{x})^2$ is a weighted average of $\lambda_i$'s, which is always less than or equal to the largest $\lambda_i$ value so:
+From our above reformulation with the unit sphere, we know that $\|\textbf{x}\|^2 = 1$ so $\|\textbf{x}\|^2 = \sum_{i=1}^{n} x_i^2 = \sum_{j=i}^{n} (\xiij \textbf{x})^2 = 1$.
+Thus the quantity $\sum_{i=1}^n \lambda_i(\xi_i^T \textbf{x})^2$ is a weighted average of $\lambda_i$'s with weights $(\xi_i^T \textbf{x})^2$, which is always less than or equal to the largest $\lambda_i$ value so:
 $$\textbf{x}^T\textbf{Ax} =\sum_{i=1}^n \lambda_i(\xi_i^T \textbf{x})^2 \leq \lambda_1$$
@@
 $$\textbf{x} = \sum_{i=k+1}^n (\xi_i^T \textbf{x}) \xi_i$$
-Using the same argument as in (b) we can bound $\textbf{x}^T\textbf{Ax}}$:
+Using the same line of argument as in part (b) we can bound $\textbf{x}^T\textbf{Ax}}$:
 $$\textbf{x}^T\textbf{Ax}} = \sum_{i=k+1}^n \lambda_i(\xi_i^T \textbf{x})^2 \leq \max\{\lambda_{k+1}, ..., \lambda_n\}$$

### 2022-11-14 submission.tex
insert missing dimensionality for R

@@
 $$\sup_{\big\{\textbf{x}\in \mathbb{R}^n \bigm| \| \textbf{x}\|=1\big\}} \frac{\textbf{x}^T\textbf{Ax}}{\|\textbf{x}\|^2}$$
-Because $\| \textbf{x}\|=1$ (i.e. choosing $\textbf{w} \in \mathbb{R}$ and $\textbf{x}=\frac{\textbf{w}}{\|\textbf{w}\|}$ to reformulate the problem over the unit sphere), we can equivalently write:
+Because $\| \textbf{x}\|=1$ (i.e. choosing $\textbf{w} \in \mathbb{R}^n$ and $\textbf{x}=\frac{\textbf{w}}{\|\textbf{w}\|}$ to reformulate the problem over the unit sphere), we can equivalently write:
 $$\sup_{\big\{\textbf{x}\in \mathbb{R}^n \bigm| \| \textbf{x}\|=1\big\}} \textbf{x}^T\textbf{Ax}}$$

### 2022-11-14 submission.tex
remove extra R

@@
 $$\textbf{x}^T\texRtbf{Ax} =\bigg(\sum_{i=1}^n (\xi_i^T \textbf{x}) \xi_i^T\bigg)\textbf{A}\bigg(\sum_{i=1}^n (\xi_i^T \textbf{x}) \xi_i\bigg)$$
-Given that $\xi_i$R are eigenvectors of $\textbf{A}$ corresponding to eigenvalues $\lambda_i$:
+Given that $\xi_i$ are eigenvectors of $\textbf{A}$ corresponding to eigenvalues $\lambda_i$:
 $$\textbf{x}^T\textbf{Ax} =\bigg(\sum_{i=1}^n (\xi_i^T \textbf{x}) \xi_i^T\bigg)\bigg(\sum_{i=1}^n \lambda_i(\xi_i^T \textbf{x}) \xi_i\bigg)$$

### 2022-11-14 submission.tex
insert missing negative in log bits

@@
 
-$$\log_2(P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\theta))$$
+$$-\log_2(P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\theta))$$
 
@@
 
-$$rate = 1 - \frac{\log_2(P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\theta))}{N \cdot D}$$
+$$rate = 1 - \frac{-\log_2(P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\theta))}{N \cdot D}$$
 
@@
 
-$$\log_2(P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\theta)) + M \cdot K \cdot D + M \cdot K$$
+$$-\log_2(P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\theta)) + M \cdot K \cdot D + M \cdot K$$
 
@@
 This means that as $k$ increases, our compression rate deteriorates.
-Looking at the total compression ratio $\frac{log_2(P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\theta)) + M \cdot K \cdot D + M \cdot K}{N\cdot D}$ in a table:
+Looking at the total compression ratio $\frac{-log_2(P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\theta)) + M \cdot K \cdot D + M \cdot K}{N\cdot D}$ in a table:
 
@@
 This is due to the high cost of storing each value in the model parameters being 64 bits.
-However, because this remains constant with respect to our data set size, as $N$ increases these ratios will approach $\frac{\log_2(P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\theta)}{N\cdot D}$ and we'll recover our compression rates from part (f).
+However, because this remains constant with respect to our data set size, as $N$ increases these ratios will approach $\frac{-\log_2(P(\{\textbf{x}^{(n)}\}_{n=1}^{N}|\theta)}{N\cdot D}$ and we'll recover our compression rates from part (f).
 By increasing $k$, we see that the ratio increases, and thus our compression rate is worse.
